\begin{abstract}
The Euclid--Mullin sequence is defined by $a(0)=2$,
$a(n{+}1) = \text{lpf}(a(0)\cdots a(n)+1)$, where $\text{lpf}$ is
the least prime factor.
Mullin's Conjecture (MC, 1963) asserts that every prime eventually
appears.  We present a Lean~4 formalization
(${\sim}78{,}000$~lines, 148~files, zero~\texttt{sorry}) that reduces MC
to a single open hypothesis and explores five variant selection rules.

The main result is the machine-verified chain
\[
  \mathrm{DSL}
  \;\Longrightarrow\;
  \mathrm{CME}
  \;\Longrightarrow\;
  \mathrm{CCSB}
  \;\Longrightarrow\;
  \mathrm{MC},
\]
where \textbf{CME} (Conditional Multiplier Equidistribution) states
that $\minFac$ destroys correlation between consecutive multiplier
residues mod~$q$, and \textbf{DSL} (Deterministic Stability Lemma)
asserts that population-level equidistribution of least prime
factors transfers to the specific EM orbit.
An inductive bootstrap recasts MC as walk-hitting on
$(\ZZ/q\ZZ)^\times$ with SubgroupEscape proved free;
a character-analytic bridge handles all character orders;
ensemble averaging over squarefree starting points collapses
the landscape to DSL\@.

The \emph{factor tree} $\mathcal{T}(m)$---rooted at~$m$, branching at
every prime factor of~$P{+}1$---provides a unifying framework for
variants: the standard walk is the leftmost path, and
$\mathrm{GenMixedMC}(m)$ asks that every prime appear somewhere
in~$\mathcal{T}(m)$.  Under PSCD (Population Sieve Confinement
Decay), a density-$1$ set of squarefree starting points satisfies
GenMixedMC; the full chain PEAP $\Rightarrow$ PSCD is
machine-checked with no residual open hypothesis:
a weak unconditional CRT bound (sqfreeCount $\geq X/4$)
absorbs the sieve constant since $\prod(1-1/p)\to 0$.
The $\{\minFac, \operatorname{secondMinFac}\}$ two-point variant
reduces to a single spectral hypothesis~(UFDStrong); for $q = 3$,
faithful character escape closes this unconditionally.
The function field analog over~$\FF_p[t]$, where the Weil bound gives
population equidistribution for free, confirms that the residual
difficulty is orbit-specificity transfer, not analytic number theory.
Over 138~dead ends are documented.

\medskip\noindent\textbf{Statements and proofs.}
Theorems marked {\color{leanink}\sffamily\footnotesize$\checkmark$\,name}
are formally verified in Lean~4; clicking the identifier links to the
source code.
Most proofs are sketched rather than fully reported;
correctness is guaranteed by the type-checker.
Some misalignment may exist between what is stated here and what is
proved in the formalization; the Lean source is the authoritative reference.
To mitigate this risk, we used
LeanArchitect~\cite{LeanArchitect2026} to cross-check the 20~key
declarations against the paper's theorem statements;
this process uncovered one missing precondition, now corrected.
All published results and open hypotheses are tagged with
\texttt{@[publish]} and \texttt{@[open\_point]} annotations in the
GitHub repository (\url{https://github.com/marcellop71/EM}), using the CA
content-addressing framework for Lean~4; a machine-readable registry
is available in the \texttt{registry/} folder.
\end{abstract}

