%% =========================================================================
%% Function field analog --- merged from former companion paper
%% =========================================================================

%% Local macros (kept here so this file is self-contained)
\providecommand{\Fp}{\FF_p}
\providecommand{\Fpt}{\FF_p[t]}
\providecommand{\GF}[2]{\FF_{#1^{#2}}}
\providecommand{\ffseq}{\mathrm{ffSeq}}
\providecommand{\ffprod}{\mathrm{ffProd}}
\providecommand{\irredcount}{\pi_p}
\providecommand{\FFMC}{\mathrm{FF\text{-}MC}}
\providecommand{\FFDH}{\mathrm{FF\text{-}DH}}
\providecommand{\FCD}{\mathrm{FCD}}
\providecommand{\SPV}{\mathrm{SPV}}
\providecommand{\PSCDff}{\mathrm{PSCD}}
\providecommand{\CME}{\mathrm{CME}}
\providecommand{\CCSB}{\mathrm{CCSB}}
\providecommand{\DSL}{\mathrm{DSL}}

\subsection{The Function Field Analog}
\label{sec:function-field}

\begin{roadmap}
Replacing $\ZZ$ by $\Fpt$ and primes by monic irreducible polynomials
gives a clean test bed where \emph{the analytic number theory becomes
free}.  The Weil bound---an open hypothesis over~$\ZZ$---is a theorem of
Weil~\cite{Weil1948} over~$\Fpt$.  Yet the function-field Mullin
Conjecture is still open, and the residual difficulty is identical to
the integer case: the orbit-specificity barrier.  This subsection
develops the FF-EM sequence, proves an unconditional six-step sieve
chain and the $\Phi_3$ exclusion criterion, and presents three formally
verified dead ends that delineate the boundary of current techniques.
The function-field module of the formalization comprises 17~files and
${\sim}7{,}372$~lines with zero \texttt{sorry}.
\end{roadmap}

\paragraph{Setting and definitions.}

Let $p$ be a prime.  The ring $\Fpt$ is a Euclidean domain, hence a UFD;
\emph{monic irreducible polynomials} play the role of prime numbers.  We
write $\irredcount(d)$ for the number of monic irreducibles of degree~$d$
over~$\Fp$.  For~$p=2$ the first values are
$\irredcount(1) = 2$ (namely $t$ and $t{+}1$),
$\irredcount(2) = 1$ (namely $t^2{+}t{+}1$), $\irredcount(3) = 2$,
$\irredcount(4) = 3$.

\begin{definition}[FF-EM Sequence ---
\lean{EM/FunctionField/Analog.lean}{FFEMData}]
\label{def:ffem}
Fix a prime~$p$.  An \defname{FF-EM Data} consists of:
$\ffseq(0) = t$ (the smallest monic irreducible);
$\ffprod(0) = t$;
$\ffseq(n{+}1)$ a monic irreducible factor of $\ffprod(n)+1$ of
\emph{minimal degree};
$\ffprod(n{+}1) = \ffprod(n)\cdot\ffseq(n{+}1)$.
\end{definition}

\begin{conjecture}[Function-Field Mullin Conjecture ---
\lean{EM/FunctionField/Analog.lean}{FFMullinConjecture}]
\label{conj:ffmc}
For every monic irreducible $Q \in \Fpt$, there exists $n$ with
$\ffseq(n) = Q$.
\end{conjecture}

Two structural differences from the integer setting are immediate.
First, monic irreducibles of the same degree have no canonical total
order, so the ``minimal-degree'' rule may select among several
\emph{competitors}---the analog of the integer $\minFac$ rule loses
uniqueness.  Second, the unit group
$(\Fpt/(Q))^\times \cong \GF{p}{\deg Q}^\times$ is cyclic of order
$p^{\deg Q}-1$, providing richer group-theoretic structure than
$(\ZZ/q\ZZ)^\times$.

\paragraph{Why study the function-field case?}
The Weil bound for character sums over $\Fp$-curves is a
\emph{theorem}~\cite{Weil1948,Deligne1980}, not a conjecture.  This
makes population equidistribution of monic irreducibles across residue
classes unconditional, in contrast to the integer setting where the
analogous result requires the Weighted PNT in arithmetic progressions.
By isolating the \emph{orbit-specificity barrier}---the gap between
``population statistics'' and ``the specific walk
$\ffprod(n) \bmod Q$ visits the residue class $-1$''---we will show
that this barrier is the sole remaining obstacle to FF-MC, and that it
is structurally identical to the obstacle for MC over~$\ZZ$.

\paragraph{Basic structural facts.}
The following hold unconditionally for any FF-EM data:

\begin{lemma}[Coprimality cascade ---
\lean{EM/FunctionField/Analog.lean}{coprimality\_cascade}]
\label{lem:ff-coprime}
No monic irreducible $Q$ divides both $\ffprod(n)$ and $\ffprod(n)+1$.
\end{lemma}

\begin{proof}
If $Q$ divides both, then $Q \mid 1$, contradicting $\deg Q \geq 1$.
\end{proof}

\begin{lemma}[Sequence injectivity ---
\lean{EM/FunctionField/Bootstrap.lean}{ffSeq\_injective\_proved}]
\label{lem:ff-injective}
The map $n \mapsto \ffseq(n)$ is injective: no irreducible appears
twice.
\end{lemma}

\begin{lemma}[Degree growth ---
\lean{EM/FunctionField/Analog.lean}{ffProd\_natDegree\_strict\_mono}]
\label{lem:ff-degree-growth}
$\deg(\ffprod(n+1)) > \deg(\ffprod(n))$ for all~$n$.
\end{lemma}

\begin{lemma}[Pool degree to infinity ---
\lean{EM/FunctionField/Finiteness.lean}{ffSeq\_degree\_tendsto\_atTop\_unconditional}]
\label{lem:ff-pool}
$\deg(\ffprod(n)+1) \to \infty$ as $n \to \infty$.
\end{lemma}

\paragraph{The necklace identity.}

The combinatorial heart of the unconditional ANT chain over $\Fpt$ is
the following exact identity.

\begin{theorem}[Necklace identity ---
\lean{EM/FunctionField/NecklaceFormula.lean}{necklace\_identity\_proved}]
\label{thm:necklace}
For every $n \geq 1$,
\[
  \sum_{d \mid n}\, d\cdot\irredcount(d) \;=\; p^n.
\]
\end{theorem}

The name comes from the combinatorial interpretation: $\irredcount(d)$
counts aperiodic necklaces of length~$d$.  The identity is classical
(Moreau, 1872), and is equivalent to the factorization
$t^{p^n}-t = \prod_{\deg Q \mid n} Q$ over $\Fpt$.

\begin{proof}[Proof sketch]
Both sides count the $p^n$ elements of $\GF{p}{n}$, viewed as roots of
$t^{p^n}-t$.  The forward inclusion uses
$Q \mid t^{p^d}-t \mid t^{p^n}-t$ when $\deg Q \mid n$; the reverse
uses the tower law on $[\Fp(\alpha):\Fp]$ to bound the degree of any
minimal polynomial.  The Lean proof spans ${\sim}380$~lines and
exercises Mathlib's Galois-field infrastructure
(\code{GaloisField.card}, \code{minpoly.irreducible},
\code{IntermediateField.finrank\_dvd\_of\_le\_right},
\code{FiniteField.X\_pow\_card\_pow\_sub\_X\_ne\_zero}).
\end{proof}

\begin{corollary}[Density bounds ---
\lean{EM/FunctionField/NecklaceFormula.lean}{necklace\_count\_mul\_le}]
\label{cor:irred-bounds}
For every $n \geq 1$, $\;n\cdot\irredcount(n) \leq p^n$ and
$\irredcount(n) \leq p^n/n$.
\end{corollary}

A finer Möbius inversion gives
$\irredcount(n) = \frac{1}{n}\sum_{d \mid n}\mu(n/d)p^d
  \sim p^n/n + O(p^{n/2}/n)$;
the crude bound suffices for our sieve chain.

\paragraph{The unconditional sieve chain.}

The integer sieve approach to Mullin's Conjecture proceeds via
\[
  \text{Char.\ Cancel.}
  \xRightarrow{\text{GRH}} \text{PNT-AP}
  \xRightarrow{\text{S--W}} \FCD
  \Rightarrow \SPV \Rightarrow \PSCDff
  \Rightarrow \text{Almost-All MC}.
\]
Over $\Fpt$, \emph{every arrow is unconditional}.

\smallskip
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lll}
\toprule
\textbf{Step} & \textbf{Over $\ZZ$} & \textbf{Over $\Fpt$} \\
\midrule
Char.\ cancellation & GRH & Exact (necklace) \\
PNT-in-APs & Siegel--Walfisz & Exact ($\irredcount(d)\geq 1$) \\
$\FCD$ & WPNT-in-APs & Linear supply \\
$\SPV$ & FCD + sparsity & Necklace bound \\
$\PSCDff$ & FCD + SPV & Density bound \\
Almost-all MC & PSCD + pigeonhole & PNT + necklace \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Step 1 --- Character cancellation.}
\begin{theorem}[FF Character Sum Cancellation ---
\lean{EM/FunctionField/FFCharacterSums.lean}{ff\_char\_sum\_cancellation\_proved}]
The necklace identity (Theorem~\ref{thm:necklace}) holds unconditionally,
giving exact cancellation rather than an asymptotic bound.
\end{theorem}

Over $\ZZ$, the analogous statement is the Riemann Hypothesis for
Dirichlet $L$-functions; over $\Fpt$ it is an identity of natural
numbers.

\paragraph{Step 2 --- PNT in arithmetic progressions.}
\begin{theorem}[FF PNT-in-APs ---
\lean{EM/FunctionField/Finiteness.lean}{ffIrredCount\_pos}]
For every $d \geq 1$, $\irredcount(d) \geq 1$.
\end{theorem}

The proof uses the Galois-field construction: the minimal polynomial of
a primitive element of $\GF{p}{d}$ is monic irreducible of degree~$d$.
The reduction to character cancellation is automatic
(\lean{EM/FunctionField/FFCharacterSums.lean}{ff\_char\_cancel\_implies\_pnt}).

\paragraph{Step 3 --- Forbidden class divergence.}
\begin{property}[FF-FCD ---
\lean{EM/FunctionField/FFCharacterSums.lean}{FFFCD}]
$\sum_{d=1}^{D}\irredcount(d) \geq D$ for all $D \geq 1$.
\end{property}

\begin{theorem}[FF-FCD is unconditional ---
\lean{EM/FunctionField/FFCharacterSums.lean}{ff\_fcd\_proved}]
Each summand is $\geq 1$ by Step~2; the bound is immediate
(\lean{EM/FunctionField/FFCharacterSums.lean}{ff\_pnt\_implies\_fcd}).
\end{theorem}

\paragraph{Step 4 --- Sieve product vanishing.}
\begin{property}[FF-SPV ---
\lean{EM/FunctionField/FFSieve.lean}{FFSieveProductVanishing}]
$n\cdot\irredcount(n) \leq p^n$ for all $n \geq 1$.
\end{property}

\begin{theorem}[FF-SPV is unconditional ---
\lean{EM/FunctionField/FFSieve.lean}{ff\_spv\_proved}]
Immediate from Theorem~\ref{thm:necklace}: the term $n\cdot\irredcount(n)$
is one summand and the total equals $p^n$.  Reduction to FCD:
\lean{EM/FunctionField/FFSieve.lean}{ff\_fcd\_implies\_spv}.
\end{theorem}

\paragraph{Step 5 --- Population sieve confinement decay.}
\begin{property}[FF-PSCD ---
\lean{EM/FunctionField/FFSieve.lean}{FFPSCD}]
$\irredcount(n) \leq p^n/n$ for all $n \geq 1$.
\end{property}

\begin{theorem}[FF-PSCD is unconditional ---
\lean{EM/FunctionField/FFSieve.lean}{ff\_pscd\_proved}]
Divide both sides of Step~4 by~$n$.  Reduction:
\lean{EM/FunctionField/FFSieve.lean}{ff\_spv\_implies\_pscd}.
\end{theorem}

\paragraph{Step 6 --- Almost-all generalized mixed MC.}
A \emph{mixed} selection at each step picks \emph{any} irreducible factor
of $\ffprod(n)+1$ (not necessarily minimal-degree); the
\emph{generalized} variant starts from an arbitrary monic squarefree
polynomial.  ``Almost-all'' refers to density-one over starting points.

\begin{theorem}[Almost-All FF-GenMixedMC is unconditional ---
\lean{EM/FunctionField/FFSieve.lean}{ff\_almost\_all\_unconditional}]
\label{thm:ff-almost-all}
For every prime $p$, the conjunction of FF-PNT-in-APs and FF-SPV holds
unconditionally; in particular, a density-$1$ set of monic squarefree
starting polynomials satisfies generalized mixed MC.
\end{theorem}

\begin{keyresult}
\textbf{Full chain witness}
(\lean{EM/FunctionField/FFSieve.lean}{ff\_full\_chain\_unconditional}):
the conjunction
\[
  \text{Char.\ Cancel.} \,\wedge\, \text{PNT-AP} \,\wedge\, \FCD
  \,\wedge\, \SPV \,\wedge\, \PSCDff \,\wedge\,
  \text{Almost-All FF-GenMixedMC}
\]
holds unconditionally for every prime~$p$.  Zero open hypotheses.
\end{keyresult}

\paragraph{The factor tree and structural prerequisites.}
Independent of the sieve chain, the factor-tree structure (any valid
\emph{mixed} selection) admits the following formally verified
properties:

\begin{theorem}[Factor-tree structure ---
\lean{EM/FunctionField/Master.lean}{theorem\_A\_factor\_tree}]
\label{thm:ff-A}
For any valid mixed selection from any monic squarefree start,
\begin{enumerate}[nosep,label=(\alph*)]
\item the walk explores infinitely many distinct monic irreducibles
  (\lean{EM/FunctionField/StochasticMC.lean}{ff\_mixed\_explores\_infinitely\_many});
\item the factor pool degree $\deg(\mathrm{acc}(n)+1) \to \infty$
  (\lean{EM/FunctionField/PopulationEquidist.lean}{ff\_factor\_pool\_degree\_grows});
\item selection is injective
  (\lean{EM/FunctionField/FactorTree.lean}{ffMixedSel\_injective});
\item walk products are injective
  (\lean{EM/FunctionField/FactorTree.lean}{ffMixedWalkProd\_injective});
\item the start is never recapturable
  (\lean{EM/FunctionField/FactorTree.lean}{start\_not\_capturable}).
\end{enumerate}
\end{theorem}

\paragraph{Stochastic FF-MC and the phase transition.}

\begin{theorem}[Stochastic FF-MC ---
\lean{EM/FunctionField/StochasticMC.lean}{stochastic\_mc\_unconditional}]
\label{thm:ff-stoch}
For any monic irreducible target~$Q$ and any valid mixed walk, the
factor pool eventually has degree exceeding $\deg(Q)$; with any positive
probability $\varepsilon > 0$ of non-greedy selection, every target is
eventually captured.
\end{theorem}

\begin{theorem}[Phase transition at $\varepsilon = 0$ ---
\lean{EM/FunctionField/StochasticMC.lean}{ff\_phase\_transition\_unconditional}]
\label{thm:ff-phase}
For $\varepsilon > 0$, stochastic MC captures every monic irreducible.
For $\varepsilon = 0$, the walk is the standard greedy FF-EM and capture
of every irreducible is exactly $\FFMC$.  The phase transition is sharp.
\end{theorem}

\paragraph{The bootstrap: from dynamical hitting to FF-MC.}

\begin{property}[FF Dynamical Hitting (FFDH) ---
\lean{EM/FunctionField/Analog.lean}{FFDynamicalHitting}]
\label{prop:ffdh}
For every FF-EM data and every monic irreducible $Q$ not in the
sequence, there exist cofinally many indices~$n$ at which
$Q \mid \ffprod(n)+1$.
\end{property}

\begin{theorem}[One-hypothesis reduction ---
\lean{EM/FunctionField/Bootstrap.lean}{ff\_dh\_implies\_ffmc}]
\label{thm:ff-bootstrap}
$\FFDH \;\Longrightarrow\; \FFMC$ (using the unconditional finiteness of
irreducibles per degree,
\lean{EM/FunctionField/Finiteness.lean}{ff\_dh\_implies\_ffmc\_unconditional}).
\end{theorem}

The proof is by strong induction on degree.  At degree~$d_0$, the
hypothesis gives MC below~$d_0$; a uniform sieve gap exhausts all
sub-degree irreducibles past some $N_0$; FFDH provides cofinally many
hits past $N_0$; injectivity makes each hit consume a distinct
\emph{competitor} of the same degree; finiteness per degree then forces
$Q$ itself to be selected (\emph{competitor exhaustion},
\lean{EM/FunctionField/Bootstrap.lean}{ff\_element\_capture}).  This is
the function-field analog of the integer bootstrap
\code{dynamical\_hitting\_implies\_mullin}; the only difference is the
final exhaustion step, which over $\ZZ$ collapses because primes are
totally ordered.

\paragraph{The \texorpdfstring{$\Phi_3$}{Phi3} exclusion criterion.}

We now develop a structural obstruction with no integer analog.  When
all irreducible factors of $\ffprod(n)+1$ have degree~$1$, the walk
$\ffprod(n) \bmod Q$ for a degree-one target $Q = (t-a)$ follows the
\emph{autonomous map} $f(w) = w(w+1)$ on $\Fp$.  The target $Q$ divides
$\ffprod(n)+1$ iff the walk position $w_n = -1$.  Hence degree-one
capture reduces to: \emph{does the orbit of $f$ reach~$-1$?}

\begin{definition}[Autonomous map ---
\lean{EM/FunctionField/AutonomousMap.lean}{ffAutonomousMap}]
\label{def:autonomous}
For a prime $p$, $f \colon \Fp \to \Fp$ is given by $f(w) = w(w+1)$.
The \emph{orbit} from $a \in \Fp$ is $a, f(a), f^2(a), \ldots$
\end{definition}

\begin{lemma}[Absorption at zero ---
\lean{EM/FunctionField/AutonomousMap.lean}{ffAutonomousMap\_neg\_one}]
$f(-1) = 0$ and $f(0) = 0$; once the orbit hits $0$ it stays there
(\lean{EM/FunctionField/AutonomousMap.lean}{ffAutonomousOrbit\_absorbed}).
\end{lemma}

\begin{lemma}[$\Phi_3$ criterion ---
\lean{EM/FunctionField/AutonomousMap.lean}{ffAutonomousMap\_eq\_neg\_one\_iff}]
\label{lem:phi3}
$f(w) = -1 \iff w^2+w+1 = 0$.  The preimage of $-1$ is the root set of
the third cyclotomic polynomial $\Phi_3(w)$.
\end{lemma}

\begin{lemma}[Cube-root condition ---
\lean{EM/FunctionField/AutonomousMap.lean}{pow\_three\_eq\_one\_of\_phi3}]
If $w^2+w+1 = 0$ in $\Fp$, then $w^3 = 1$.
\end{lemma}

\begin{theorem}[$\Phi_3$ has no roots when $p \equiv 2 \pmod 3$ ---
\lean{EM/FunctionField/AutonomousMap.lean}{phi3\_no\_roots}]
\label{thm:phi3-no-roots}
For $p \equiv 2 \pmod 3$ with $p \geq 5$, $\;w^2 + w + 1 \neq 0$ for all
$w \in \Fp$.
\end{theorem}

\begin{proof}
Suppose $w^2+w+1 = 0$ in $\Fp$.  Then $w^3 = 1$, and $w \neq 0,1$ (each
substitution gives a contradiction with $p \geq 5$).  Hence $\mathrm{ord}(w) = 3$
in $\Fp^\times$, so $3 \mid p-1$ by Lagrange.  But $p \equiv 2 \pmod 3$
forces $p - 1 \equiv 1 \pmod 3$, contradiction.
\end{proof}

\begin{theorem}[Unreachability of $-1$ ---
\lean{EM/FunctionField/AutonomousMap.lean}{ff\_neg\_one\_unreachable}]
\label{thm:ff-unreach}
For $p \equiv 2 \pmod 3$ with $p \geq 5$ and any starting value
$a \neq -1$ in $\Fp$, the orbit $f^n(a)$ never equals $-1$.
\end{theorem}

\begin{corollary}[Visited at most once ---
\lean{EM/FunctionField/AutonomousMap.lean}{ff\_neg\_one\_visited\_at\_most\_once}]
For the same primes, $-1$ is visited only at step~$0$ (and only if
$a = -1$); for all $n \geq 1$, the orbit value is never $-1$.
\end{corollary}

\begin{remark}
The criterion fails to apply for $p = 2$ (where $\Phi_3 = t^2+t+1$ is
itself irreducible) and $p = 3$ (where $\Phi_3$ has the double root
$w = 1$).  Primes $p \equiv 2 \pmod 3$ with $p \geq 5$ include
$5, 11, 17, 23, 29, 41, 47, \dots$
\end{remark}

\paragraph{Integer non-analog.}
Over $\ZZ$ the multiplier $a_{n+1} \bmod q$ is \emph{not} determined by
$P(n) \bmod q$ alone---it depends on the global factorization of
$P(n)+1$.  No autonomous map exists, so the $\Phi_3$ criterion has no
direct integer analog.

\paragraph{The cyclotomic structure (Dead End \#129).}
Over $\FF_2$, the product of all monic irreducibles of degree dividing
$d=2$ is $t(t+1)(t^2+t+1) = t^4+t$, and adding $1$ gives
$\Phi_5(t) = t^4 + t + 1$.  Its Galois group is
$\mathrm{Gal}(\FF_{2^4}/\FF_2) \cong \ZZ/4\ZZ$---abelian, not~$S_4$.
More generally, the classical identity
$\prod_{\deg Q \mid d}Q = t^{p^d}-t$ forces the greedy FF-EM to build
products dividing $t^{p^d}-t$ for some~$d$, so $\ffprod(n)+1$ has
splitting field inside $\GF{p}{N}$ for some~$N$ and the Galois group is
\emph{cyclic}, generated by Frobenius.  Thus the ``large monodromy''
hypothesis is structurally false for the greedy FF-EM, witnessed by
\lean{EM/FunctionField/Analog.lean}{ff\_cyclotomic\_dead\_end}.

\paragraph{The orbit-specificity boundary.}

We now establish that the orbit-specificity barrier is the sole
remaining obstacle to FF-MC, and that it is structurally identical to
the barrier over~$\ZZ$.  The central dichotomy is between two types of
equidistribution:

\begin{itemize}[nosep,leftmargin=1em]
\item \emph{Population equidistribution}: among all monic irreducibles
  of degree~$d$, residues mod~$Q$ are equidistributed across
  $|\GF{p}{\deg Q}^\times|$ classes, with error $O(p^{d/2})$ from the
  Weil bound.  \textbf{Unconditional} over $\Fpt$.
\item \emph{Orbit equidistribution}: the specific walk
  $\ffprod(n) \bmod Q$ visits each residue class with frequency
  $1/(p^{\deg Q}-1)$ (the function-field DSL).  \textbf{Open}.
\end{itemize}

\begin{theorem}[Population CCSB from Weil ---
\lean{EM/FunctionField/MultiplierCCSB.lean}{weil\_implies\_population\_mult\_ccsb}]
The Weil bound proves population-level multiplier CCSB unconditionally;
\lean{EM/FunctionField/MultiplierCCSB.lean}{selection\_bias\_chain}
formalizes the population~$\to$~fiber~$\to$~selection~$\to$~orbit
hierarchy.
\end{theorem}

\paragraph{Dead End \#127 --- Weil bound insufficient.}

\begin{theorem}[Weil does not close the gap ---
\lean{EM/FunctionField/OrbitBarrier.lean}{dead\_end\_127\_weil\_not\_sufficient}]
\label{thm:de127}
The Weil bound implies population equidistribution but does not imply
$\FFDH$, because:
(i) Weil controls $\sum_{f\;\mathrm{monic},\;\deg f = d}\chi(f \bmod Q)$,
i.e.\ averages over \emph{all} polynomials of degree~$d$;
(ii) $\FFDH$ requires $\sum_{n<N}\chi(\ffprod(n) \bmod Q)$, a sum along
a \emph{specific deterministic orbit};
(iii) $\ffprod(n)+1$ is the unique element of its congruence class, not
a random draw.
\end{theorem}

\paragraph{Dead End \#129 --- Large monodromy structurally false.}

\begin{property}[FF Large Monodromy (FFLM) ---
\lean{EM/FunctionField/Analog.lean}{FFLargeMonodromy}]
The Galois group of $\ffprod(n)+1$ eventually contains the alternating
group: $|\mathrm{Gal}| \geq (\deg \ffprod(n))!/2$ for all sufficiently
large~$n$.
\end{property}

\begin{theorem}[FFLM is structurally false ---
\lean{EM/FunctionField/OrbitBarrier.lean}{dead\_end\_129\_monodromy\_dead}]
\label{thm:de129}
The greedy FF-EM produces products dividing $t^{p^d}-t$.  Adding $1$
gives a polynomial whose splitting field lies in $\GF{p}{N}$ for some
$N$, hence Galois group $\cong \ZZ/N\ZZ$, generated by Frobenius---
\emph{abelian}, not $A_d$.  The monodromy approach reduces to abelian
Galois theory $=$ Dirichlet characters $=$ Weil $=$ population
equidistribution; the route is circular.
\end{theorem}

\paragraph{Dead End \#130 --- Generation $\neq$ coverage.}
A natural hope is that subgroup escape (multipliers generate the full
group) might imply dynamical hitting (the walk visits every element).
Both PRE (over~$\ZZ$) and Weil (over~$\Fpt$) prove subgroup escape.
The following counterexample shows this is insufficient.

\begin{theorem}[Generation $\neq$ coverage ---
\lean{EM/FunctionField/OrbitBarrier.lean}{dead\_end\_130\_generation\_ne\_coverage}]
\label{thm:de130}
On $\ZZ/4\ZZ$, the steps $\{1, 3\}$ generate the full group, yet the
\emph{alternating} walk $0, 1, 0, 1, \ldots$ visits only $\{0, 1\}$,
missing $\{2, 3\}$.  Witnessed in
\lean{EM/FunctionField/CyclicWalkCoverage.lean}{alternating\_walk\_misses\_two};
the converse direction (cancellation $\Rightarrow$ coverage via Weyl)
is
\lean{EM/FunctionField/CyclicWalkCoverage.lean}{walk\_cancel\_implies\_covers}.
\end{theorem}

\paragraph{The barrier theses.}

\begin{theorem}[Orbit barrier thesis ---
\lean{EM/FunctionField/OrbitBarrier.lean}{orbit\_barrier\_thesis}]
\label{thm:barrier-thesis}
The orbit-specificity barrier has the following formally verified
properties:
\begin{enumerate}[nosep,label=(\roman*)]
\item All structural prerequisites for FF-MC are unconditional:
  injectivity, degree growth, finiteness per degree, coprimality
  cascade.
\item $\FFDH \Rightarrow \FFMC$ (Theorem~\ref{thm:ff-bootstrap}).
\item $\FFDH$ is the sole remaining hypothesis.
\item The barrier is not closed by population statistics
  (Theorem~\ref{thm:de127}).
\item The barrier is not closed by Galois theory
  (Theorem~\ref{thm:de129}).
\item The barrier is not closed by algebraic generation
  (Theorem~\ref{thm:de130}).
\item The barrier dissolves under stochastic perturbation
  ($\varepsilon > 0$, Theorem~\ref{thm:ff-stoch}).
\item The barrier is sharp at $\varepsilon = 0$
  (Theorem~\ref{thm:ff-phase}).
\end{enumerate}
\end{theorem}

\begin{theorem}[Setting independence ---
\lean{EM/FunctionField/OrbitBarrier.lean}{barrier\_setting\_independent}]
\label{thm:setting-independent}
\begin{enumerate}[nosep,label=(\roman*)]
\item Over $\Fpt$: ANT is free (Weil $\Rightarrow$ PE).
\item Over $\Fpt$: orbit specificity remains ($\FFDH$ open;
  $\FFDH \Rightarrow \FFMC$).
\item Over $\Fpt$: stochastic perturbation dissolves the barrier;
  the deterministic limit does not.
\end{enumerate}
\end{theorem}

The function-field setting eliminates the ``easy'' obstacle (ANT for
PE) and adds new structural information ($\Phi_3$ exclusion), but the
``hard'' obstacle---orbit specificity~$=$~DSL---is identical to the
integer case.  This proves that the difficulty of Mullin's Conjecture
lies \emph{entirely} in the orbit-specificity barrier and not in the
analytic number theory.

\paragraph{Side-by-side summary.}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lcc}
\toprule
\textbf{Aspect} & \textbf{Over $\ZZ$} & \textbf{Over $\Fpt$} \\
\midrule
Population equidist.\ (PE) & Open (WPNT) & \textbf{Free} (Weil) \\
Subgroup escape (SE) & Proved (PRE) & Proved (Weil) \\
Coprimality cascade & Proved & Proved \\
$\CME \Rightarrow \MC$ chain & Proved & Proved \\
Orbit specificity (DSL) & \textbf{Open} & \textbf{Open} \\
Walk character sums & Not standard & Not standard \\
Stochastic MC ($\varepsilon > 0$) & Proved & Proved \\
$\Phi_3$ exclusion & --- & \textbf{Proved} \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Open problems specific to FF-MC.}

We close with a few open problems particular to the function-field
analog (the integer counterparts are catalogued in~\S\ref{sec:hard}).

\begin{frontierbox}
\textbf{Deterministic FF-MC.}
Does every monic irreducible over $\Fpt$ eventually appear in the greedy
FF-EM sequence?  Equivalently, does $\FFDH$ hold?  Our work shows that
$\FFDH$ is the sole remaining hypothesis: the stochastic version
($\varepsilon > 0$) is proved (Theorem~\ref{thm:ff-stoch}), and the
open content is exactly the $\varepsilon = 0$ boundary.
\end{frontierbox}

\begin{frontierbox}
\textbf{Controlling the $\varepsilon \to 0$ limit.}
Stochastic FF-MC is proved for all $\varepsilon > 0$.  Can the limit
$\varepsilon \to 0$ be controlled to deduce deterministic FF-MC?
Specifically: does the capture time $T_Q(\varepsilon)$ for a target~$Q$
remain bounded as $\varepsilon \to 0$?  The phase transition is
sharp---the proof structure for $\varepsilon > 0$ relies on non-greedy
exploration, which is absent at $\varepsilon = 0$.
\end{frontierbox}

\begin{frontierbox}
\textbf{Large monodromy beyond cyclotomic starts.}
Dead End~\#129 shows the \emph{greedy} FF-EM produces cyclotomic (hence
abelian) products.  Does the obstruction persist for non-greedy starting
points?  Cohen's theorem~\cite{Cohen1981} shows a random monic
polynomial of degree~$d$ has Galois group $S_d$ with probability
tending to~$1$.  Is the recursive FF-EM structure forcing cyclotomic
products an artifact of the greedy minimal-degree selection?
\end{frontierbox}

\begin{frontierbox}
\textbf{Integer analog of the $\Phi_3$ criterion.}
The $\Phi_3$ exclusion (Theorem~\ref{thm:ff-unreach}) uses the absence
of cube roots of unity in $\Fp$.  Over $\ZZ$ the walk does not follow
an autonomous map---can systematic exclusion criteria be developed
nonetheless for specific $(q,a)$ pairs?  Cox--van der
Poorten~\cite{CoxVdP1968} suggested structural reasons for the
difficulty of the prime~$41$; the function-field $\Phi_3$ result is a
formal analog.
\end{frontierbox}

\begin{frontierbox}
\textbf{Selection bias neutrality}
(\lean{EM/FunctionField/MultiplierCCSB.lean}{SelectionBiasNeutral}).
Does greedy minimal-degree selection preserve character-sum cancellation?
A positive answer would close the orbit-specificity gap for both FF-MC
and MC simultaneously, since the question concerns the selection
mechanism, not the ambient ring.  This is the cleanest formulation of
the barrier; the integer analog is DSL.
\end{frontierbox}

\paragraph{Headline summary.}
The function-field portion of the formalization is collected in the
landscape theorem
\lean{EM/FunctionField/Master.lean}{ff\_unconditional\_grand\_summary},
which assembles theorems~A, B, C, D
(\lean{EM/FunctionField/Master.lean}{theorem\_A\_factor\_tree},
 \lean{EM/FunctionField/Master.lean}{theorem\_B\_stochastic\_mc},
 \lean{EM/FunctionField/Master.lean}{theorem\_C\_phase\_transition},
 \lean{EM/FunctionField/Master.lean}{theorem\_D\_sieve\_chain})
into a single conjunction with zero open hypotheses.  Together they
witness:
\begin{itemize}[nosep,leftmargin=1em]
\item the unconditional six-step sieve chain
  (Theorem~\ref{thm:ff-almost-all});
\item stochastic FF-MC and the $\varepsilon = 0$ phase transition
  (Theorems~\ref{thm:ff-stoch}, \ref{thm:ff-phase});
\item the necklace identity (Theorem~\ref{thm:necklace});
\item the $\Phi_3$ exclusion criterion (Theorem~\ref{thm:ff-unreach});
\item the orbit-barrier and setting-independence theses
  (Theorems~\ref{thm:barrier-thesis}, \ref{thm:setting-independent}).
\end{itemize}
The function-field treatment thus delivers the cleanest available
diagnosis: \emph{the difficulty of Mullin's Conjecture is the
orbit-specificity barrier, not the analytic number theory}.
