%% =========================================================================
\section{Introduction}
\label{sec:intro}
%% =========================================================================

Euclid's proposition IX.20 of the \emph{Elements} shows that for any
finite set of primes, each prime factor of their product plus one is outside the
set~\cite{Hardy2009}: to grow your set of primes, you can pick any of them. The \textbf{Euclid--Mullin
sequence} (OEIS A000945), introduced by Mullin~\cite{Mullin1963}, makes a
definite choice: always take the \emph{smallest} prime factor.
\begin{align}
  a(0) &= 2, &
  a(n+1) &= \text{smallest prime factor of } \bigl(a(0) \cdots a(n) + 1\bigr).
  \label{eq:em-def}
\end{align}
The first twenty terms (0-indexed) are
\begin{gather*}
  \underbrace{2}_{a(0)},\;
  3,\; 7,\; 43,\; 13,\; 53,\;
  \underbrace{5}_{a(6)},\;
  \underbrace{6221671}_{a(7)},\;
  38709183810571, \\
  139,\; 2801,\;
  \underbrace{11}_{a(11)},\;
  17,\; 5471,\; 52662739,\; 23003,\;
  30693651606209, \\
  \underbrace{37}_{a(17)},\;
  1741,\;
  \underbrace{1313797957}_{a(19)},\; \ldots
\end{gather*}
The sequence shows an erratic behavior: small primes appear out of their
natural order ($5$ not until position~$6$, $11$ at position~$11$,
$37$ at position~$17$), while enormous primes---a 7-digit number at
position~$7$, a 14-digit number at position~$8$---appear early.  As of
2025, only 51~terms are known and some primes like $41$ and $47$ are
not yet observed.  Computing further terms requires \emph{complete}
factorization of $\Prod{2}(n)+1$ (where $\Prod{2}(n) \coloneqq a(0)\cdots a(n)$
is the running product): finding any prime factor is not
enough---one must certify that no smaller factor exists.  Since
$\Prod{2}(n)$ grows super-exponentially, this quickly exceeds the reach
of all known factoring algorithms.  By construction, no prime can
appear twice.

\begin{conjecture}[Mullin, 1963]\label{conj:mullin}
Every prime number eventually appears in the Euclid--Mullin sequence.
\end{conjecture}

The conjecture has resisted proof for over sixty years.  The difficulty
is showing that the deterministic $\minFac$ rule eventually \emph{selects} each prime.
That the rule matters is not idle speculation: Cox and van der
Poorten~\cite{CoxVdP1968} showed that replacing $\minFac$ with the
\emph{largest} prime factor provably misses infinitely many primes.

Each step couples the next prime to the full factorization history, creating a recursive dependency that
defeats both probabilistic heuristics and standard sieve methods.

At each step, $\Prod{2}(n)+1$ may have many prime factors, and just the \emph{smallest} is chosen.
So, a target prime~$q$ may possibly divide $\Prod{2}(n)+1$ for many (maybe infinite) $n$,
yet never be selected if a smaller prime always divides $\Prod{2}(n)+1$ as well.
Our formalization tries to make this tension precise.

\paragraph{The accumulator structure.}
The running product $\Prod{2}(n)$ is an \emph{accumulator}: a single
number that commits to the entire sequence history via irreversible
multiplication, analogous to the cumulative digest in a hash chain.
The accumulator poses a challenge~$\Prod{2}(n)+1$; the response
$\minFac(\Prod{2}(n)+1)$ extends the chain; and the updated accumulator
$\Prod{2}(n{+}1) = \Prod{2}(n) \cdot \minFac(\Prod{2}(n){+}1)$ absorbs the
response irreversibly---once a prime enters the product, it divides
every future product and can never appear again
(Theorem~\ref{thm:extinct}).

This accumulator coupling is what makes standard tools fail.  Sieve
methods require approximate independence between the events
``$p \mid m$'' for different~$m$; here successive challenges
$\Prod{2}(n)+1$ share a cumulative history.  Ergodic methods require a
fixed or state-dependent map; here the map at step~$n$ depends on the
full accumulator, not just the current residue.  The walk
reformulation (Section~\ref{sec:walk}) tames this coupling by
projecting the accumulator onto a finite group: $\walkZ{2}{q}{n} =
\Prod{2}(n) \bmod q$ preserves the divisibility information relevant
to~$q$ while discarding the accumulator's combinatorial complexity.
But the information lost in this projection is exactly the source of
difficulty: the walk position determines \emph{whether} $q$ can divide
$\Prod{2}(n)+1$; the full accumulator determines \emph{which} prime is
actually selected.  Every open hypothesis in this paper addresses this gap.

\paragraph{The factoring channel.}
The accumulator structure of the EM sequence is closely analogous to a
cryptographic hash chain: each term is deterministically derived from
its predecessor through an operation---integer factorization---that
destroys algebraic relationships.  In an iterated SHA-256 chain
$x_0, H(x_0), H(H(x_0)), \ldots$, each value is uniquely determined
by the seed, yet consecutive terms pass every reasonable statistical
test for independence.  In the EM sequence, the running product
$\Prod{2}(n)$ determines $\Prod{2}(n)+1$, whose smallest prime factor
becomes the next multiplier; but the $O(\log q)$ bits visible in the
residue $\Prod{2}(n) \bmod q$ cannot control the outcome of factoring the
${\sim}2^n$-bit integer $\Prod{2}(n)+1$.  This \emph{information
bottleneck} is why the multiplier residues behave as if
independent---and why the walk on $(\ZZ/q\ZZ)^\times$ should visit
every element, including the ``death state'' $-1$ that would put~$q$
in the sequence.  The analogy is tighter than it may appear: in both
cases the pseudorandomness claim concerns a fully deterministic process
whose forward map is easy but whose outputs resist structural
prediction---the difference being that $\minFac$ achieves this not by
cryptographic design but accidentally, through the information
bottleneck of projecting a ${\sim}2^n$-bit integer onto $O(\log q)$
bits.  Our formalization makes this intuition precise: we
identify the exact mathematical content---Conditional Multiplier
Equidistribution (CME)---needed to convert ``the factoring channel destroys
correlation'' into a proof of MC, and show it implies the conjecture
through a verified chain of reductions.

\

Our main result is the machine-verified reduction of MC to a single
open hypothesis:

\begin{theorem}[\lean{EM/Reduction/Master.lean}{full\_chain\_dsl}]
\label{thm:main-chain}
$\mathrm{DSL} \;\Longrightarrow\;
 \mathrm{CME} \;\Longrightarrow\;
 \mathrm{CCSB} \;\Longrightarrow\; \MC$.
\end{theorem}

\noindent The Deterministic Stability Lemma~(DSL) is the sole
remaining hypothesis.  It asserts that \emph{population
equidistribution} of least prime factors among shifted squarefree
integers---a standard consequence of Alladi's
theorem~\cite{Alladi1977} and Dirichlet's theorem on primes in
arithmetic progressions---implies \emph{conditional multiplier
equidistribution} (CME) for the specific EM orbit.  In symbols,
$\mathrm{DSL} = (\mathrm{PE} \Rightarrow \mathrm{CME})$, where PE is
the population equidistribution that follows from analytic number theory.
The entire mathematical content of this paper is the gap between PE
(known) and CME (sufficient for MC): does generic equidistribution
transfer to one deterministic orbit?

The strategy proceeds in three stages:

\

\begin{enumerate}[nosep]
\item \textbf{Reformulation.}  We recast ``does prime~$q$ appear?'' as
  ``does a multiplicative walk on the cyclic group $(\ZZ/q\ZZ)^\times$
  hit the element~$-1$?''  This translation is exact
  (Section~\ref{sec:walk}).
\item \textbf{Bootstrap.}  We show that the algebraic precondition for
  the walk to reach~$-1$---that the multipliers generate the full
  group---is \emph{free}, following from the inductive hypothesis
  that all primes below~$p$ have already appeared
  (Section~\ref{sec:bootstrap}).  A single hit on~$-1$ past the
  stage where all smaller primes are excluded suffices for MC\@.
\item \textbf{Reduction to DSL.}  We develop the character-analytic
  infrastructure to show CME~$\Rightarrow$~CCSB~$\Rightarrow$~MC
  (Section~\ref{sec:character}), then an ensemble averaging framework
  that collapses the full landscape to DSL
  (Section~\ref{sec:ensemble}).  Section~\ref{sec:hard} explains
  why the remaining hypothesis is hard.
\end{enumerate}

\

The formalization serves two purposes:
(i)~it guarantees that every reduction is logically sound, and (ii)~it
precisely delineates the boundary between what is proved and what
remains open, preventing the kind of subtle gap that plagues pencil-and-paper
reductions involving multiple interacting hypotheses.

\

The intermediate reductions are:

\begin{theorem}[\defname{Single Hit Theorem} --- \lean{EM/Equidist/Bootstrap.lean}{single\_hit\_implies\_mc}]
\label{thm:main}
$\mathrm{SingleHitHypothesis} \;\Longrightarrow\; \MC$.
\end{theorem}

SingleHitHypothesis asks: for every missing prime~$q$, if every prime
below~$q$ already appears in the sequence and the multiplier residues
generate the full group $(\ZZ/q\ZZ)^\times$ (SubgroupEscape,
Property~\ref{def:se}), then $q \mid \Prod{2}(n)+1$ for some~$n$ past
the stage where all smaller primes are excluded.  The proof is by strong
induction on~$p$: the inductive hypothesis ensures all primes below~$p$
have appeared, PrimeResidueEscape (proved elementarily) bootstraps
SubgroupEscape, and a single hit gives $\seq{2}(n{+}1) = p$.

The algebraic precondition---that the multipliers generate
$(\ZZ/q\ZZ)^\times$---is free; the open problem is purely dynamical.
Multiple strategies for producing the required hit are formally verified:

\begin{theorem}[\lean{EM/Equidist/Bootstrap.lean}{dynamical\_hitting\_implies\_mullin}]
\label{thm:dh-main}
DynamicalHitting (DH: the walk hits $-1$ cofinally)
$\;\Longrightarrow\; \MC$.
\end{theorem}

\begin{theorem}[\lean{EM/Equidist/SelfCorrecting.lean}{complex\_csb\_mc'}]
ComplexCharSumBound (CCSB: character sums of the walk are $o(N)$)
$\;\Longrightarrow\; \MC$.
\end{theorem}

\begin{theorem}[\lean{EM/LargeSieve/Spectral.lean}{cme\_implies\_mc}]
$\mathrm{CME} \;\Longrightarrow\; \MC$ \emph{(sharpest sufficient condition)}.
\end{theorem}

\noindent An independent route bypasses CME entirely:
Visit Equidistribution~(VE)---the walk visits every
residue class with equal frequency---also implies MC through a
separate machine-verified chain
(\code{sve\_implies\_mc}, \code{cancel\_weyl\_implies\_mc}).
VE and CME are genuinely different conditions (neither implies
the other), but both reach MC\@.

All reductions are fully machine-verified with zero \code{sorry}.
The ensemble averaging framework
(Section~\ref{sec:ensemble}) uses generalized EM sequences parametrized
by arbitrary squarefree starting points, CRT-based decorrelation, and a
first-moment/variance argument to collapse
all routes to DSL\@.  The formalization thus
provides a precise ``roadmap'': prove DSL (the sole remaining
hypothesis), or equivalently any of DH, CCSB, CME, VE, or the
equivalent formulations catalogued in \S\ref{sec:character},
\S\ref{sec:open}, and the glossary (Appendix~\ref{sec:glossary}),
and the rest follows by machine-checked deduction.

\paragraph{Organization.}
Section~\ref{app:bag} analyzes the bag-theoretic and
information-theoretic structure of the EM dynamics.
Section~\ref{app:agents} describes the human--AI methodology.
The technical development begins in Section~\ref{sec:walk}, which
reformulates MC as a walk-hitting problem,
establishes the algebraic prerequisites, and derives the first missing prime's death channel avoidance.
Section~\ref{sec:bootstrap} presents the inductive bootstrap---the
core insight that SubgroupEscape is free---and proves the Single Hit
Theorem, reducing MC to producing one hit at each prime.
Section~\ref{sec:character} develops the character-analytic reduction
(CCSB~$\Rightarrow$~MC), including the large sieve infrastructure, the
spectral energy bridge, and the van der Corput--autocorrelation route.
Section~\ref{sec:ensemble} develops the ensemble averaging framework,
reducing MC to a single Deterministic Stability Lemma~(DSL).
Section~\ref{sec:hard} explains \emph{why} the remaining hypothesis is
difficult by analyzing dead ends and structural barriers.
Section~\ref{sec:open} identifies the open problems whose resolution
would close the conjecture and assesses the evidence for and against.
Section~\ref{sec:variants} studies variants of the conjecture:
the random-factor variant (easy by mixing), the
$\{\minFac, \operatorname{secondMinFac}\}$ two-point variant
(almost closed under UFDStrong, \S\ref{sec:minfac-secondminfac}),
and the function field analog over~$\FF_p[t]$
(\S\ref{sec:function-field}).
Section~\ref{sec:lean} describes the Lean formalization and
catalogues all verified results (\S\ref{sec:verified-results}).
Appendix~\ref{app:history} collects the historical background;
Appendix~\ref{app:analogies} discusses analogies with
Artin's conjecture, multiplicative walks, Sarnak's program, and
sieve theory;
Appendix~\ref{app:routes} presents additional sieve and spectral routes to MC;
and Appendix~\ref{sec:glossary} provides a glossary of all definitions and hypotheses.

