%% =========================================================================
\section{Open problems and summary of formally verified results}
\label{sec:open}
%% =========================================================================

The formalization certifies the reductions above and pinpoints the
frontier.  This section identifies the open problems whose resolution
would close Mullin's Conjecture.

\paragraph{CCSB as the Precise Frontier}

The formalization identifies \textbf{ComplexCharSumBound} as the
irreducible analytic content.  The question:

\begin{center}
\emph{Are the walk character sums $\sum_{n<N} \chi(\walkZ{2}{q}{n})$ bounded
$o(N)$ for every non-trivial $\chi$?}
\end{center}

CCSB is a single hypothesis that implies MC with no additional
conditions.  As shown in Section~\ref{sec:character}, CCSB rules out
permanent avoidance of~$-1$: the Fourier bridge forces the hit count
at~$-1$ to be $N/(q{-}1) + o(N)$, contradicting the zero hits that
the first missing prime's walk must achieve past the sieve gap.

The walk telescoping identities (Section~\ref{sec:character}) provide
precise structural constraints.  The identity
$\sum_n \chi(w(n))(\chi(m(n))-1) = O(1)$ means that the walk sum
$S_N$ and the multiplier sum $M_N = \sum_n \chi(m(n))$ satisfy
$S_N \approx S_N + (M_N - S_N) = M_N + O(1)$ only in the crude sense;
the telescoping does \emph{not} separate them.

\paragraph{Connection to Bombieri--Vinogradov}

A Bombieri--Vinogradov type result for EM walk residues would give:
\[
\sum_{\substack{q \leq Q \\ q\text{ prime}}}
\max_{a}
\Bigl| |\{n \leq N : w(n) \equiv a \pmod{q}\}|
  - \frac{N}{q-1} \Bigr|
\;\ll\; \frac{NQ}{(\log N)^A}.
\]
For non-exceptional primes, the hit count at~$-1$ would be
$\sim N/(q{-}1) > 0$, contradicting permanent avoidance; the finitely
many exceptional primes below~$Q_0$ are handled by the inductive
bootstrap (\S\ref{sec:bootstrap}), closing the conjecture.

The difficulty is that BV applies to the set of \emph{all} primes,
not to a specific subsequence.  The EM walk is deterministic and
self-referential: the walk at step~$n$ depends on the factorization
of~$\Prod{2}(n)+1$, which depends on all previous walk values.  Standard
BV does not apply.

\paragraph{Connection to Chebotarev}

The \textbf{EffectiveKummerEscape} hypothesis asserts: for each
prime~$\ell$, there exists~$B$ such that for $q \geq B$ with
$\ell \mid q\!-\!1$, some multiplier among the first~$B$ escapes the
$\ell$-th power kernel.  This is a Chebotarev-type statement for the
Kummer extension $\mathbb{Q}(\zeta_\ell, 3^{1/\ell}, \ldots,
53^{1/\ell})$: the Frobenius at~$q$ determines which multiplier
primes are $\ell$-th power residues.

An effective Chebotarev density theorem~\cite{LagariasOdlyzko1977}
for this fixed number field
would give EKE for all but finitely many~$q$ (effectively bounded).
Combined with finite verification for the remaining~$q$, this would
prove PRE and hence SE unconditionally---but SE is
\emph{already} proved unconditionally via the elementary PRE.
The Chebotarev approach would give a stronger \emph{effective}
bound on how quickly SE kicks in, refining the density argument of
\S\ref{sec:bootstrap}.

\paragraph{The Sieve-Theoretic Approach}

\textbf{Mertens\-Escape}: for any prime~$q$ and proper
subgroup~$H$, infinitely many primes outside~$H$ exist
(Dirichlet content).
\textbf{Sieve\-Amplification}: Mertens escape should force
eventual $\minFac(\Prod{2}(n){+}1)$ escape from~$H$, via
super-exponential growth and mutual coprimality of
successive Euclid numbers.

The formally verified chain:
$\mathrm{MertensEscape} + \mathrm{SieveAmplification}
\xrightarrow{\text{proved}}
\mathrm{TailSE} \xrightarrow{\text{proved}}
\mathrm{CofinalEscape} \xrightarrow{\text{proved}}
\mathrm{QuotientDH}$.

The formalization articulates a richer sieve infrastructure in two parallel
routes:

\paragraph{Cumulative route.}
\begin{gather*}
\mathrm{PDE}
\xrightarrow{\text{Alladi}}
\mathrm{GLPFE}
\xrightarrow{\mathrm{SieveTransfer}}
\mathrm{SieveEquidist}
\xrightarrow{\text{open}}
\mathrm{NoLongRuns} \\
\xrightarrow{\text{proved}}
\mathrm{PED}
\xrightarrow{\text{open}}
\mathrm{CCSB}
\xrightarrow{\text{proved}}
\mathrm{MC}.
\end{gather*}
Here PDE is PrimeDensityEquipartition (PNT in arithmetic progressions,
a known theorem not yet in Mathlib), and GLPFE is GenericLPFEquidist
(Alladi's theorem~\cite{Alladi1977} on $\minFac$ distribution of generic integers,
also known but not formalized).  Both ends of the chain---from PDE to
GLPFE via Alladi, and from CCSB to MC via Fourier inversion---are
formally proved.

\paragraph{Window route.}
\[
\mathrm{StrongSieveEquidist}
\xrightarrow{\text{proved}}
\mathrm{NoLongRunsAt}
\xrightarrow{\text{proved}}
\mathrm{PEDAt}
\xrightarrow{\text{open}}
\mathrm{CCSB}
\xrightarrow{\text{proved}}
\mathrm{MC}.
\]
StrongSieveEquidist asserts that EM multipliers are equidistributed
within sliding windows; NoLongRunsAt and PEDAt are per-prime variants
proved by pigeonhole and block-counting respectively
(\lean{EM/Equidist/SieveTransfer.lean\#L185}{strongSieveEquidist\_noLongRunsAt},
\lean{EM/Equidist/SieveTransfer.lean\#L254}{noLongRunsAt\_ped}).

\paragraph{The genuine frontier.}

\textbf{SieveTransfer} is the critical open hypothesis: does the
equidistribution of $\minFac$ residues for generic integers transfer
to the specific EM orbit?  Everything above SieveTransfer is known
mathematics; everything below it is proved.  SieveTransfer is where
``known but not formalized'' meets ``genuinely open.''

The difficulty: for \emph{generic} $q$-rough integers,
$\minFac$ residues are equidistributed (by CRT + Mertens)---the
death channel gets its fair share of multipliers.  But the EM
Euclid numbers are not generic: each is determined by the entire
walk history.  Transferring the generic equidistribution to this
specific orbit is the open step.

\paragraph{Sieve-to-harmonic convergence.}

The sieve hierarchy (\S36--\S39 of \code{Equidist/SelfCorrecting.lean})
and the harmonic hierarchy (\S30--\S35) converge: both produce
DecorrelationHypothesis as output.  The full chain
\[
\mathrm{SieveEquidist}
\xrightarrow{\text{proved}}
\mathrm{Dec}
\xrightarrow{\text{proved}}
\mathrm{PED}
\xrightarrow[\text{sole gap}]{\text{open}}
\mathrm{CCSB}
\xrightarrow{\text{proved}}
\mathrm{MC}
\]
is formalized, with the first two arrows machine-verified
(\lean{EM/LargeSieve/Basic.lean\#L1617}{sieve\_equidist\_implies\_decorrelation},
\lean{EM/Equidist/SelfCorrecting.lean\#L139}{decorrelation\_implies\_ped}).
The sieve route achieves SieveEquidist~$\Rightarrow$~Dec via a
counting-to-character-sum bridge: SieveEquidistribution produces
\code{EMMultCharSumBound} with $Q_0 = 0$, meaning multiplier character
sums cancel for \emph{all} primes~$q$, which is exactly
DecorrelationHypothesis.  The sole remaining gap on this route is
\textbf{PEDImpliesComplexCSB}
(\lean{EM/Equidist/SelfCorrecting.lean\#L109}{PEDImpliesComplexCSB}):
does positive escape density for all primes imply walk character sum
cancellation?  Any proof of SieveEquidistribution (e.g., from PNT
in APs $+$ Alladi's theorem) would immediately yield Dec and PED
for free, isolating this single bridge as the only open step.

\paragraph{DSL: The Sole Remaining Hypothesis}

The ensemble averaging framework (Section~\ref{sec:ensemble}) collapses
the entire reduction landscape to a single open statement:
the \textbf{Deterministic Stability Lemma}~(DSL,
Property~\ref{def:dsl}).  DSL asserts that population
equidistribution of multiplier residues (a consequence of standard
ANT) implies conditional multiplier equidistribution for the
standard EM orbit ($n = 2$).  The assembly
$\mathrm{DSL} \Rightarrow \mathrm{CME} \Rightarrow \mathrm{CCSB}
\Rightarrow \mathrm{MC}$
is fully machine-verified
(\lean{EM/Reduction/Master.lean}{full\_chain\_dsl}).
DSL is the orbit-transfer principle: it bridges from population-level
statistics (which hold for almost all starting points) to the specific
orbit.

A weaker variant also suffices:

\begin{theorem}[\lean{EM/Ensemble/Structure.lean}{pe\_dsl\_hitting\_implies\_mc}]
$\mathrm{PE} + \mathrm{DSLHitting} \;\Longrightarrow\; \MC$,
\end{theorem}

\noindent where DSLHitting only requires that the walk hits~$-1$ for a
positive density of starting points, rather than full equidistribution.

\paragraph{The Adelic Decomposition of CME}
\label{sec:adelic}

The EM walk does not privilege any single prime: at each step, the race
across \emph{all} coordinates of $\hat{\ZZ}^\times = \prod_p (\ZZ/p\ZZ)^\times$
is run and the winner ($\minFac$) is selected.  The process is
intrinsically adelic.  CRT multiplier invariance
(\lean{EM/Group/CRT.lean}{crt\_multiplier\_invariance}) shows that
$\minFac(\Prod{2}(n){+}1) \bmod q$ depends only on the residues of
$\Prod{2}(n)$ at primes $r \neq q$: the multiplier at~$q$ is determined
by the \emph{non-$q$ coordinates} of the walk.

This structural observation decomposes CME into two independent conditions:

\begin{itemize}[nosep]
\item \textbf{Multiplier Marginal Equidistribution}~(MME): the
  unconditional multiplier character sums
  $\sum_{n<N} \chi(\multZ{2}{q}{n})$ vanish for each non-trivial~$\chi$.
  This captures the \emph{analytic} content---the multiplier has no
  preferred direction.
\item \textbf{Multiplier--Walk Independence}~(MWI): the multiplier
  character sum conditioned on $\walkZ{2}{q}{n} = a$ is bounded by
  the unconditional sum plus $o(N)$.  This captures the
  \emph{structural} content---knowing where the walk is does not predict
  the multiplier.
\end{itemize}

\begin{theorem}[\lean{EM/Adelic/Equidist.lean}{cme\_iff\_adelic}]
\label{thm:cme-iff-adelic}
$\mathrm{CME} \;\Longleftrightarrow\; \mathrm{MWI} + \mathrm{MME}$.
\end{theorem}

\begin{proof}
$(\Rightarrow)$: CME gives MWI trivially (the conditional sum
is already $o(N)$); CME gives MME by decomposing the unconditional sum
into fibers
$\sum_n \chi(m(n)) = \sum_{a} \sum_{\{n : w(n) = a\}} \chi(m(n))$
via \code{Finset.sum\_fiberwise}, applying CME at each fiber with
$\varepsilon' = \varepsilon / |\GG|$, and summing
(\lean{EM/Adelic/Equidist.lean}{cme\_implies\_mme}).

$(\Leftarrow)$: Given $\chi \neq 1$ and walk position~$a$, MWI bounds
the conditional sum by $\|\sum \chi(m)\| + (\varepsilon/2)\cdot N$,
and MME bounds the unconditional sum by $(\varepsilon/2)\cdot N$.
Together: $\leq \varepsilon N$.
\end{proof}

The equivalence shows that MME (analytic, from population equidistribution)
and MWI (structural, from CRT + cross-prime decorrelation) are genuinely
independent components.  The proved identity
$\mathrm{MME} \Leftrightarrow \text{vanishing walk lag-1 autocorrelation}$
(\lean{EM/Adelic/Equidist.lean}{mme\_iff\_walk\_autocorrelation},
via \code{walk\_shift\_one\_correlation}) clarifies that CCSB (walk equidistribution)
and MME (multiplier equidistribution, equivalently walk autocorrelation)
are logically independent: neither implies the other.

\medskip
A key bridge is the \textbf{CRT Multiplier Fiber} hypothesis: for
non-trivial $\chi, \psi : (\ZZ/q\ZZ)^\times \to \CC^\times$, the mixed sum
$\sum_{n<N} \chi(\multZ{2}{q}{n}) \cdot \psi(\walkZ{2}{q}{n}) = o(N)$.
This says that the multiplier and walk at prime~$q$ are decorrelated
at the character level.

\begin{theorem}[\lean{EM/Adelic/Equidist.lean}{crt\_fiber\_implies\_mwi\_proved}]
\label{thm:crt-fiber-mwi}
$\mathrm{CRTFiber} + \mathrm{MME} \;\Longrightarrow\; \mathrm{MWI}$.
Hence $\mathrm{CRTFiber} + \mathrm{MME} \;\Longrightarrow\; \mathrm{CME}
\;\Longrightarrow\; \mathrm{MC}$, with no open~Props.
\end{theorem}

\begin{proof}
By Fourier inversion on the finite abelian group $\GG = (\ZZ/q\ZZ)^\times$:
\[
  \mathbf{1}_{w(n)=a}
  \;=\; \frac{1}{|\GG|} \sum_{\psi \in \hat\GG}
    \bar\psi(a) \cdot \psi(w(n)).
\]
Substituting into the conditional sum
$\sum_{\{n : w(n)=a\}} \chi(m(n))$ and exchanging sums gives
\[
  \frac{1}{|\GG|}\,\sum_\psi \bar\psi(a) \cdot
  \sum_{n<N} \chi(m(n)) \cdot \psi(w(n)).
\]
The trivial character $\psi = 1$ contributes
$(1/|\GG|) \cdot \|\sum \chi(m)\|$, bounded by MME\@.
Each non-trivial~$\psi$ contributes a mixed character sum
bounded by CRTFiber.  Triangle inequality
(using $|\bar\psi(a)| = 1$) over the $|\GG|-1$ non-trivial characters
gives $\|\text{conditional}\| \leq \|\text{unconditional}\| + \varepsilon N$.

The formal proof (145~lines) uses \code{char\_indicator\_expansion} for
the Fourier inversion identity and \code{MulChar.equivToUnitHom} to
bridge between \code{DirichletCharacter\;\CC\;q}
(which is \code{Fintype} and has the orthogonality infrastructure)
and the \code{(ZMod\;q)\textsuperscript{x}\;→*\;\CC\textsuperscript{x}}
type used in the hypothesis definitions.
\end{proof}

\paragraph{The Profinite Weyl Criterion}

The walk lives in $\hat{\ZZ}^\times = \prod_p (\ZZ/p\ZZ)^\times$.
Uniform Profinite Equidistribution~(UPE) is the Weyl criterion for
this compact abelian group: for any finite set of distinct primes
$q_1, \ldots, q_k$ (not appearing in the sequence) and any product
of non-trivial characters $\chi_i : (\ZZ/q_i\ZZ)^\times \to \CC^\times$,
\[
  \Bigl\|\sum_{n<N} \prod_{i=1}^k
    \chi_i(\walkZ{2}{q_i}{n})\Bigr\| \;\leq\; \varepsilon N
  \qquad\text{for all } N \geq N_0(\varepsilon).
\]

\begin{theorem}[\lean{EM/Adelic/Profinite.lean}{uniform\_profinite\_implies\_ccsb},
  \lean{EM/Adelic/Profinite.lean}{uniform\_profinite\_implies\_cpd}]
\label{thm:upe-specializations}
$\mathrm{UPE}$ implies both\/ $\mathrm{CCSB}$
(specialization to $k{=}1$, via $\mathsf{Fin.prod\_univ\_one}$)
and\/ $\mathrm{CPD}$
(specialization to $k{=}2$, via $\mathsf{Fin.prod\_univ\_two}$).
Hence $\mathrm{UPE} \Rightarrow \mathrm{MC}$.
\end{theorem}

The natural question is whether the converse holds: does pairwise
decorrelation (CCSB at $k{=}1$ plus CPD at $k{=}2$) bootstrap to the
full $k$-wise condition~UPE\@?

\begin{remark}[Dead End~\#125]
\label{rem:pairwise-k-wise}
\textbf{No.}  The abstract XOR counterexample is decisive: three
unit-modulus sequences $X_1, X_2, X_3 := X_1 \cdot X_2$ satisfy all
pairwise cancellation conditions
($\sum X_i \bar X_j = o(N)$) but have
$\sum X_1 X_2 X_3 = \sum |X_1|^2 |X_2|^2 = N$.
No standard inequality (Cauchy--Schwarz, H\"older, van~der Corput)
bridges pairwise to $k$-wise.
The Tao--Ter\"av\"ainen theorem~\cite{TaoTeravainenPairwise}
(``pairwise implies higher'' for multiplicative functions) requires
multiplicativity, which the EM walk lacks.  UPE therefore requires
$k$-wise decorrelation for \emph{all}~$k$ as a primitive hypothesis,
not derivable from $k{=}1$ and $k{=}2$ data alone.
\end{remark}

The hierarchy is:
\[
  \mathrm{UPE}
  \;\Longrightarrow\;
  \bigl\{\mathrm{CCSB},\;\mathrm{CPD}\bigr\}
  \;\Longrightarrow\;
  \mathrm{MC},
  \qquad\text{but}\quad
  \mathrm{CCSB} + \mathrm{CPD}
  \;\not\Rightarrow\;
  \mathrm{UPE}.
\]
CPD combined with CRT multiplier invariance gives MWI
(conditional on \code{CPDImpliesCRTFiber}, the sole remaining bridge
in the adelic framework).  The full adelic landscape
(\lean{EM/Adelic/Equidist.lean}{adelic\_landscape}) witnesses five
proved routes to MC through the adelic framework:
CME~$\Leftrightarrow$~AdelicEquidist,
AdelicEquidist~$\Rightarrow$~MC,
MWI~$+$~MME~$\Rightarrow$~MC,
CRTFiber~$+$~MME~$\Rightarrow$~CME (fully proved),
and CPD~$+$~MME~$\Rightarrow$~MC (conditional on one bridge).

\paragraph{What Would Close the Conjecture}

The cleanest paths to MC:
\begin{enumerate}
\item \textbf{Prove DSL} (sole remaining hypothesis): show that the
  character energy of generalized EM sequences concentrates when averaged
  over squarefree starting points.  This is the weakest known hypothesis
  that implies MC, and the only one that does so without additional
  conditions.  The assembly
  DSL~$\Rightarrow$~CME~$\Rightarrow$~CCSB~$\Rightarrow$~MC is fully
  machine-verified (\lean{EM/Reduction/Master.lean}{full\_chain\_dsl}).

\item \textbf{Prove CME}: show that the multiplier
  character sum $\sum_{\substack{n < N \\ w(n) = c}} \chi(m(n))$ is
  $o(N)$ for each walk position~$c$.  CME is strictly weaker than
  CCSB, and $\mathrm{CME} \Rightarrow \mathrm{CCSB}$ is proved
  (\lean{EM/LargeSieve/Spectral.lean}{cme\_implies\_ccsb}).
  CME asks only about the \emph{conditional} distribution of
  multipliers given walk state---it does not require controlling the
  walk character sum itself.  This bypasses the $d \geq 3$ barrier
  entirely.

\item \textbf{Prove CCSB directly}: show that no non-trivial
  character sum can sustain the $\Omega(N)$ bias that permanent
  avoidance of~$-1$ would require.  The self-correcting sieve
  (concentration of EM primes in a residue class is exponentially
  self-limiting) is the strongest heuristic argument.

\item \textbf{Prove a BV-type estimate}: even an averaged bound
  over~$q$ would suffice---for $q \geq Q_0$ it rules out permanent
  avoidance, and the finitely many $q < Q_0$ are handled by the
  inductive bootstrap (\S\ref{sec:bootstrap}).

\item \textbf{Close the orbit chain gap}: show that at each cofinally
  visited walk position, at least two distinct multiplier classes
  appear.  This would force the orbit chain to expand to the full group.

\item \textbf{Prove DH directly}: show that a multiplicative walk
  on a cyclic group with a generating set of multipliers must hit
  every element cofinally.  This is a combinatorial question about
  deterministic walks.

\item \textbf{Prove SieveTransfer}: show that the EM orbit's $\minFac$
  distribution matches that of generic integers, at least on average.
  The cumulative sieve route (\S38) reduces this to known number theory
  (PNT in APs + Alladi's theorem); closing SieveTransfer gives CME
  (and hence CCSB) via the conditional multiplier equidistribution
  framework.

\item \textbf{Prove BVImpliesMMCSB or PrimeArithLSImpliesMMCSB}: the
  large sieve route (\S41--\S65) reduces MC to a transfer hypothesis.
  Given Bombieri--Vinogradov or the analytic large sieve, the remaining
  open step is showing that the EM orbit's multipliers receive their
  fair share of each residue class---in particular, that the death
  channel is not systematically avoided.  This is the same
  orbit-specificity gap as SieveTransfer, approached from a different
  mathematical toolkit.
\end{enumerate}

\paragraph{Does the Walk Hit $-1$?}

The irreducible open question is not whether DH holds (cofinal
hitting is a convenient sufficient condition) but whether the walk
hits~$-1$ \emph{at least once} past the sieve gap.  The Single Hit
Theorem (Theorem~\ref{thm:single-hit}) shows that a single hit at
each prime suffices for MC; DH, CCSB, and CME are strategies for
producing that hit.

\paragraph{Evidence for.}
Three independent lines of evidence suggest the walk does hit~$-1$
(see also Remark~\ref{rem:why-dsl}).
(1)~\emph{The information bottleneck}: $\Prod{2}(n)$ has
${\sim}2^n$ bits, but the walk position carries only $O(\log q)$ bits.
The smallest prime factor of $\Prod{2}(n)+1$ depends on the full
value, not on its residue mod~$q$, so conditioning on walk position
constrains almost nothing about the factoring outcome---the
conditional and unconditional $\minFac$ distributions should agree.
(2)~\emph{Ensemble evidence}: for almost all squarefree starting
points, CME holds (the proved SD~$\Rightarrow$~cancellation chain).
The standard orbit $n = 2$ has no special arithmetic property that
would make it an outlier.  DSL is a typicality assertion, not a
structural miracle.
(3)~\emph{Analogy with Artin's conjecture}: Artin's conjecture
(that every non-square integer is a primitive root for infinitely
many primes) has the same orbit-specificity structure and is believed
true; Hooley~\cite{Hooley1967} proved it conditional on GRH.  The
walk-hitting question is the analogous statement for the EM walk
and would follow from an analogous uniformity hypothesis.

\paragraph{Evidence against.}
Two features of the EM sequence give pause.
(1)~\emph{Cox--van der Poorten}~\cite{CoxVdP1968}: the ``largest factor'' variant of
the Euclid--Mullin sequence provably misses primes.  The EM
sequence's completeness is not a soft consequence of the Euclid
construction but depends sensitively on the $\minFac$ selection
rule.  This fragility means heuristic arguments
(``it should work because Euclid numbers have many factors'') are
not reliable.
(2)~\emph{The $d \geq 3$ barrier}: the formalization proves that
the most natural route from multiplier escape density to walk
character sum cancellation (PED~$\Rightarrow$~BRE~$\Rightarrow$~CCSB) is
\emph{impossible} for character orders $d \geq 3$.  This is not
evidence against the hit, but it shows that producing one
requires mechanisms beyond the simplest equidistribution
framework.  The CME bypass sidesteps this barrier, but the barrier's
existence means any proof must be genuinely subtle.

\paragraph{Collapse certificates.}
Two natural-seeming proof strategies---modelling the EM walk as a
non-homogeneous Markov chain (Doeblin--Dobrushin ergodic theory) and
adapting the multiplicative large sieve to the $\minFac$ selection
rule---were assessed by formalization.  Both \emph{collapse
definitionally} to existing open hypotheses:
\begin{itemize}
\item
  \lean{EM/Advanced/MarkovSieve.lean}{doeblin\_eq\_cme}:
  \texttt{DoeblinConvergenceForEM = CME} by \texttt{rfl}.
  The Doeblin coefficient of ergodicity approaching~$1$ for the
  empirical transition matrix of the EM walk is literally the
  statement that conditional multiplier character sums vanish, which
  is CME.
\item
  \lean{EM/Advanced/MarkovSieve.lean}{sieve\_orbit\_control\_eq\_ccsb}:
  \texttt{SieveOrbitControl = CCSB} by \texttt{rfl}.
  Pointwise-in-$q$ walk character sum bounds from sieve methods are
  literally CCSB\@.
\end{itemize}
A definitional equality (\texttt{rfl}) is the strongest possible dead-end
certificate: the type-checker confirms that the ``new'' hypothesis
contains zero mathematical content beyond the existing one.
The underlying obstruction is the \emph{orbit-specificity gap}
(average-over-$q$ versus pointwise-in-$q$), which is structurally
analogous to the gap between Chebotarev's density theorem
(unconditional) and Artin's primitive root conjecture
(conditional on GRH; Hooley 1967).

The root cause is algebraic.  The identity
\begin{equation}
  \label{eq:minfac-min}
  \minFac(nm) = \min\bigl(\minFac(n),\;\minFac(m)\bigr)
  \qquad (n,m > 1),
  \tag{$\checkmark$ \code{minFac\_mul\_eq\_min}}
\end{equation}
proved without any coprimality hypothesis, shows that $\minFac$ is a
\emph{lattice} operation (min), not an algebraic one (product).
Multiplicative number theory---Dirichlet series, the Hal\'asz theorem,
pretentious distance---is designed for functions $f$ satisfying
$f(nm) = f(n)f(m)$, and is structurally incompatible with a function
obeying~\eqref{eq:minfac-min}.

\paragraph{Assessment.}
We believe the walk does hit~$-1$ at every prime, primarily because
the self-correcting sieve mechanism provides a concrete dynamical
reason (not merely a probabilistic heuristic) for the walk to
eventually hit~$-1$.  The strongest form of this belief: CME should
hold because the EM multipliers, conditioned on walk position, have
no arithmetic reason to avoid the death channel systematically.
But we acknowledge that no existing technique can prove this.
The formalization has exhaustively explored all known proof strategies
across 134~dead ends: the $d \geq 3$ barrier, the MultCancelToWalkCancel
counterexample (Dead End~\#117), the cofactor identity's failure to
escape the Marginal/Joint Barrier (Dead End~\#115), and the exhaustion
of all four DSL proof directions (algebraic identities, cofactor
completeness, exponential sums, sum-product theory) collectively show
that the proof, when found, will require genuinely new mathematical
techniques---likely exploiting the specific arithmetic of the
$\minFac$ rule on super-exponentially growing squarefree numbers in
ways that current analytic number theory does not.

\renewcommand{\arraystretch}{1.25}
\footnotesize
\begin{longtable}{@{}p{5.5cm}l>{\raggedright\arraybackslash}p{4.5cm}@{}}
\toprule
\textbf{Result} & \textbf{Status} & \textbf{Lean identifier} \\
\midrule
\endfirsthead
\toprule
\textbf{Result} & \textbf{Status} & \textbf{Lean identifier} \\
\midrule
\endhead
\midrule
\multicolumn{3}{r}{\emph{continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot
\multicolumn{3}{l}{\emph{Sequence foundations}} \\
\quad Every $\seq{2}(n)$ is prime & Proved & \code{seq\_isPrime} \\
\quad No prime repeats & Proved & \code{seq\_injective} \\
\midrule
\multicolumn{3}{l}{\emph{Main reductions to MC}} \\
\quad \textbf{DSL $\Rightarrow$ MC} (sole hypothesis, via ensemble) & \textbf{Proved} & \code{full\_chain\_dsl} \\
\quad \textbf{SHH $\Rightarrow$ MC} (Single Hit Theorem) & Proved & \code{single\_hit\_implies\_mc} \\
\quad \textbf{DH $\Rightarrow$ MC} (cofinal hitting route) & Proved & \code{dynamical\_hitting\_implies\_mullin} \\
\quad \textbf{CCSB $\Rightarrow$ MC} (single hypothesis) & Proved & \code{complex\_csb\_mc'} \\
\quad \textbf{CME $\Rightarrow$ MC} (sharpest, all $d$) & Proved & \code{cme\_implies\_mc} \\
\quad PE + DSL $\Rightarrow$ MC & Proved & \code{pe\_dsl\_implies\_mc} \\
\quad PE + DSLHitting $\Rightarrow$ MC & Proved & \code{pe\_dsl\_hitting\_implies\_mc} \\
\quad PE $\Rightarrow$ MC & Proved & \code{pe\_implies\_mullin} \\
\quad HH $\Rightarrow$ MC & Proved & \code{hh\_implies\_mullin} \\
\quad SE + MH $\Rightarrow$ MC & Proved & \code{se\_mixing\_implies\_mullin} \\
\quad \textbf{WE $\Rightarrow$ MC} (single Prop) & Proved & \code{walk\_equidist\_mc} \\
\midrule
\multicolumn{3}{l}{\emph{Inductive bootstrap}} \\
\quad PrimeResidueEscape (elementary) & Proved & \code{prime\_residue\_escape} \\
\quad MC($<\!p$) + PRE $\Rightarrow$ SE($p$) & Proved & \code{mcBelow\_pre\_implies\_se} \\
\quad $q$-roughness from MC($<\!q$) & Proved & \code{mcBelow\_implies\_seq\_ge} \\
\quad One-prime gap & Proved & \code{mcBelow\_cofinal\_hit\_implies\_mc\_at} \\
\quad mcBelow 11 & Proved & \code{concrete\_mcBelow\_11} \\
\midrule
\multicolumn{3}{l}{\emph{Algebraic framework}} \\
\quad Confinement Theorem & Proved & \code{confinement\_forward/reverse} \\
\quad PRE $\Leftrightarrow$ SE & Proved & \code{pre\_iff\_se} \\
\quad SE $\Leftrightarrow$ character detection & Proved & \code{se\_iff\_char\_detection} \\
\quad Maximal subgroup reduction & Proved & \code{se\_of\_maximal\_escape} \\
\quad WHP $\Leftrightarrow$ HH & Proved & \code{whp\_iff\_hh} \\
\quad SE density argument (CRT + QR) & Proved & \code{se\_qr\_obstruction} \\
\midrule
\multicolumn{3}{l}{\emph{Character sum chain}} \\
\quad Fourier bridge: CCSB $\Rightarrow$ hit count lb & Proved & \code{complex\_csb\_implies\_hit\_count\_lb\_proved} \\
\quad Decorrelation $\Rightarrow$ PED & Proved & \code{decorrelation\_implies\_ped} \\
\quad NoLongRuns$(L)$ $\Rightarrow$ PED & Proved & \code{noLongRuns\_implies\_ped} \\
\quad BRE $\Rightarrow$ PEDImpliesCSB & Proved & \code{block\_rotation\_implies\_ped\_csb} \\
\quad CME $\Rightarrow$ CCSB (all $d$, bypasses BRE) & Proved & \code{cme\_implies\_ccsb} \\
\quad CME $\Rightarrow$ MC & Proved & \code{cme\_implies\_mc} \\
\quad Walk char recurrence ($\mathbb{C}$-valued) & Proved & \code{char\_walk\_recurrence} \\
\quad Telescoping identity & Proved & \code{walk\_telescope\_identity} \\
\quad Telescoping norm $\leq 2$ & Proved & \code{walk\_telescope\_norm\_bound} \\
\quad Shift-one autocorrelation & Proved & \code{walk\_shift\_one\_correlation} \\
\quad Order-2 sign-flip chain & Proved & \code{order2\_noLongRuns\_mc} \\
\midrule
\multicolumn{3}{l}{\emph{Walk dynamics}} \\
\quad Walk--divisibility bridge & Proved & \code{walkZ\_eq\_neg\_one\_iff} \\
\quad Products strictly monotone & Proved & \code{prod\_strictMono} \\
\quad Fundamental trichotomy & Proved & \code{avoidance\_contradicts\_se\_mixing} \\
\quad Self-avoidance dichotomy & Proved & \code{self\_avoidance\_dichotomy} \\
\quad Scheduled walk coverage & Proved & \code{scheduled\_walk\_covers\_all} \\
\midrule
\multicolumn{3}{l}{\emph{Selectability analysis}} \\
\quad Divisor freshness & Proved & \code{divisor\_not\_yet\_in\_seq} \\
\quad Passed-over persistence & Proved & \code{passed\_over\_persists} \\
\quad Selectability extinction & Proved & \code{selectability\_extinguished} \\
\quad MC $\Rightarrow$ $\neg$InfinitelySelectable & Proved & \code{mc\_implies\_not\_infinitely\_selectable} \\
\quad DH $\Rightarrow$ InfinitelySelectable & Proved & \code{dh\_implies\_infinitely\_selectable} \\
\midrule
\multicolumn{3}{l}{\emph{Sieve and orbit analysis}} \\
\quad EMFE $\Leftrightarrow$ TailSE & Proved & \code{emfe\_iff\_tail\_se\_at} \\
\quad TailSE $\Rightarrow$ CofinalEscape $\Rightarrow$ QuotientDH & Proved & \code{tail\_se\_gives\_sub\_dh} \\
\quad Dirichlet: $\infty$ primes per residue class & Proved & \code{dirichlet\_residues\_independent} \\
\quad Minimality sieve + coupling & Proved & \code{minimality\_sieve} \\
\quad StrongSieveEquidist $\Rightarrow$ NoLongRunsAt & Proved & \code{strongSieveEquidist\_noLongRunsAt} \\
\quad NoLongRunsAt $\Rightarrow$ PEDAt & Proved & \code{noLongRunsAt\_ped} \\
\quad DPED $\Rightarrow$ PED & Proved & \code{dped\_implies\_ped} \\
\quad PDE $+$ sieve chain $\Rightarrow$ MC & Proved & \code{primeDensity\_chain\_mc} \\
\quad GLPFE $+$ SieveTransfer $\Rightarrow$ MC & Proved & \code{genericLPF\_chain\_mc} \\
\quad SieveEquidist $\Rightarrow$ Dec & Proved & \code{sieve\_equidist\_implies\_decorrelation} \\
\quad SieveEquidist $\Rightarrow$ PED & Proved & \code{sieve\_equidist\_implies\_ped} \\
\midrule
\multicolumn{3}{l}{\emph{Ensemble and DSL framework}} \\
\quad Generalized EM: genProd, genSeq & Proved & \code{genProd\_two\_eq\_prod} \\
\quad genSeq injective & Proved & \code{genSeq\_injective} \\
\quad $n \mid \text{genProd}(n, k)$ & Proved & \code{start\_dvd\_genProd} \\
\quad $\Prod{2}(n)$ squarefree & Proved & \code{prod\_squarefree} \\
\quad Buchstab weight $\kappa(2) = 2/3$ & Proved & \code{buchstabWeight\_two} \\
\quad FM + VB $\Rightarrow$ RSD & Proved & \code{first\_moment\_variance\_implies\_rsd} \\
\quad CharConcentration $\Rightarrow$ Cancellation & Proved & \code{char\_concentration\_implies\_cancellation} \\
\quad FEB $\Rightarrow$ CME (fiber energy) & Proved & \code{feb\_implies\_cme} \\
\quad Cross term $=$ sum of fiber cross terms & Proved & \code{total\_cross\_term\_eq\_sum\_fiber} \\
\quad $\|\chi(u)\| = 1$ for group chars & Proved & \code{char\_value\_norm\_one} \\
\quad Energy $= N + 2 \sum$ cross terms & Proved & \code{char\_sum\_energy\_eq\_N\_plus\_cross} \\
\quad $o(N)$ cross terms $\Rightarrow$ linear energy & Proved & \code{cross\_term\_controls\_growth} \\
\quad ExcessEnergy $=$ visit deviation & Proved & \code{excessEnergy\_eq\_visit\_deviation} \\
\quad SVE $\Leftrightarrow$ VE & Proved & \code{sve\_iff\_ve} \\
\quad Population energy $\geq 0$ & Proved & \code{populationCharEnergy\_nonneg} \\
\quad Population energy $=$ ensemble avg & Proved & \code{populationCharEnergy\_eq\_ensembleAvg} \\
\quad Cross-term pair symmetry & Proved & \code{crossTermPair\_comm} \\
\quad CSVB $\Rightarrow$ SecondMomentBound & Proved & \code{charSumVarianceBound\_implies\_secondMomentBound} \\
\quad SD $\Rightarrow$ PairwiseCrossTermVanishing & Proved & \code{stepDecorrelation\_implies\_pairwiseVanishing} \\
\quad Variance chain from bridges & Proved & \code{variance\_chain\_from\_bridges} \\
\quad EnsembleConcentration $\Rightarrow$ AlmostAllSqfreeEqd & Proved & \code{ensemble\_concentration\_implies\_eqd} \\
\quad SE $\Rightarrow$ NAO & Proved & \code{se\_implies\_nao} \\
\quad Full CRT chain $\Rightarrow$ cancellation & Proved & \code{full\_crt\_chain\_implies\_cancellation} \\
\quad WeylHittingBridge & Proved & \code{weyl\_hitting\_bridge\_proved} \\
\quad PerChiCancellationBridge & Proved & \code{per\_chi\_cancellation\_bridge\_proved} \\
\quad Walk cancellation ($n{=}2$) $\Rightarrow$ MC & Proved & \code{cancel\_weyl\_implies\_mc} \\
\quad JSE $+$ MultCancelToWalkCancel $\Rightarrow$ MC & Proved & \code{jse\_transfer\_implies\_mc} \\
\quad SD $\Rightarrow$ a.a.\ char sum cancellation & Proved & \code{sd\_implies\_cancellation} \\
\quad 4-hypothesis master (SRE+CRT+Bridge+Decorr) & Proved & \code{ensemble\_pt\_master\_simplified} \\
\quad GenHittingImpliesGenMC & Proved & \code{gen\_hitting\_implies\_gen\_mc\_proved} \\
\quad DecorrelationImpliesVariance & Proved & \code{decorrelation\_implies\_variance\_proved} \\
\quad CharVarianceImpliesConcentration & Proved & \code{char\_variance\_implies\_concentration\_proved} \\
\quad EnsembleMultEquidistImpliesCharMeanZero & Proved & \code{ensemble\_mult\_equidist\_char\_mean\_zero} \\
\quad JSE base case (tower rule) & Proved & \code{jse\_base\_case} \\
\quad Cofactor identity: $w(n){+}1 = m(n) \cdot \mathrm{cof}(n)$ & Proved & \code{shifted\_walk\_eq\_mult\_mul\_cof} \\
\quad Hit $\Leftrightarrow$ cofZ $= 0$ & Proved & \code{walkZ\_eq\_neg\_one\_iff\_cofZ\_zero} \\
\midrule
\multicolumn{3}{l}{\emph{UCE (sieve-based) route}} \\
\quad UCE + UCEImpliesCME $\Rightarrow$ MC & Proved & \code{uce\_cme\_implies\_mc} \\
\quad UCE + Alladi + DSL $\Rightarrow$ MC & Proved & \code{uce\_dsl\_implies\_mc} \\
\quad $\Prod{2}(n)$ coprime to primes not in seq & Proved & \code{prod\_coprime\_of\_not\_in\_seq} \\
\quad UCE landscape (summary conjunction) & Proved & \code{uce\_landscape} \\
\midrule
\multicolumn{3}{l}{\emph{FF-MC large-$p$ chain}} \\
\quad FFDH $\Rightarrow$ FF-MC (competitor exhaustion) & Proved & \code{ff\_dh\_implies\_ffmc} \\
\quad FF-EM sequence injective & Proved & \code{ffSeq\_injective\_proved} \\
\quad Weil $\Rightarrow$ FF-SE for $p \geq (d{-}1)^2{+}1$ & Proved & \code{weil\_implies\_ff\_se} \\
\quad FF-SE unconditional for $d = 1$ & Proved & \code{ff\_se\_degree\_one} \\
\quad $\ZZ/4\ZZ$ counterexample (generation $\neq$ coverage) & Proved & \code{alternating\_walk\_misses\_two} \\
\quad Weil $\Rightarrow$ population multiplier CCSB & Proved & \code{weil\_implies\_population\_mult\_ccsb} \\
\quad Selection bias chain & Proved & \code{selection\_bias\_chain} \\
\midrule
\multicolumn{3}{l}{\emph{Large sieve route}} \\
\quad MultiModularCSB $\Rightarrow$ MC & Proved & \code{mmcsb\_implies\_mc} \\
\quad BV chain $\Rightarrow$ MC & Proved & \code{bv\_chain\_mc} \\
\quad ArithLS chain $\Rightarrow$ MC & Proved & \code{arith\_ls\_chain\_mc} \\
\quad ALS chain $\Rightarrow$ MC & Proved & \code{als\_prime\_arith\_ls\_chain\_mc} \\
\quad \textbf{WeakALS} (\S58) & \textbf{Proved} & \code{weak\_als\_from\_card\_bound} \\
\quad Gauss sum inversion (\S57) & Proved & \code{char\_sum\_to\_exp\_sum} \\
\quad \textbf{ALS $\Rightarrow$ PrimeArithLS} (\S65) & \textbf{Proved} & \code{als\_implies\_prime\_arith\_ls} \\
\quad Hilbert chain $\Rightarrow$ ALS & Proved & \code{hilbert\_chain\_als} \\
\quad HilbertInequality1 $\Rightarrow$ HilbertInequality (rescaling) & Proved & \code{hilbert\_rescale} \\
\quad MittagLefflerCsc (Mittag-Leffler for $\csc$) & Proved & \code{mittag\_leffler\_csc\_proved} \\
\quad CscBilinearBound $\Rightarrow$ GramOffDiag & Proved & \code{csc\_bilinear\_implies\_gram\_offdiag} \\
\quad CrossRCesaroConvergence & Proved & \code{cross\_r\_cesaro\_convergence\_proved} \\
\quad GramOffDiag $\Rightarrow$ ALS & Proved & \code{gram\_offdiag\_bilinear\_implies\_als} \\
\quad Refined Gram off-diagonal bound & Proved & \code{gramMatrix\_offdiag\_bound\_dist} \\
\quad Packing bound (pigeonhole) & Proved & \code{round\_sep\_card\_le} \\
\quad Jordan's inequality (\S56) & Proved & \code{sin\_pi\_ge\_two\_mul} \\
\quad Geometric sum bound (\S56) & Proved & \code{norm\_eAN\_geom\_sum\_le\_inv} \\
\quad Parseval for ZMod.dft (\S53) & Proved & \code{zmod\_dft\_parseval} \\
\quad Gauss sum norm $\|\tau\|^2 = p$ (\S54) & Proved & \code{gaussSum\_norm\_sq\_eq\_prime} \\
\quad Walk autocorrelation identities (\S53) & Proved & \code{walkAutocorrelation\_*} \\
\quad Character Parseval (\S60) & Proved & \code{char\_parseval\_units} \\
\quad All 8 GCT internal lemmas (\S56--\S62) & Proved & \code{gct\_nontrivial\_char\_sum\_le} \\
\midrule
\multicolumn{3}{l}{\emph{Open hypotheses --- live targets}} \\
\quad \textbf{DeterministicStabilityLemma} (sole hypothesis) & \textbf{Open} & \code{DeterministicStabilityLemma} \\
\quad \textbf{DSLHitting} (weaker variant) & \textbf{Open} & \code{DSLHitting} \\
\quad \textbf{SingleHitHypothesis} (weakest sufficient) & \textbf{Open} & \code{SingleHitHypothesis} \\
\quad \textbf{DynamicalHitting} & \textbf{Open} & \code{DynamicalHitting} \\
\quad \textbf{ComplexCharSumBound} & \textbf{Open} & \code{ComplexCharSumBound} \\
\quad \textbf{MultiModularCSB} & \textbf{Open} & \code{MultiModularCSB} \\
\quad DecorrelationHypothesis & \textbf{Open} & \code{DecorrelationHypothesis} \\
\quad PositiveEscapeDensity & \textbf{Open} & \code{PositiveEscapeDensity} \\
\quad \textbf{PEDImpliesComplexCSB} (sole sieve-route gap) & \textbf{Open} & \code{PEDImpliesComplexCSB} \\
\quad NoLongRuns$(L)$ & \textbf{Open} & \code{NoLongRuns} \\
\quad SieveEquidistribution & \textbf{Open} & \code{SieveEquidistribution} \\
\quad MertensEscape & \textbf{Open} & \code{MertensEscape} \\
\quad SieveAmplification & \textbf{Open} & \code{SieveAmplification} \\
\quad \textbf{SieveTransfer} (genuine frontier) & \textbf{Open} & \code{SieveTransfer} \\
\quad StrongSieveEquidist & \textbf{Open} & \code{StrongSieveEquidist} \\
\quad DistributionalPED & \textbf{Open} & \code{DistributionalPED} \\
\quad \textbf{BVImpliesMMCSB} (genuine frontier) & \textbf{Open} & \code{BVImpliesMMCSB} \\
\quad GaussConductorTransfer (all lemmas proved) & \textbf{Open} & \code{GaussConductorTransfer} \\
\quad PrimeArithLSImpliesMMCSB & \textbf{Open} & \code{PrimeArithLSImpliesMMCSB} \\
\quad HilbertInequality (IK Lemma~7.8, $\delta$-separated) & \textbf{Open} & \code{HilbertInequality} \\
\quad HilbertInequality1 (1-separated core) & \textbf{Open} & \code{HilbertInequality1} \\
\quad HilbertCscBilinearBridge & \textbf{Open} & \code{HilbertCscBilinearBridge} \\
\quad EnsembleConcentration & \textbf{Open} & \code{EnsembleConcentration} \\
\quad RecipSumConcentration & \textbf{Open} & \code{RecipSumConcentration} \\
\quad CharSumVarianceBound & \textbf{Open} & \code{CharSumVarianceBound} \\
\quad PopulationEquidist & \textbf{Open} & \code{PopulationEquidist} \\
\quad PopulationTransfer & \textbf{Open} & \code{PopulationTransfer} \\
\quad EMDImpliesCME & \textbf{Open} & \code{EMDImpliesCME} \\
\quad UniformConductorEquidist (sieve uniformity) & \textbf{Open} & \code{UniformConductorEquidist} \\
\quad UCEImpliesCME (orbit-specificity bridge) & \textbf{Open} & \code{UCEImpliesCME} \\
\quad JointStepEquidist (ensemble route gap) & \textbf{Open} & \code{JointStepEquidist} \\
\quad MultCancelToWalkCancel ($\equiv$ CCSB, Dead End \#117) & \textbf{Open} & \code{MultCancelToWalkCancel} \\
\quad PairwiseVanishingImpliesSMB (variance route) & \textbf{Open} & \code{PairwiseVanishingImpliesSMB} \\
\quad SMBImpliesDSL (second moment $\Rightarrow$ DSL) & \textbf{Open} & \code{SMBImpliesDSL} \\
\quad VarianceChainImpliesDSL (full variance chain) & \textbf{Open} & \code{VarianceChainImpliesDSL} \\
\quad FFDynamicalHitting (cofinal $-1$ hits for FF-EM) & \textbf{Open} & \code{FFDynamicalHitting} \\
\quad SelectionBiasNeutral (min-degree preserves cancel.) & \textbf{Open} & \code{SelectionBiasNeutral} \\
\quad WalkCharExpansion (Fourier inversion on finite groups) & \textbf{Open} & \code{WalkCharExpansion} \\
\midrule
\multicolumn{3}{l}{\emph{Known theorems --- not yet in Mathlib}} \\
\quad PrimeDensityEquipartition (PNT in APs) & Known & \code{PrimeDensityEquipartition} \\
\quad GenericLPFEquidist (Alladi~\cite{Alladi1977}) & Known & \code{GenericLPFEquidist} \\
\quad BombieriVinogradov & Known & \code{BombieriVinogradov} \\
\quad AnalyticLargeSieve & Known & \code{AnalyticLargeSieve} \\
\quad ArithmeticLargeSieve & Known & \code{ArithmeticLargeSieve} \\
\midrule
\multicolumn{3}{l}{\emph{Dead ends --- false or blocked}} \\
\quad MultCSBImpliesMMCSB (false in general, \S\ref{sec:walk-bridge-false}) & \textbf{Dead} & \code{MultCSBImpliesMMCSB} \\
\quad BlockRotationEstimate (impossible for $d \geq 3$, Remark~\ref{rem:bre-impossible}) & \textbf{Dead} & \code{BlockRotationEstimate} \\
\quad MultCancelToWalkCancel (counterexample, Dead End~\#117) & \textbf{Dead} & \code{MultCancelToWalkCancel} \\
\quad Cofactor identity (joint variable, Dead End~\#115) & \textbf{Dead} & shifted\_walk\_eq\_mult\_mul\_cof \\
\quad Function field analog (Weil $\not\Rightarrow$ FF-DSL, Dead End~\#127) & \textbf{Dead} & \code{WeilDoesNotImplyFFDSL} \\
\quad $p$-adic geometry (technique mismatch, Dead End~\#128) & \textbf{Dead} & \code{padicGeometryDeadEnd} \\
\quad FFLM structurally false (cyclotomic, Dead End~\#129) & \textbf{Dead} & \code{ff\_cyclotomic\_dead\_end} \\
\quad SE $\not\Rightarrow$ DH (generation $\neq$ coverage, Dead End~\#130) & \textbf{Dead} & \code{se\_does\_not\_imply\_dh} \\
\quad Dobrushin coefficient $= 1$ (deterministic, Dead End~\#131) & \textbf{Dead} & \code{dobrushin\_coefficient\_one} \\
\quad $L$-function factorization circular (Dead End~\#132) & \textbf{Dead} & \code{dead\_end\_132\_witness} \\
\quad Self-similar FE $\neq$ Lapidus framework (Dead End~\#133) & \textbf{Dead} & \code{dead\_end\_133\_witness} \\
\quad No Tauberian lever for $L_{\mathrm{EM}}$ (Dead End~\#134) & \textbf{Dead} & \code{tauberian\_lever\_absent\_witness} \\
\end{longtable}
\renewcommand{\arraystretch}{1.0}
\normalsize


