%% =========================================================================
\section{Ensemble Averaging and the DSL Route}
\label{sec:ensemble}
%% =========================================================================

The orbit-specificity barrier---the difficulty of transferring
population statistics to one deterministic orbit---can be circumvented
by averaging \emph{over the orbit parameter itself}.  This section
develops the ensemble framework: we parametrize the EM construction by
its starting point, average over squarefree starting points via the
Chinese Remainder Theorem, and reduce MC to a single
\emph{Deterministic Stability Lemma}~(DSL)\@.

\medskip\noindent\textbf{What is new vs.\ what is infrastructure.}\enspace
The generalized EM sequences, the CRT propagation framework, and the
DSL~$\Rightarrow$~MC reduction (Theorem~\ref{thm:dsl-mc}) are original
contributions of this formalization.  The decorrelation--variance--concentration
chain (\S\ref{sec:decorr-chain}) is proved with explicit constants
($C = 2$, Theorem~\ref{thm:variance-bound}).  The Weyl hitting bridge
(\S\ref{sec:weyl-bridge}) and the generalized MC argument
(\S\ref{sec:gen-mc}) are fully original.  The sieve density function
and Buchstab weight (\S\ref{sec:first-moment}) formalize known sieve
theory.  The visit equidistribution identity (\S\ref{sec:visit-eqd})
connecting spectral and geometric viewpoints is new.

Sections~\ref{sec:walk}--\ref{sec:hard} worked with the standard orbit
$s = 2$.  We now exploit the starting-point parameter~$s$ introduced in
\eqref{eq:gen-em}: the \emph{ensemble} is the family of EM trajectories
$\{\Prod{n} : n \text{ squarefree}\}$.  The structural properties
established in Section~\ref{sec:walk}---divisibility, monotonicity,
injectivity, squarefreeness, and the death channel
equivalence---hold for every starting point and are used throughout.

\subsection{CRT Equidistribution and Propagation}
\label{sec:crt-equidist}

The ensemble is the collection of generalized EM sequences indexed by
squarefree starting points $n \in [1, X]$.  For each prime~$q$, we
study how the residues $\Prod{n}(k) \bmod q$ and
$\seq{n}(k) \bmod q$ distribute as $n$ ranges over
squarefree integers.

\begin{definition}[\defname{Accumulator density}]
For prime~$q$, step~$k$, and residue class $a \in (\ZZ/q\ZZ)^\times$,
define the \emph{accumulator density}:
\[
  \rho_k(a) \;=\; \lim_{X \to \infty} \frac{|\{n \in [1,X] : n
    \text{ squarefree},\; \Prod{n}(k) \equiv a \pmod{q}\}|}
    {|\{n \in [1, X] : n \text{ squarefree}\}|}.
\]
\end{definition}

\begin{remark}[\defname{SquarefreeResidueEquidist} (\abbr{SRE})]
Squarefree integers equidistribute among non-zero residue classes
modulo any prime~$q$: $\rho_0(a) = 1/(q{-}1)$ for all
$a \in (\ZZ/q\ZZ)^\times$.
This is a standard consequence of the inclusion-exclusion formula for
squarefree counting (a known result, not yet in Mathlib).
\end{remark}

SRE provides the base case of a CRT induction:

\begin{property}[\defname{CRTPropagationStep}]
For all primes~$q$ and steps~$k$: if $\rho_k$ is uniform on
$(\ZZ/q\ZZ)^\times$, then so is $\rho_{k+1}$.
\end{property}

The structural basis for CRT propagation is the \emph{CRT multiplier
invariance}
(\lean{EM/MullinCRT.lean}{crt\_multiplier\_invariance}): the generalized
multiplier $\seq{n}(k) = \minFac(\Prod{n}(k) + 1)$
depends on the full integer $\Prod{n}(k) + 1$, not just its residue
modulo~$q$.  By CRT, knowing $\Prod{n}(k) \bmod q$ does not
constrain $\Prod{n}(k) \bmod r$ for any other prime~$r$, so the
factoring outcome is ``$q$-blind.''  This decorrelates the mod-$q$
coordinate of the accumulator at step~$k$ from the multiplier at
step~$k$.

The propagation proceeds by induction on~$k$:

\begin{theorem}[\lean{EM/EnsembleCRT.lean}{sre\_crt\_implies\_accum\_equidist}]
\label{thm:sre-crt}
$\mathrm{SRE} + \mathrm{CRTPropagationStep}$ implies
accumulator equidistribution:
$\rho_k(a) = 1/(q{-}1)$ for all~$k$, $q$,
$a \in (\ZZ/q\ZZ)^\times$.
\end{theorem}

From accumulator equidistribution, a population transfer bridge gives
\emph{ensemble multiplier equidistribution}~(EME): the density of
squarefree $n$ with $\seq{n}(k) \equiv a \pmod{q}$ tends
to $1/(q{-}1)$ for each non-zero~$a$.

\paragraph{CRT independence and joint equidistribution.}
A deeper structural property concerns the \emph{joint} distribution of
residues at different steps.  The \emph{marginal product} version---the
product $\rho_j(a) \cdot \rho_k(b)$---is trivially $1/(q{-}1)^2$ from
accumulator equidistribution.  The \emph{genuine joint}
version---counting squarefree~$n$ with $\Prod{n}(j) \equiv a$ and
$\Prod{n}(k) \equiv b$ \emph{simultaneously}---is not trivial: it
requires genuine independence of the mod-$q$ coordinates at different
steps, formalized as \defname{SquarefreeCRTIndependence}
(\code{EnsembleCRTFreedom.lean}).

The CRT freedom framework proves:

\begin{theorem}[\lean{EM/EnsembleCRTFreedom.lean}{scrti\_sre\_implies\_aep}]
$\mathrm{SCRTI} + \mathrm{SRE}$ implies
$\mathrm{AccumulatorEquidistPropagation}$.
\end{theorem}

\noindent This provides a clean decomposition: SRE is the base case
(a consequence of standard analytic number theory),
SquarefreeCRTIndependence is the inductive step (CRT structure), and
AccumulatorEquidistPropagation is the output.

\paragraph{Standard analytic number theory (ANT).}
\label{sec:ant-def}
Throughout this paper, \textbf{ANT} denotes the conjunction of two
classical results: (i)~Dirichlet's theorem on primes in arithmetic
progressions, and (ii)~sieve estimates for shifted squarefree
integers.  Together they imply
\defname{MinFacResidueEquidist}~(\abbr{MFRE}): among $q$-rough
shifted squarefree integers $m$ (those with $\mu^2(m{-}1)=1$ and
$\minFac(m) > q$), the smallest prime factor $\minFac(m)$ is
equidistributed modulo~$q$
(\lean{EM/PopulationEquidistProof.lean}{MinFacResidueEquidist}).
MFRE in turn implies PopulationEquidist~(PE) via
\code{pe\_of\_equidist}.  ANT is not formalized in Lean---it is
assumed as a hypothesis in the master reduction---but it follows from
results that are entirely standard in analytic number theory.

\subsection{The Decorrelation--Variance--Concentration Chain}
\label{sec:decorr-chain}

The ensemble character energy is the key object connecting CRT
equidistribution to MC\@.  For a character~$\chi$ and starting
point~$n$, the \emph{multiplier character energy} at length~$K$ is
\[
  E(n, K) \;=\; \Bigl\|\sum_{k < K} \chi\bigl(\seq{n}(k) \bmod q\bigr)\Bigr\|^2.
\]
This is an energy in the character spectrum: $E(n, K)$ lies between $0$
(perfect cancellation) and $K^2$ (no cancellation).  CCSB asks that
$E(n, K) = o(K^2)$ for the standard orbit $n = 2$; the ensemble
framework asks about the distribution of $E(n, K)$ over squarefree~$n$.

\begin{property}[\defname{StepDecorrelation} (\abbr{SD})]
For every prime~$q$, character~$\chi$ with $\|\chi(a)\|^2 \leq 1$, and
steps $j < k$: the ensemble cross-term vanishes:
\[
  \lim_{X \to \infty}
  \mathbb{E}_{n \leq X}^{\mathrm{sqf}}
  \bigl[\mathrm{Re}\bigl(\chi(m_j) \cdot \overline{\chi(m_k)}\bigr)\bigr]
  \;=\; 0,
\]
where $m_j = \seq{n}(j) \bmod q$ and the expectation is
over squarefree~$n$.
\end{property}

Step Decorrelation says that the character values at different steps are
asymptotically uncorrelated when averaged over starting points.  CRT
independence provides the structural basis: between steps $j$ and $k$,
each ``$+1$ shift'' in the Euclid construction decorrelates the mod-$q$
information.

\begin{property}[\defname{CharSumVarianceBound} (\abbr{CSVB})]
There exists $C > 0$ such that for all $q$, $\chi$, and $K$:
$\mathbb{E}_{n}^{\mathrm{sqf}}[E(n, K)] \leq C \cdot K$
eventually in~$X$.
\end{property}

\begin{theorem}[\lean{EM/EnsemblePT.lean}{decorrelation\_implies\_variance\_proved}]
\label{thm:variance-bound}
$\mathrm{SD} \;\Longrightarrow\; \mathrm{CSVB}$ with $C = 2$.
\end{theorem}

\begin{proof}
By induction on~$K$.

\emph{Base case.}  $E(n, 0) = 0 \leq 0$ for all~$n$.

\emph{Inductive step.}  The energy recurrence gives
\[
  E(n, K{+}1) \;=\; E(n, K) + \|\chi(m_K)\|^2 +
  2\,\mathrm{Re}\Bigl(\sum_{j < K} \chi(m_j) \cdot \overline{\chi(m_K)}\Bigr).
\]
Taking ensemble averages:
\begin{itemize}[nosep]
\item \textbf{Diagonal}: $\mathbb{E}[\|\chi(m_K)\|^2] \leq 1$ since
  $\|\chi(a)\|^2 \leq 1$ pointwise.
\item \textbf{Inductive term}: $\mathbb{E}[E(n, K)] \leq 2K$ by the
  inductive hypothesis (for $X \geq X_0^{(K)}$).
\item \textbf{Cross-terms}: for each $j < K$, SD gives
  $|\mathbb{E}[\mathrm{Re}(\chi(m_j)\overline{\chi(m_K)})]|
  < 1/(2(K{+}1))$ for~$X$ large enough.
  Summing the $K$ terms: $2|\text{cross sum}| \leq 1$.
\end{itemize}
Combining: $\mathbb{E}[E(n, K{+}1)] \leq 2K + 1 + 1 = 2(K{+}1)$.
The threshold $X_0^{(K+1)}$ is the maximum of $X_0^{(K)}$ and the $K$
individual cross-term thresholds.
\end{proof}

The constant $C = 2$ is sharp for this method: each step contributes
at most $1$ from the diagonal and at most $1$ from the cross-terms.
For comparison, independent random variables with $|\chi(m)| = 1$ give
$\mathbb{E}[E(n, K)] = K$ exactly (perfect cancellation of
cross-terms); $C = 2$ allows the cross-terms to contribute up to their
maximum.

\begin{property}[\defname{EnsembleCharSumConcentration}]
For every $q$, $\chi$, and $\varepsilon > 0$: the density of
squarefree $n \leq X$ with $E(n, K) > (\varepsilon K)^2$ tends to~$0$
as $X \to \infty$, uniformly in~$K$ past a threshold.
\end{property}

\begin{theorem}[\lean{EM/EnsembleDecorrelation.lean}{char\_variance\_implies\_concentration\_proved}]
CSVB implies EnsembleCharSumConcentration.
\end{theorem}

\begin{proof}[Proof sketch]
Markov's inequality.  Fix $\varepsilon > 0$.  The density of squarefree
$n$ with $E(n, K) > (\varepsilon K)^2$ is at most
$\mathbb{E}[E(n, K)] / (\varepsilon K)^2 \leq 2K / (\varepsilon^2 K^2)
= 2/(\varepsilon^2 K)$, which tends to $0$ as $K \to \infty$.
\end{proof}

\begin{theorem}[\lean{EM/EnsembleDecorrelation.lean}{char\_concentration\_implies\_cancellation}]
EnsembleCharSumConcentration $\Longrightarrow$ for almost all
squarefree~$n$, the character energy $E(n, K) = o(K^2)$.
\end{theorem}

\begin{proof}[Proof sketch]
The set of $n$ with $E(n, K) > (\varepsilon K)^2$ \emph{for all}~$K$
is contained in the set for any specific~$K_0$.  Choosing $K_0$
large enough (from concentration) makes this density arbitrarily
small.
\end{proof}

\noindent The three theorems compose into a single proved reduction:

\begin{theorem}[\lean{EM/EnsemblePT.lean}{sd\_implies\_cancellation}]
\label{thm:sd-cancel}
$\mathrm{StepDecorrelation}$ alone $\Longrightarrow$ character sum
cancellation for almost all squarefree starting points.
\end{theorem}

\noindent This composition uses zero open hypotheses: StepDecorrelation is
the sole input, and the three bridges (variance, concentration,
cancellation) are all machine-verified.

\paragraph{First Moment, Sieve Density, and the Buchstab Weight}
\label{sec:first-moment}

The ensemble average of the reciprocal $1/\seq{n}(k)$
connects to classical sieve theory via the \emph{sieve density
function}:
\[
  g(r) \;=\; \frac{r}{r^2 - 1},
\]
which gives the conditional probability that a random shifted
squarefree number $m$ (with $\mu^2(m{-}1) = 1$) is divisible by a
prime~$r$.  The algebraic properties of~$g$ are fully proved in
\code{ShiftedSquarefreeDensity.lean}:

\begin{itemize}[nosep]
\item $g(r) > 0$ for $r \geq 2$
  (\lean{EM/ShiftedSquarefreeDensity.lean}{sieveDensity\_pos}).
\item $g(r) < 1$ for $r \geq 2$
  (\lean{EM/ShiftedSquarefreeDensity.lean}{sieveDensity\_lt\_one}).
\item $g(r) > 1/r$
  (\lean{EM/ShiftedSquarefreeDensity.lean}{sieveDensity\_gt\_inv}):
  a density boost from automatic squarefreeness.
\item $g(r) = \tfrac{1}{2}\bigl(\tfrac{1}{r-1} + \tfrac{1}{r+1}\bigr)$
  (\lean{EM/ShiftedSquarefreeDensity.lean}{sieveDensity\_partial\_frac}).
\item $g$ is strictly decreasing on $[2, \infty)$
  (\lean{EM/ShiftedSquarefreeDensity.lean}{sieveDensity\_strict\_anti}).
\end{itemize}

The \emph{Buchstab weight} $\kappa(p)$ is the sieve density at the
base prime~$p$:

\begin{theorem}[\lean{EM/EnsembleFirstMoment.lean}{buchstabWeight\_two}]
$\kappa(2) = g(2) = 2/3$.
\end{theorem}

The first-moment/variance pattern gives a concentration result for
reciprocal sums:

\begin{theorem}[\lean{EM/EnsembleFirstMoment.lean}{first\_moment\_variance\_implies\_rsd}]
First Moment + Variance Bound $\Longrightarrow$ Reciprocal Sum
Divergence: $\sum_{k < K} 1/\seq{n}(k) \to \infty$ for
almost all squarefree~$n$.
\end{theorem}

\subsection{The DSL Route to MC}
\label{sec:dsl-route}

\begin{property}[\defname{DSL} ---
  \lean{EM/PopulationTransferStrategy.lean}{Deterministic\-Stability\-Lemma}]
\label{def:dsl}
Population equidistribution of multiplier residues implies conditional
multiplier equidistribution for the standard orbit ($n = 2$):
\[
  \mathrm{PE} \;\Longrightarrow\; \mathrm{CME}.
\]
Explicitly: if for every prime~$q$ and
$a \in (\ZZ/q\ZZ)^\times$ the density of
primes $p \equiv a \pmod{q}$ tends to $1/(q{-}1)$, then for
every missing~$q$, non-trivial~$\chi$, walk position
$c \in (\ZZ/q\ZZ)^\times$, and $\varepsilon > 0$:
\[
  \exists\, N_0 \;\;\forall\, N \geq N_0 :\quad
  \Bigl\|\sum_{\substack{n < N \\ \walkZ{2}{q}{n} = c}}
    \chi\bigl(\multZ{2}{q}{n}\bigr)\Bigr\|
  \;\leq\; \varepsilon \cdot N.
\]
\end{property}

DSL is an \emph{orbit-transfer} statement: it asserts that
population-level equidistribution (which classical analytic number
theory can establish) propagates to the specific deterministic orbit
$n = 2$.  Note the contrast with the \emph{population concentration}
property (Theorem~\ref{thm:sd-cancel}), which is already proved:
SD implies that \emph{almost all} squarefree starting points have
cancelling character energy.  But ``almost all'' does not cover
$n = 2$ specifically.  DSL bridges this gap.

\begin{theorem}[\lean{EM/MasterReduction.lean}{full\_chain\_dsl}]
\label{thm:dsl-mc}
$\mathrm{DSL} \;\Longrightarrow\; \MC$.
\end{theorem}

\begin{proof}[Proof sketch]
The proof assembles a five-link chain:
\begin{enumerate}[nosep]
\item \textbf{ANT $\Rightarrow$ PE}: MinFacResidueEquidist (a standard
  consequence of PNT in arithmetic progressions) gives
  PopulationEquidist
  (\lean{EM/PopulationEquidistProof.lean}{pe\_of\_equidist}).
\item \textbf{DSL: PE $\Rightarrow$ CME}: DSL converts population
  equidistribution to conditional multiplier equidistribution for the
  actual orbit.
\item \textbf{CME $\Rightarrow$ CCSB}: proved for all character
  orders, bypassing the $d \geq 3$ barrier
  (\lean{EM/LargeSieveSpectral.lean}{cme\_implies\_ccsb}).
\item \textbf{CCSB $\Rightarrow$ hits}: the Fourier bridge
  (Theorem~\ref{thm:fourier-step}) forces the hit count at~$-1$ to be
  eventually positive.
\item \textbf{Hits $\Rightarrow$ MC}: the Single Hit Theorem
  (Theorem~\ref{thm:single-hit}).
\end{enumerate}
Every arrow is fully machine-verified.  This collapses the entire
reduction landscape to a single open hypothesis.
\end{proof}

\begin{remark}[Why DSL should hold]
\label{rem:why-dsl}
DSL asserts that population equidistribution of $\minFac$ residues
transfers to the specific EM orbit.  Three structural features of the
EM sequence support this expectation.

\smallskip
\noindent\textbf{(1) The information bottleneck.}\enspace
At step~$n$, $\Prod{2}(n)$ has ${\sim}2^n$ bits, but the walk
position $\walkZ{2}{q}{n} = \Prod{2}(n) \bmod q$ carries only
$O(\log q)$ bits.  The smallest prime factor of $\Prod{2}(n)+1$
depends on the full ${\sim}2^n$-bit value, not on its $O(\log q)$-bit
residue.  Conditioning on $\walkZ{2}{q}{n} = c$ constrains almost
nothing about the factoring behavior of $\Prod{2}(n)+1$: the
conditional and unconditional distributions of
$\minFac(\Prod{2}(n)+1) \bmod q$ should agree, which is exactly what
CME says.  This is the same mechanism that makes iterated hash chains
pass statistical tests for independence: the forward map is easy, but
projecting an exponentially large integer onto a small residue
destroys all correlation.

\smallskip
\noindent\textbf{(2) No arithmetic obstruction.}\enspace
The products $\Prod{2}(n)$ are squarefree
(\lean{EM/WeakErgodicity.lean}{prod\_squarefree}), pairwise
multiplicatively entangled (each divides the next), and each
$\Prod{2}(n)+1$ has a distinct smallest prime factor
($\seq{2}$ is injective).  There is no known
arithmetic mechanism---congruence obstruction, algebraic identity, or
Galois constraint---that would systematically bias
$\minFac(\Prod{2}(n)+1)$ toward or away from particular residue
classes mod~$q$, conditional on the walk position.  Every structural
feature that \emph{could} create such a bias (coprime cascade, CRT
propagation, cofactor identity) has been exhaustively checked and
found insufficient (\S\ref{sec:hard}).

\smallskip
\noindent\textbf{(3) The ensemble evidence.}\enspace
The ensemble framework parametrizes the EM construction by squarefree
starting points.  For \emph{almost all} starting points~$n$, CME
holds: the proved StepDecorrelation~$\Rightarrow$~cancellation chain
(\lean{EM/EnsemblePT.lean}{sd\_implies\_cancellation})
shows that generic orbits satisfy the character sum bounds.  DSL asks
that the standard orbit $n = 2$ is not exceptional within this
ensemble.  Since $n = 2$ has no special arithmetic property that
distinguishes it from a typical squarefree starting point---it is not
a prime power, not a perfect square, not in any special congruence
class---there is no reason to expect it to be an outlier.

\smallskip
DSL is therefore a \emph{typicality assertion}: the EM orbit behaves
like a generic member of its ensemble.  This is analogous to
Hooley's conditional proof of Artin's conjecture~\cite{Hooley1967},
which shows the conjecture for all integers conditional on GRH---the
hypothesis being that the specific integers of interest are not
exceptional among all integers.  The difficulty of proving DSL is
the difficulty of proving that a specific deterministic sequence
satisfies a statistical law.
\end{remark}

\begin{remark}[\lean{EM/CRTPointwiseTransfer.lean}{oce\_eq\_cme}]
An alternative perspective on DSL is the \emph{CRT pointwise transfer}:
condition on the walk position $c = \walkZ{2}{q}{n}$ and ask whether
the multiplier residues in each fiber are equidistributed.
This \emph{orbit-conditional equidistribution}~(OCE) turns out to be
definitionally equal to CME---proved by \code{rfl}, since the walk
position \emph{is} the residue class that CME conditions on.
Any route to OCE automatically gives CME, and the CRT bridge
(population conditional equidistribution~$\Rightarrow$~OCE) is
subsumed by DSL
(\lean{EM/CRTPointwiseTransfer.lean}{dsl\_implies\_crt\_bridge}).
\end{remark}

A weaker variant also suffices:

\begin{property}[\defname{DSLHitting}]
$\mathrm{PE} \Rightarrow \mathrm{DH}$: population equidistribution
implies that the walk hits~$-1$ cofinally.  This is strictly weaker
than DSL (which gives $\mathrm{PE} \Rightarrow \mathrm{CME}$):
DSLHitting asks only for the walk to visit $-1$, not for full
character sum cancellation.
\end{property}

\begin{theorem}[\lean{EM/EnsembleStructure.lean}{pe\_dsl\_hitting\_implies\_mc}]
$\mathrm{PE} + \mathrm{DSLHitting} \;\Longrightarrow\; \MC$.
\end{theorem}

\noindent The reduction is immediate:
$\mathrm{PE} \xrightarrow{\text{DSLHitting}} \mathrm{DH}
\xrightarrow{\text{proved}} \mathrm{MC}$.

\begin{remark}[The tail identity and DSL]\label{rem:tail-dsl}
The tail identification (Theorem~\ref{thm:tail-id}) gives a precise
structural link between the standard orbit and the ensemble.  Since
$\Prod{2}(M)$ is squarefree for all~$M$
(\lean{EM/WeakErgodicity.lean}{prod\_squarefree}), the $M$-th
accumulator is itself a valid ensemble member.  The tail identity says
the orbit starting from $\Prod{2}(M)$ \emph{is} the suffix of the
standard orbit from step~$M$ onward:
\[
  \Prod{\Prod{2}(M)}(k) \;=\; \Prod{2}(M + k),
  \qquad
  \seq{\Prod{2}(M)}(k) \;=\; \seq{2}(M + k).
\]
Consequently, the character energy of the ensemble member
$n = \Prod{2}(M)$ over a window of length~$K$ equals the character
energy of the standard orbit over~$[M, M{+}K)$:
\[
  E\bigl(\Prod{2}(M),\, K\bigr)
  \;=\; E_{\mathrm{tail}}(2,\, M,\, K).
\]
The ensemble concentration result
(SD~$\Rightarrow$~CSVB~$\Rightarrow$~concentration,
Theorem~\ref{thm:sd-cancel}) says: for almost all squarefree~$n$,
$E(n, K) = o(K^2)$.  So the question ``does the standard orbit have
cancelling character energy on long windows?''\ is the same as ``does
the ensemble member $\Prod{2}(M)$ fall outside the bad set?''---for
every~$M$.

\medskip
\noindent\textbf{The Borel--Cantelli program.}\enspace
Let $B_\varepsilon(K)$ denote the set of squarefree~$n$ with
$E(n,K) > (\varepsilon K)^2$.  The first-moment (Markov) bound gives
bad density~$O(1/K)$, which is not summable over windows and
therefore insufficient for a Borel--Cantelli argument.  The
\emph{second-moment} (Chebyshev) approach resolves this: if the
population fourth moment satisfies
$\mathbb{E}_{\mathrm{sqf}}[E(n,K)^2] = O(K^2)$, then Chebyshev
gives
\[
  \Pr_{\mathrm{sqf}}\bigl[n \in B_\varepsilon(K)\bigr]
  \;\leq\; \frac{D}{\varepsilon^4 K^2},
\]
which \emph{is} summable.  Applying this to dyadic windows
$[2^i, 2^{i+1})$ with $K = 2^i$:
$\sum_i D/(\varepsilon^4 \cdot 4^i) < \infty$, so by
Borel--Cantelli, for all large~$i$ the character energy on the window
$[2^i, 2^{i+1})$ is at most $(\varepsilon \cdot 2^i)^2$.  Summing
over windows covering~$[0,N)$:
\[
  |F(c,\,[0,N))| \;\leq\;
  \sum_{i:\, 2^i < N} \varepsilon \cdot 2^i
  \;\leq\; 2\varepsilon N,
\]
which gives $|F(c,N)| = o(N)$, i.e.\ CME.

\medskip
\noindent\textbf{The sole new content: four-point PCV.}\enspace
The fourth moment $\mathbb{E}[E(n,K)^2]$ decomposes into a diagonal
contribution~$K^2$, a cross-term contribution that vanishes by PCV,
and a \emph{four-point correlation}
\[
  \Gamma_4(j_1,k_1,j_2,k_2)
  \;=\; \mathbb{E}_{\mathrm{sqf}}\bigl[
    \mathrm{Re}(\chi(m_{j_1})\overline{\chi(m_{k_1})})
    \cdot
    \mathrm{Re}(\chi(m_{j_2})\overline{\chi(m_{k_2})})
  \bigr].
\]
For distinct pairs $(j_1,k_1) \neq (j_2,k_2)$ with all four indices
distinct, CRT independence should force $\Gamma_4 \to 0$ via
factorization into two-point correlations.  If this \emph{four-point
PCV} holds, $\mathbb{E}[E(n,K)^2] = O(K^2)$, and the
Borel--Cantelli program closes DSL.

\medskip
\noindent\textbf{The Substitution Principle (and its limits).}\enspace
An equivalent reformulation of CME, proved definitionally equal
(\code{sp\_eq\_cme}), works at the level of \emph{fiber return visits}:
the steps~$k$ where $\walkZ(q,k) = a$.  At such steps the multiplier
primes $\seq(k{+}1)$ are \emph{distinct} (\code{return\_visit\_mult\_injective})
and \emph{pairwise coprime} (\code{return\_visit\_mult\_coprime}).
The structural package is recorded in a 4-tuple witness
\code{sp\_structural\_summary}: injectivity, coprimality, primality,
and divisibility.  Four independent routes from SP to MC are assembled
in \code{all\_routes\_to\_mc\_with\_sp}.

However, the coprime cascade structure alone does \emph{not} force
character cancellation: for any $q \geq 3$, one can construct
pairwise coprime integers $N_i = p_i r_i$ in a fixed residue class
$\bmod q$ whose least prime factors all lie in the same residue class
(e.g.\ $p_i \equiv 2$, $r_i \equiv 3 \pmod{5}$ gives
$\sum \chi(\mathrm{minFac}(N_i)) = I \cdot \chi(2) = \Theta(I)$).
Thus the cancellation in CME, if true, must rely on the specific
feedback dynamics of the EM sequence---the fact that each
$\seq(n{+}1) = \mathrm{minFac}(\Prod{}(n){+}1)$ depends on all prior
terms---not merely on coprimality.  The ``Substitution Principle'' is
therefore just CME under a structurally suggestive name; the name does
not isolate a new mechanism.

\medskip
\noindent\textbf{The orbit-specificity gap.}\enspace
Even the Chebyshev bound is a statement about the \emph{density} of
bad starting points, not a pointwise statement about the specific
values~$\Prod{2}(M)$.  A density-zero set can still contain a
specific super-exponentially growing sequence.  However, the
$O(1/K^2)$ rate makes the gap much narrower: combined with CRT
typicality of $\Prod{2}(M) \bmod q$ (the residues form a walk on
$(\mathbb{Z}/q\mathbb{Z})^\times$ visiting every class cofinally, by
the proved \code{SubgroupEscape} and
\code{prime\_residue\_escape}), the gap reduces to a quantitative
equidistribution-in-residue-classes statement.  This is the sharpest
form of the orbit-specificity barrier.
\end{remark}

\begin{remark}[Self-correcting feedback and visit equidistribution]
\label{rem:self-correcting-ve}
The preceding attacks target CME (conditional multiplier
equidistribution) or its equivalents.  A structurally different route
bypasses CME entirely: \textbf{visit equidistribution} (VE).

Define the empirical frequency vector
\[
  f_c(N) \;=\; \frac{1}{N}\,\bigl|\{n < N : \walkZ{2}{q}{n} = c\}\bigr|,
  \qquad c \in (\ZZ/q\ZZ)^\times.
\]
VE asserts $f_c(N) \to 1/\varphi(q)$ for every~$c$.  By the
Weyl criterion this is equivalent to walk character cancellation
$\|\sum_{n<N} \chi(\walkZ{2}{q}{n})\|^2 = o(N^2)$, and the chain
\[
  \text{VE}
  \;\xleftrightarrow[\text{proved}]{\text{Weyl}}\;
  \text{WalkCharCancel}
  \;\xrightarrow[\text{proved}]{\text{WeylHittingBridge}}\;
  \text{MC}
\]
is fully machine-verified
(\code{sve\_iff\_ve}, \code{cancel\_weyl\_implies\_mc}).

\medskip
\noindent\textbf{VE is strictly different from CME.}\enspace
CME controls the \emph{conditional} distribution of multipliers given
walk position; VE controls the \emph{marginal} distribution of walk
positions.  Neither implies the other: equidistributed walk positions
do not force equidistributed conditional multipliers, and vice versa.
This is not an equivalence collapse.

\medskip
\noindent\textbf{The self-correcting feedback mechanism.}\enspace
The EM walk has a built-in negative feedback loop.  Suppose $f_c(N)$
is anomalously large: the walk visits position~$c$ too often.
At each such visit, $\Prod{2}(n) \equiv c \pmod{q}$, so
$\Prod{2}(n) + 1 \equiv c + 1 \pmod{q}$.  The next multiplier is
$\seq{}{n{+}1} = \minFac(\Prod{2}(n) + 1)$, an integer whose residue
class mod~$q$ is determined by $\Prod{2}(n) + 1$.  By Alladi's
theorem~\cite{Alladi1977} (axiomatized as
\code{GenericLPFEquidist}), the least prime factors of rough integers
in a fixed residue class are equidistributed among all nonzero
residues mod~$q$.  If the multiplier at position~$c$ is approximately
equidistributed, the transition $c \mapsto c \cdot m(n)$ scatters the
walk away from~$c$ toward all other positions---a negative feedback
that opposes further concentration at~$c$.

This has been formalized via a Lyapunov function
(\code{SelfCorrectingDrift.lean}).  Define
\[
  L(N) \;=\; \sum_{c \in (\ZZ/q\ZZ)^\times}
  \Bigl(V_c(N) - \frac{N}{\varphi(q)}\Bigr)^{\!2},
\]
where $V_c(N) = |\{n < N : \walkZ{2}{q}{n} = c\}|$ is the visit count.
The one-step algebraic identity
(\lean{EM/SelfCorrectingDrift.lean}{lyapunov\_one\_step})
\[
  L(N{+}1) = L(N) + 2\,d_{\walkZ{2}{q}{N}}(N) + \frac{q-2}{q-1},
\]
where $d_c(N) = V_c(N) - N/(q{-}1)$ is the visit deviation at
position~$c$, telescopes to
(\lean{EM/SelfCorrectingDrift.lean}{lyapunov\_telescope})
\[
  L(N) = 2\,R(N) + N\,\frac{q-2}{q-1},
  \qquad R(N) = \sum_{n<N} d_{\walkZ{2}{q}{n}}(n).
\]
The cumulative drift $R(N)$ records how often the walk visits
positions that are already over-represented.  Under a uniform
multiplier measure, $\mathbb{E}[d_c] = 0$ for every position
(\lean{EM/SelfCorrectingDrift.lean}{uniform\_multiplier\_zero\_drift}),
and the transition matrix is doubly stochastic
(\lean{EM/SelfCorrectingDrift.lean}{group\_walk\_doubly\_stochastic}),
so the drift has zero bias.
\textbf{SelfCorrectingDrift}~(SCD) asserts $R(N) = o(N^2)$, which
gives $L(N) = o(N^2)$, hence VE, and therefore MC via the proved
chain (\lean{EM/SelfCorrectingDrift.lean}{scd\_implies\_mc}).

\medskip
\noindent\textbf{SCD $\equiv$ SVE (Dead End~\#120).}\enspace
The Lyapunov telescope $L(N) = 2R(N) + O(N)$ shows that SCD
($R(N) = o(N^2)$) is algebraically equivalent to SVE ($L(N) = o(N^2)$)
for each fixed prime~$q$ above the threshold.  The converse is also
proved: SVE implies SCD above the SVE threshold
(\lean{EM/SelfCorrectingDrift.lean}{sve\_implies\_scd\_above\_threshold}).
Therefore SCD provides no new proof leverage beyond SVE; it is a
Lyapunov reformulation that clarifies the dynamical mechanism
(self-correcting feedback) but does not reduce the mathematical
difficulty.

\medskip
\noindent\textbf{What remains.}\enspace
The residual difficulty is orbit-specificity: the Alladi-type
equidistribution holds for ``generic'' rough integers, but
$\Prod{2}(n) + 1$ is a specific value, not a random draw from its
residue class.  The Lyapunov analysis shows that proving
$R(N) = o(N^2)$ requires controlling how often the walk visits
positions that are already over-represented---and this control depends
on the specific EM orbit, not just generic multiplier statistics.
This is the same orbit-specificity barrier
from~\S\ref{sec:landscape}, now expressed at the level of single-step
drift rather than long-range character cancellation.

\medskip
\noindent\textbf{Per-class bad-set escape (Dead End~\#121).}\enspace
A natural attempt: use SMSB (SecondMomentSquaredBound) to bound the
\emph{global} density of ``bad'' steps (those whose tail has large
character energy), then argue that each walk-position class has
proportionally few bad steps, then use SubgroupEscape to get escape
from each class.  The per-class bad sets partition the total bad set
(\lean{EM/SMSBPerClassEscape.lean}{badVisitsInClass\_card\_sum\_eq}),
and pigeonhole guarantees \emph{some} class with small bad density
(\lean{EM/SMSBPerClassEscape.lean}{exists\_class\_small\_bad\_set}).
But MC requires the specific class~$-1$, not an existential over
classes.  Targeting~$-1$ requires that bad steps do not concentrate
in one class---a per-fiber density statement
(BadSetEquidistribution) that is a consequence of CME\@.  Going from
the marginal bad-density bound (what SMSB provides) to the conditional
bound (what per-class escape requires) is blocked by the
Marginal/Joint Barrier: the joint distribution of (walk position,
badness) is controlled by CME, not by SMSB alone.
\end{remark}

\subsection{Generalized MC and the Weyl Hitting Bridge}
\label{sec:gen-mc}
\label{sec:weyl-bridge}

The ensemble framework requires proving MC not just for $n = 2$ but for
arbitrary squarefree starting points.  This motivates two key results.

\begin{definition}[\defname{GenMullinConjecture}$(n)$]
Every prime appears in the generalized EM sequence from starting
point~$n$: $\forall\, q$ prime, $\exists\, k$,
$\seq{n}(k) = q$.
\end{definition}

\begin{theorem}[\lean{EM/EnsemblePT.lean}{gen\_hitting\_implies\_gen\_mc\_proved}]
\label{thm:gen-mc}
For squarefree $n \geq 1$: if the generalized walk hits~$-1$
cofinally for every prime~$q$,
then $\mathrm{GenMullinConjecture}(n)$ holds.
\end{theorem}

\begin{proof}
Strong induction on~$q$.  Assume every prime $p < q$ appears as
$\seq{n}(k_p)$ for some~$k_p$.  Collect all witnesses into
a uniform bound~$N$.  Cofinal hitting at~$q$ provides $k \geq N$ with
$q \mid \Prod{n}(k) + 1$.  Since all primes $p < q$ already appeared
at steps $< k$, they divide $\Prod{n}(k)$ and hence cannot divide
$\Prod{n}(k) + 1$.  So $q = \minFac(\Prod{n}(k) + 1) =
\seq{n}(k)$.

The argument is a generalized \code{captures\_target} lemma
(\lean{EM/EnsemblePT.lean}{gen\_captures\_target}), paralleling the
standard version from \S\ref{sec:bootstrap}.
\end{proof}

The connection between character cancellation and hitting uses the Weyl
equidistribution criterion, adapted to the multiplicative setting:

\begin{theorem}[\defname{WeylHittingBridge} ---
  \lean{EM/EnsemblePT.lean}{weyl\_hitting\_bridge\_proved}]
\label{thm:weyl-bridge}
If all non-trivial walk character sums satisfy
$\|\sum_{k<K} \chi(\walkZ{n}{q}{k})\|^2 \leq (\varepsilon K)^2$
for large~$K$, then either $q$ appears in $\seq{n}(\cdot)$
or the walk hits~$-1$ cofinally.
\end{theorem}

\begin{proof}
By contradiction using a \emph{Weyl test function}.  Suppose $q$ never
appears and the walk stops hitting~$-1$ past some step~$N_0$.  Then the
hit count $H(K) = |\{k < K : \Prod{n}(k) \equiv -1 \pmod{q}\}|$ is
bounded by the constant $C = H(N_0)$ for all~$K$.

Define a centered test function $\chi_{\mathrm{test}} : \NN \to \mathbb{C}$:
\[
  \chi_{\mathrm{test}}(x) \;=\;
  \begin{cases}
    0 & \text{if } x \equiv 0 \pmod{q}, \\
    (q{-}2)/(q{-}1) & \text{if } x \equiv -1 \pmod{q}, \\
    -1/(q{-}1) & \text{otherwise.}
  \end{cases}
\]
The test function satisfies $\|\chi_{\mathrm{test}}\|^2 \leq 1$,
$\chi_{\mathrm{test}}(0) = 0$, and
$\sum_{a \in \ZZ/q\ZZ} \chi_{\mathrm{test}}(a) = 0$ (the sum of
$(q{-}2)/(q{-}1)$ from $a = q{-}1$ and $(q{-}2) \cdot (-1/(q{-}1))$
from the remaining units is zero).

The partial sum of the test function along the walk telescopes to
$\sum_{k<K} \chi_{\mathrm{test}}(\Prod{n}(k)) = H(K) \cdot
(q{-}2)/(q{-}1) + (K - H(K)) \cdot (-1/(q{-}1))
= H(K) - K/(q{-}1)$.
So the walk character energy equals
\[
  E_{\mathrm{test}}(K) \;=\; \bigl(H(K) - K/(q{-}1)\bigr)^2.
\]
Since $H(K) \leq C$ while $K/(q{-}1) \to \infty$, for large~$K$ we have
$E_{\mathrm{test}}(K) \geq (K/(q{-}1) - C)^2$.  Setting
$\varepsilon = 1/(2(q{-}1))$, the hypothesis gives
$E_{\mathrm{test}}(K) \leq (K/(2(q{-}1)))^2$.  Taking square roots:
$K/(q{-}1) - C \leq K/(2(q{-}1))$, hence $K/(2(q{-}1)) \leq C$,
so $K \leq 2C(q{-}1)$.  But $K$ was chosen larger than $2Cq + 1$:
contradiction.
\end{proof}

\begin{theorem}[\lean{EM/EnsemblePT.lean}{cancel\_weyl\_implies\_mc}]
\label{thm:cancel-weyl-mc}
Walk character cancellation at $n = 2$ implies MC.
\end{theorem}

\begin{proof}
For $q = 2$: $\seq{2}(0) = 2$.  For $q \geq 3$: $q$ does not divide~$2$,
so the Weyl Hitting Bridge gives either $q \in \seq{2}(\cdot)$
(hence $q \in \{\seq{2}(k)\}$) or cofinal hitting.  Cofinal hitting gives
$q \in \seq{2}(\cdot)$ by Theorem~\ref{thm:gen-mc}, noting
that all generalized primes from $n = 2$ are odd (since $2 \mid \Prod{2}(k)$
implies $\Prod{2}(k) + 1$ is odd), so the induction only requires
odd primes $< q$, all of which have already appeared by the
inductive hypothesis.
\end{proof}

\paragraph{The Per-Chi Cancellation Bridge}
\label{sec:per-chi}

The decorrelation chain of \S\ref{sec:decorr-chain} works with
\emph{universal} StepDecorrelation (quantified over all characters).
For the JSE route, only \emph{per-chi} decorrelation is available (for
each individual non-trivial character).  The PerChiCancellationBridge
shows the same three-step chain works per-chi:

\begin{theorem}[\lean{EM/EnsemblePT.lean}{per\_chi\_cancellation\_bridge\_proved}]
\label{thm:per-chi}
\defname{PerChiCancellationBridge}: for each prime~$q$ and
character~$\chi$ with $\|\chi\|^2 \leq 1$, if the cross-terms
$\mathbb{E}^{\mathrm{sqf}}[\mathrm{Re}(\chi(m_j) \cdot
\overline{\chi(m_k)})] \to 0$ for all $j < k$, then the density of
squarefree~$n$ with $E(n, K) > (\varepsilon K)^2$ for all~$K$ tends
to zero.
\end{theorem}

\begin{proof}
The proof is identical to the universal chain (Theorem~\ref{thm:variance-bound}
$\to$ concentration $\to$ cancellation), with the StepDecorrelation
hypothesis restricted to the fixed character~$\chi$.  Each step
(variance with $C = 2$, Markov concentration, squeeze to zero) goes
through unchanged because none requires decorrelation of \emph{other}
characters.
\end{proof}

\paragraph{The Ensemble Chain in Detail}
\label{sec:ensemble-chain}

The full ensemble assembly uses the following master theorem:

\begin{theorem}[\lean{EM/EnsemblePT.lean}{ensemble\_pt\_master\_simplified}]
Under four hypotheses (SRE, CRTPropagationStep,
AccumEquidistImpliesMultEquidist, EnsembleEquidistImpliesDecorrelation),
character sums cancel for almost all squarefree starting points.
\end{theorem}

\noindent The chain:
\[
  \underbrace{\mathrm{SRE} + \mathrm{CRT} + \mathrm{Bridge}}_{\text{ANT + CRT structure}}
  \;\xrightarrow{\text{proved}}\;
  \mathrm{EME}
  \;\xrightarrow{\text{open}}\;
  \mathrm{SD}
  \;\xrightarrow{\text{proved}}\;
  \text{cancellation for a.a.~}n.
\]
The first arrow composes three proved reductions.  The second arrow
(EME~$\Rightarrow$~SD) is the sole remaining open hypothesis.  The
third arrow is Theorem~\ref{thm:sd-cancel}.

\paragraph{The JSE route.}
An alternative chain uses \emph{JointStepEquidist}~(JSE): the joint
distribution of $(\seq{n}(j), \seq{n}(k))$
is asymptotically uniform over non-zero pairs.  The JSE base case
($j = 0, k = 1$) is proved via the tower rule
(\lean{EM/EnsembleBaseCase.lean}{jse\_base\_case}):
$P(A, B) = P(A \mid B) \cdot P(B)$, decomposing the joint density
into conditional and marginal equidistribution.

\begin{theorem}[\lean{EM/EnsemblePT.lean}{jse\_transfer\_implies\_mc}]
$\mathrm{JSE} + \mathrm{MultCancelToWalkCancel}
\;\Longrightarrow\; \MC$.
\end{theorem}

\noindent The composition:
\begin{enumerate}[nosep]
\item JSE gives per-chi decorrelation for non-trivial~$\chi$
  (\lean{EM/EnsemblePT.lean}{joint\_step\_equidist\_implies\_step\_decorrelation}).
\item PerChiCancellationBridge (Theorem~\ref{thm:per-chi}) gives
  multiplier character cancellation for almost all squarefree~$n$.
\item \textbf{MultCancelToWalkCancel} (open): multiplier energy
  $o(K^2)$ implies walk energy $o(K^2)$.
\item WeylHittingBridge (Theorem~\ref{thm:weyl-bridge}) converts walk
  character cancellation into cofinal hitting.
\item Generalized MC (Theorem~\ref{thm:gen-mc}) and standard-EM
  bridge (Theorem~\ref{thm:cancel-weyl-mc}) give MC\@.
\end{enumerate}

The two remaining open hypotheses are:
\begin{itemize}[nosep]
\item \textbf{JointStepEquidist}: requires analytic number theory
  (Siegel--Walfisz or Bombieri--Vinogradov type estimates).
\item \textbf{MultCancelToWalkCancel}: multiplier character cancellation
  implies walk character cancellation.  Dead End~\#117 shows this is
  equivalent to CCSB/CME in difficulty: an explicit counterexample
  (alternating multipliers $\{2, 3\} \bmod 5$ with $\chi(2) = i$) gives
  multiplier cancellation yet $|W_K| = \Theta(K)$, and this
  counterexample is compatible with all EM structural properties
  (squarefree accumulator, coprime cascade, CRT invariance).
\end{itemize}

\noindent MultCancelToWalkCancel is the Marginal/Joint Barrier
(\S\ref{sec:oracle}) in ensemble form: the multiplier character sum is a
\emph{marginal} statistic (it depends on the individual multiplier
values), while the walk character sum is a \emph{joint} statistic (it
depends on all previous multipliers through the partial product).
Cancellation of individual terms does not imply cancellation of their
cumulative products.  This is why the DSL route
(Section~\ref{sec:dsl-route}) bypasses the ensemble-to-orbit gap
entirely: DSL gives CME directly, which implies walk character
cancellation without ever needing the multiplier-to-walk transfer.

\paragraph{The Variance Route to DSL}
\label{sec:variance-route}

The DSLVariance framework (\code{DSLVariance.lean}) provides an
alternative approach to DSL via the second moment of character energy.

\begin{definition}[\defname{Population Character Energy}]
The ensemble average of character energy:
$\mathcal{E}(q, \chi, K, X) = \mathbb{E}^{\mathrm{sqf}}_{n \leq X}[E(n, K)]$.
\end{definition}

\begin{theorem}[\lean{EM/DSLVariance.lean}{populationCharEnergy\_nonneg}]
$\mathcal{E} \geq 0$.
\end{theorem}

The population energy decomposes via cross-term pairs:

\begin{definition}[\defname{CrossTermPair}]
For steps $j < k$:
$\Gamma(j, k) = \mathbb{E}^{\mathrm{sqf}}_{n \leq X}
[\mathrm{Re}(\chi(m_j) \cdot \overline{\chi(m_k)})]$.
\end{definition}

\begin{theorem}[\lean{EM/DSLVariance.lean}{crossTermPair\_comm}]
$\Gamma(j, k) = \Gamma(k, j)$ (since $\mathrm{Re}(z \cdot \bar{w}) =
\mathrm{Re}(w \cdot \bar{z})$).
\end{theorem}

\begin{property}[\defname{PairwiseCrossTermVanishing} (\abbr{PCV})]
$\Gamma(j, k) \to 0$ as $X \to \infty$ for each fixed pair $j < k$.
\end{property}

\begin{theorem}[\lean{EM/DSLVariance.lean}{stepDecorrelation\_implies\_pairwiseVanishing}]
$\mathrm{SD} \;\Longrightarrow\; \mathrm{PCV}$.
\end{theorem}

\begin{theorem}[\lean{EM/DSLVariance.lean}{charSumVarianceBound\_implies\_secondMomentBound}]
$\mathrm{CSVB}(C) \;\Longrightarrow\; \mathrm{SecondMomentBound}(C)$:
the population energy is at most~$CK$.
\end{theorem}

The variance route aims to reach DSL via
$\mathrm{PCV} \Rightarrow \mathrm{SMB} \Rightarrow
\text{concentration} \Rightarrow \mathrm{DSL}$.
The first two arrows face a technical gap, and the last faces the
fundamental orbit-transfer barrier:
\begin{itemize}[nosep]
\item \textbf{PCV $\Rightarrow$ SMB}: requires \emph{uniform}
  convergence across all $K(K{-}1)/2$ cross-term pairs, not just
  pointwise.
\item \textbf{Concentration $\Rightarrow$ DSL}: the proved chain
  SD~$\Rightarrow$~CSVB~$\Rightarrow$~Concentration
  (Theorem~\ref{thm:sd-cancel}) already gives cancellation for
  \emph{almost all} squarefree starting points.  But DSL asks for
  CME at the specific orbit $n = 2$, and a density-one set of good
  starting points can exclude any particular~$n$.  This is the
  orbit-specificity barrier in its sharpest form.
\end{itemize}

\subsection{Visit Equidistribution and the Spectral Identity}
\label{sec:visit-eqd}

The ensemble framework connects to the spectral route via a
fundamental identity relating character energy to the geometry of the
walk.  For a walk of length~$N$ on $(\ZZ/q\ZZ)^\times$, define the
\emph{visit count} $V_N(a) = |\{n < N : \walkZ{2}{q}{n} = a\}|$.

\begin{theorem}[\lean{EM/VisitEquidistribution.lean}{excessEnergy\_eq\_visit\_deviation}]
\label{thm:excess-energy}
The excess energy $E_\chi(N) - N$ equals $(q{-}1)$ times the
squared deviation of visit counts from the uniform distribution:
\[
  \sum_{\chi \neq \chi_0} \bigl\|S_\chi(N)\bigr\|^2
  \;=\; (q{-}1) \sum_{a \in (\ZZ/q\ZZ)^\times}
  \Bigl(V_N(a) - \frac{N}{q{-}1}\Bigr)^2.
\]
\end{theorem}

This establishes the equivalence
\defname{SVE}~$\Leftrightarrow$~\defname{VE}
(\lean{EM/MasterReduction.lean}{sve\_iff\_ve}): subquadratic visit
energy (the total non-trivial character energy is $o(N^2)$) is the same
as visit equidistribution (the visit counts are approximately uniform).
The identity connects the Fourier-analytic perspective (character sums)
with the geometric perspective (visit counts) on the walk:
\begin{itemize}[nosep]
\item \textbf{SVE} says the walk has no sustained spectral bias:
  projecting onto any non-trivial character gives cancellation.
\item \textbf{VE} says the walk has no sustained spatial bias:
  it visits every position approximately equally often.
\end{itemize}
These are dual formulations of the same condition, connected by
Parseval's identity on the finite group $(\ZZ/q\ZZ)^\times$.

The fiber decomposition (\S\ref{sec:fiber-energy}) refines this further:

\begin{theorem}[\lean{EM/VisitEquidistribution.lean}{fiberMultCharSum\_step}]
The fiber character sum $F(a, N) = \sum_{\substack{n < N \\ w(n) = a}}
\chi(m(n))$ satisfies the recurrence:
$F(a, N{+}1) = F(a, N) + \chi(m(N)) \cdot \mathbf{1}_{w(N) = a}$.
\end{theorem}

\noindent This step-by-step evolution of fiber sums provides the
infrastructure for the CME~$\Rightarrow$~CCSB proof via energy
tracking: the total character energy decomposes into active-fiber
contributions at each step, and CME bounds each fiber's contribution.

\begin{remark}[Almost-all MC and the multiplier/walk gap]
\label{rem:almost-all-mc}
A natural question: since SD gives character sum cancellation for
\emph{almost all} squarefree starting points
(Theorem~\ref{thm:sd-cancel}), does this yield
$\mathrm{GenMullinConjecture}(n)$ for almost all~$n$?

It does not.  The cancellation that SD produces is for the
\textbf{multiplier} character energy
$\|\sum_{k<K} \chi(\seq{n}(k) \bmod q)\|^2 = o(K^2)$---the
individual multiplier residues have cancelling character sums.  But
the Weyl Hitting Bridge (Theorem~\ref{thm:weyl-bridge}), which
converts character cancellation into cofinal hitting, requires
\textbf{walk} character cancellation:
$\|\sum_{k<K} \chi(\walkZ{n}{q}{k})\|^2 = o(K^2)$.  The walk is a
partial product of the multipliers---$\walkZ{n}{q}{k} = \prod_{j<k}
\multZ{n}{q}{j}$---and cancellation of a sum $\sum \chi(m_k)$ does not
imply cancellation of $\sum \prod_{j<k} \chi(m_j)$.

This is Dead End~\#117 (MultCancelToWalkCancel, \S\ref{sec:oracle}):
an explicit counterexample (alternating multipliers $\{2,3\} \bmod 5$
with $\chi(2) = i$) achieves multiplier cancellation yet
$|W_K| = \Theta(K)$, and is compatible with all EM structural
properties.  The chain therefore breaks as:
\[
  \mathrm{SD}
  \;\xrightarrow{\text{proved}}\;
  \text{a.a.\ multiplier cancellation}
  \;\xrightarrow{\text{\#117}}\;
  \text{a.a.\ walk cancellation}
  \;\xrightarrow{\text{proved}}\;
  \text{a.a.\ GenMC}.
\]
The last arrow composes the Weyl bridge with
Theorem~\ref{thm:gen-mc}; the first arrow is
Theorem~\ref{thm:sd-cancel}; the middle arrow is the
marginal/joint barrier, which persists orbit by orbit.
Ensemble averaging helps with multiplier statistics but cannot cross
this barrier: the walk character sum is a \emph{joint} statistic
(depending on all previous multipliers through the partial product),
while the multiplier character sum is a \emph{marginal} one.

This is why DSL targets CME rather than multiplier cancellation:
CME goes directly from conditional multiplier equidistribution to
walk cancellation (via the fiber decomposition and telescoping),
bypassing MultCancelToWalkCancel entirely.
\end{remark}

\subsection{Tail Window Decorrelation}
\label{sec:twd}

The tail identity (Theorem~\ref{thm:tail-id}) connects temporal
blocks of the standard EM orbit to the ensemble framework: the
character sum over steps $jK{+}1, \ldots, (j{+}1)K$ of the standard
orbit equals the generalized partial sum starting from $\Prod{2}(jK)$.
This motivates decomposing the full character sum into non-overlapping
\emph{windows}.

\begin{definition}[\defname{WindowCharSum}]
The character sum over the $j$-th window of length~$K$:
$S_j = \sum_{k < K} \chi(\seq{2}(jK{+}k{+}1) \bmod q)$.
\end{definition}

\begin{theorem}[\lean{EM/TailWindowDecorrelation.lean}{windowCharSum\_eq\_genSeqCharPartialSum}]
$S_j$ equals the generalized character partial sum
starting from $\Prod{2}(jK)$:
$S_j = \sum_{k<K} \chi(\text{genSeq}(\Prod{2}(jK), k) \bmod q)$.
\end{theorem}

\begin{theorem}[\lean{EM/TailWindowDecorrelation.lean}{charSum\_eq\_window\_sum}]
\label{thm:window-decomp}
The total character sum decomposes:
$\sum_{k < JK} \chi(\seq{2}(k{+}1) \bmod q) = \sum_{j<J} S_j$.
\end{theorem}

Applying the variance identity
(\lean{EM/TailWindowDecorrelation.lean}{norm\_sq\_sum\_eq\_diag\_plus\_cross}),
the total character energy decomposes as
\[
  \Bigl\|\sum_{j<J} S_j\Bigr\|^2
  = \sum_{j<J} |S_j|^2
  + 2 \sum_{j < l < J} \mathrm{Re}(S_j \cdot \overline{S_l}).
\]
The diagonal terms are bounded by $J \cdot K^2$
(\lean{EM/TailWindowDecorrelation.lean}{diag\_energy\_le}).
The \emph{cross terms}---the off-diagonal correlations between
windows at different temporal positions---are the new quantity of
interest.

\begin{property}[\defname{TailWindowDecorrelation} (\abbr{TWD})]
The average absolute cross term between non-overlapping windows
vanishes as the window size~$K$ grows:
\[
  \frac{1}{J^2} \sum_{j < l < J}
  |\mathrm{Re}(S_j \cdot \overline{S_l})| \to 0.
\]
\end{property}

\begin{theorem}[\lean{EM/TailWindowDecorrelation.lean}{twd\_implies\_mc\_from\_bridge}]
$\mathrm{TWDImpliesCCSB} + \mathrm{TWD}
\;\Longrightarrow\; \MC$.
\end{theorem}

\noindent The proof composes the open bridge TWDImpliesCCSB with the
proved chain CCSB~$\Rightarrow$~MC\@.  The bridge TWDImpliesCCSB
would follow from TWD by choosing $K = \lceil 1/\varepsilon\rceil$:
the diagonal contributes $N \cdot K = O(\varepsilon^{-1} N)$ while the
cross terms contribute $O(\varepsilon N^2)$, giving
$\|S(N)\|^2 = O(\varepsilon N^2)$ for all~$\varepsilon > 0$, hence CCSB\@.

\begin{remark}[TWD vs.\ ensemble decorrelation]
\label{rem:twd-vs-ensemble}
TWD is a \emph{temporal} decorrelation statement about the standard EM
orbit: windows far apart in the sequence should have nearly
independent character sums.  The physical intuition is that the
accumulator $\Prod{2}(jK)$ grows super-exponentially, so the starting
points of consecutive windows are extremely different in magnitude,
making it hard for any fixed modulus~$q$ to see systematic correlation
between them.

This contrasts with the \emph{fiber} decorrelation used in the
variance route (\S\ref{sec:variance-route}): there, one averages over
squarefree starting points at two fixed time steps $j, k$.  TWD
averages over time windows at the single starting point $n = 2$.  The
two approaches face different barriers: the variance route faces the
orbit-specificity barrier (density-one results cannot pin down $n=2$),
while TWD faces the temporal-correlation barrier (consecutive windows
share the accumulator prefix $\Prod{2}(jK)$).

The file \code{TailWindowDecorrelation.lean} (477~lines, zero
\code{sorry}) formalizes the window definitions, the tail identity
connection, the variance identity, the diagonal and cross-term bounds,
and the conditional chain TWD~$\Rightarrow$~CCSB~$\Rightarrow$~MC\@.
\end{remark}

\subsection{The Reduction Landscape}
\label{sec:landscape}

Figure~\ref{fig:reduction-graph} shows the complete dependency graph
of the formalization.  Every solid arrow is a machine-verified
implication with zero \code{sorry}.  The single dashed arrow marks the
sole remaining open hypothesis: the Deterministic Stability
Lemma~(DSL), which asks whether population equidistribution (a
consequence of standard analytic number theory) transfers to
conditional multiplier equidistribution for the specific orbit
$n = 2$.

\begin{figure}[!t]
\centering
\begin{tikzpicture}[scale=0.88, every node/.append style={transform shape},
  node distance=0.85cm and 1.3cm,
  every node/.style={font=\small},
  proved/.style={-{Stealth[length=5pt]}, thick, color=leanink},
  open/.style={-{Stealth[length=5pt]}, thick, dashed, red!70!black},
  equiv/.style={{Stealth[length=5pt]}-{Stealth[length=5pt]}, thick, color=leanink},
  box/.style={draw, rounded corners=3pt, inner sep=3pt, minimum height=1.5em,
              font=\small},
  openbox/.style={box, draw=red!70!black, fill=red!5},
  provedbox/.style={box, draw=leanink, fill=blue!5},
  goalbox/.style={box, draw=black, fill=green!8, line width=1pt},
]
% --- Row 1: ANT (standard analytic number theory) ---
\node[provedbox] (ant) {ANT\rlap{\;\scriptsize(MFRE + SRE)}};

% --- Row 2: PE and AccumEquidist (parallel outputs of ANT) ---
\node[below left=0.85cm and 0.6cm of ant, provedbox] (pe) {PE};
\node[below right=0.85cm and 0.6cm of ant, provedbox] (ae) {AccumEquidist};

% --- Row 3: EME ---open--> DSL ---
\node[below=of ae, provedbox] (eme) {EME};
\node[right=2.0cm of eme, openbox] (dsl) {\textbf{DSL}};

% --- Row 5: SD, CME, SVE <-> VE, HOD, TWD ---
\node[below=of dsl, provedbox] (cme) {CME};
\node[left=1.6cm of cme, provedbox] (sd) {SD};
\node[right=1.4cm of cme, provedbox] (sve) {SVE};
\node[right=0.8cm of sve, provedbox] (ve) {VE};
\node[left=1.2cm of sd, provedbox] (hod) {HOD};
\node[right=0.8cm of ve, openbox] (twd) {TWD};

% --- Row 6: CSVB, CCSB ---
\node[below=of cme, provedbox] (ccsb) {CCSB};
\node[below=of sd, provedbox] (csvb) {CSVB};

% --- Row 7: SE, SHH, Hits, DH, Weyl ---
\node[below=1.0cm of ccsb, provedbox] (hits) {Hits $-1$};
\node[right=1.2cm of hits, provedbox] (dh) {DH};
\node[right=1.0cm of dh, provedbox] (wb) {Weyl};
\node[left=1.2cm of hits, provedbox] (shh) {SHH};
\node[left=1.0cm of shh, provedbox] (se) {SE};
\node[left=1.0cm of se, provedbox] (pre) {PRE};

% --- Row 8: MC ---
\node[below=1.0cm of hits, goalbox] (mc) {\textbf{MC}};

% --- Arrows ---
% ANT splits into PE and AccumEquidist
\draw[proved] (ant) -- (pe) node[midway, left, font=\tiny] {MFRE};
\draw[proved] (ant) -- (ae) node[midway, right, font=\tiny] {SRE+CRT};
\draw[proved] (ae) -- (eme);

% PE feeds into DSL (PE + DSL → CME)
\draw[proved] (pe) |- (dsl.west);
\draw[open] (eme) -- (dsl) node[midway, above, font=\tiny] {open};

% DSL -> CME -> CCSB -> Hits -> MC
\draw[proved] (dsl) -- (cme);
\draw[proved] (cme) -- (ccsb);
\draw[proved] (ccsb) -- (hits) node[midway, right, font=\tiny] {Fourier};
\draw[proved] (hits) -- (mc);

% Spectral equivalence: CCSB -> SVE <-> VE
\draw[proved] (ccsb.east) -- ++(0.3,0) |- (sve);
\draw[equiv] (sve) -- (ve);

% Decorrelation chain: SD -> CSVB -a.a.-> CCSB
\draw[proved] (sd) -- (csvb);
\draw[proved] (csvb) -- (ccsb) node[midway, below, font=\tiny] {a.a.};

% Weyl bridge: SVE -> Weyl -> DH -> MC
\draw[proved] (sve) |- (wb);
\draw[proved] (wb) -- (dh);
\draw[proved] (dh) -- (mc);

% Bootstrap: PRE -> SE -> SHH -> MC
\draw[proved] (pre) -- (se) node[midway, above, font=\tiny] {$+\MC({<}\,p)$};
\draw[proved] (se) -- (shh) node[midway, above, font=\tiny] {free};
\draw[proved] (shh) -- (mc);

% HOD -> CCSB (proved)
\draw[proved] (hod) |- (ccsb.west);

% TWD -> CCSB (open bridge)
\draw[open] (twd) |- (ccsb.east) node[near start, right, font=\tiny] {open};

\end{tikzpicture}
\caption{Dependency graph of the main reductions.
  Solid {\color{leanink}blue} arrows: machine-verified.
  Dashed {\color{red!70!black}red} arrows: open hypotheses
  (DSL, TWD~$\Rightarrow$~CCSB).
  All abbreviations are defined in the glossary
  (Appendix~\ref{sec:glossary}).}
\label{fig:reduction-graph}
\end{figure}

The graph reveals the architecture: multiple independent paths converge
on MC\@.
The \emph{left path} (PRE~$\to$~SE~$\to$~SHH~$\to$~MC) is the
inductive bootstrap of Section~\ref{sec:bootstrap}, proved
unconditionally---it reduces MC to producing one hit on~$-1$ at each
prime.  The \emph{central path}
(ANT~$\to$~PE~$\to$~DSL~$\to$~CME~$\to$~CCSB~$\to$~hits~$\to$~MC)
is the ensemble route of this section---it produces the required hit
via harmonic analysis, conditional on DSL\@.  The \emph{right path}
(SVE~$\leftrightarrow$~VE, Weyl~$\to$~DH~$\to$~MC) provides an
alternative spectral route.  Two additional routes reach CCSB:
HOD~$\to$~CCSB (proved, via van der Corput) and
TWD~$\to$~CCSB (open bridge, via temporal window decorrelation).
All solid arrows are machine-verified; the two dashed arrows
(DSL, TWD~$\Rightarrow$~CCSB) are the open hypotheses.

\begin{theorem}[\lean{EM/EnsemblePT.lean}{dsl\_closes\_all}]
Under standard ANT (MinFacResidueEquidist) plus DSL:
$\mathrm{MC} \;\wedge\; \mathrm{CCSB}$.
\end{theorem}
