%% =========================================================================
\section{Variants of the Conjecture}
\label{sec:variants}
%% =========================================================================

The Euclid--Mullin construction makes sense with any factor-selection
rule: at each step, $\Prod{2}(n)+1$ has a non-empty set of prime
factors, and the rule picks one.  Mullin's Conjecture is specific to
the $\minFac$ rule.  This section collects the known variants---some
easy, some provably incomplete, one almost closed---and uses them to
locate standard MC within a broader landscape.

\begin{roadmap}
\begin{enumerate}[nosep,leftmargin=*]
\item[\S\ref{sec:random-variant}] Random-factor variant --- MC by mixing; independence vs.\ selection trade-off.
\item[\S\ref{sec:factor-tree}] Factor tree and GenMixedMC --- nondeterministic access to all prime factors.
\item[\S\ref{sec:almost-all-mixed}] Almost all GenMixedMC --- density-$1$ result: PEAP alone suffices (no open hypothesis).
\item[\S\ref{sec:minfac-secondminfac}] $\{\minFac,\operatorname{secondMinFac}\}$ variant --- two-point spectral contraction, UFDStrong.
\item[\S\ref{sec:ensemble-variant}] Ensemble MC --- almost every starting point, backward dynamics.
\item[\S\ref{sec:weak-mullin}] Weak Mullin --- analytic weakenings, partial results.
\item[\S\ref{sec:function-field}] Function field analog --- $\FF_p[t]$, Weil bound, orbit-specificity barrier.
\end{enumerate}
\end{roadmap}

\paragraph{The Selection Rule Landscape}
\label{sec:selection-landscape}

Fix an initial prime $p_0 = 2$.  Given a \emph{selection rule}
$\sigma$ that, for any integer $N \geq 2$, chooses a prime factor
$\sigma(N) \mid N$, define:
\begin{align}
  P_\sigma(0) &= 2, &
  P_\sigma(n{+}1) &= P_\sigma(n) \cdot \sigma(P_\sigma(n)+1).
\end{align}
The selection rule determines the sequence; the running product
$P_\sigma(n)$ determines the sequence history.
Five rules are studied in the formalization:

\smallskip
\begin{center}
\begin{tabular}{@{}l@{\quad}l@{\quad}l@{}}
\toprule
\textbf{Selection rule} $\sigma$ & \textbf{MC status} & \textbf{Key property} \\
\midrule
Random (uniform) & Easy (mixing) & Independence \\
$\minFac$ (standard EM) & \textbf{Open} (DSL) & Deterministic, lethal \\
$\{\minFac, \operatorname{secondMinFac}\}$ & \textbf{Open} (UFDStrong) & Two-point diversity \\
$(1{-}\varepsilon)\,\minFac + \varepsilon\,$random & a.a.\ (PEAP) & Full factor access \\
Steered (Booker~\cite{BoSa2012}) & True (by construction) & Adversarial \\
$\mathrm{maxFac}$ (Cox--vdP~\cite{CoxVdP1968}) & False & Misses primes \\
\bottomrule
\end{tabular}
\end{center}

\smallskip\noindent
The landscape reveals a tension: the random rule is the easiest to
analyze (mixing theory) but the worst at selection (probability
$1/|F|$ per hit); $\minFac$ is the best at selection (lethal hits)
but the hardest to analyze (deterministic, no independence); the
$\{\minFac, \operatorname{secondMinFac}\}$ variant sits between them,
offering a two-point spectral contraction that is provably sufficient
under a single open hypothesis; and the factor tree framework
(\S\ref{sec:factor-tree}) generalizes all three by allowing access
to the full factor set at every step, reducing GenMixedMC to sieve
theory for almost all starting points.

\subsection{The Random-Factor Variant}
\label{sec:random-variant}

Consider a variant where, instead of the smallest prime factor, one
picks a \emph{random} prime factor of~$\Prod{2}(n)+1$ at each step.
MC follows from classical mixing theory.  For any target prime~$p$,
the residue $r_n = \Prod{2}(n) \bmod p$ performs a multiplicative walk
on $(\ZZ/p\ZZ)^\times$.  Each multiplier is a random element of the
group; once the multipliers generate the full group (which PRE
guarantees, Theorem~\ref{thm:pre}), the walk is a genuine random walk
with full support.  By Diaconis--Shahshahani equidistribution, the walk
mixes in $O(\log p)$ steps: $r_n = -1$ occurs with frequency
$\to 1/(p{-}1)$---infinitely often with probability~1.  When $p$
divides the Euclid number, the random rule selects~$p$ with probability
$1/|F(S_n)|$, but cofinal hitting provides infinitely many chances,
so~$p$ is eventually selected.

\paragraph{Comparison with $\minFac$.}
The actual EM sequence uses
$\seq{2}(n{+}1) = \minFac(\Prod{2}(n){+}1)$, a \emph{deterministic}
function of the walk position.  Compared to the random-factor variant,
$\minFac$ is strictly better in every measurable dimension:

\smallskip
\begin{center}
\begin{tabular}{l@{\quad}l@{\quad}l}
\toprule
 & \textbf{Random factor} & $\boldsymbol{\minFac}$ \\
\midrule
\textbf{Selection when $q \mid P(n){+}1$}
  & Probability $1/|F(S_n)|$ & Probability~1 \\
\textbf{Accumulator growth}
  & Faster (random $\geq \minFac$) & Slowest possible \\
\textbf{Hits on $-1$ needed}
  & Infinitely many & \textbf{One} (lethal hit) \\
\bottomrule
\end{tabular}
\end{center}

\smallskip\noindent
\emph{Selection efficiency.}\enspace Past the sieve gap, 0-accessibility
guarantees: if $q \mid \Prod{2}(n)+1$ then
$q = \minFac(\Prod{2}(n)+1)$---selection is certain, not
probabilistic.  \emph{Growth rate.}\enspace Since
$\minFac(N) \leq p$ for any factor~$p$ of~$N$, the accumulator
$\Prod{2}(n)$ grows minimally under the $\minFac$ rule, giving
exponentially more steps before it becomes astronomically large---more
chances for~$q$ to divide.  \emph{Lethality.}\enspace The inductive
bootstrap shows one single hit suffices: $\MC({<}\,q)$ ensures~$q$
is minimal whenever it divides, so hitting $-1$ once means~$q$
enters the sequence.  The random variant requires infinitely many
hits to overcome low per-hit selection probability.

\emph{MinFac is strictly better in every measurable dimension.
So why can't we prove it works?}

\paragraph{The independence trade-off.}
The answer is surgical: the random variant possesses a single property
that $\minFac$ lacks, and that property is the \emph{only thing the
proof uses}.

The property is \textbf{independence of multiplier from walk
position}.  In the random-factor variant, the multiplier at step~$n$
is selected independently of which residue class
$\Prod{2}(n) \bmod q$ occupies.  Walk position and multiplier are
independent random variables, and independence is the engine that
drives mixing.  Diaconis--Shahshahani's theorem applies whenever the
multipliers are i.i.d.\ from a distribution whose support generates
the group.

In the $\minFac$ variant, the multiplier is
$\minFac(\Prod{2}(n){+}1) \bmod q$---a deterministic function of the
full integer~$\Prod{2}(n)+1$.  The CRT multiplier invariance
(\lean{EM/Group/CRT.lean}{crt\_multiplier\_invariance}) shows
that the outcome is independent of $\Prod{2}(n) \bmod q$ at the
\emph{population} level: for generic integers, knowing the mod-$q$
residue does not predict $\minFac$.  But for the specific orbit~$n=2$,
the multiplier is a deterministic value, and mixing theorems for
deterministic sequences cannot be invoked.

The random variant thus sacrifices selection efficiency (it may
miss~$q$ even when $q$ divides) in exchange for statistical
independence.  The $\minFac$ variant has perfect selection but no
independence:

\smallskip
\begin{center}
\begin{tabular}{l@{\quad}l@{\quad}l}
\toprule
 & \textbf{Random factor} & $\boldsymbol{\minFac}$ \\
\midrule
Selection & Bad ($1/k$ chance) & Perfect \\
Mixing & Random walk theorem & Unproved \\
\bottomrule
\end{tabular}
\end{center}

\smallskip\noindent
The random variant's proof: mixing $\to$ cofinal hitting $\to$
infinitely many chances compensate for low per-hit probability.
The $\minFac$ variant needs only one hit---absurdly less than the
random variant achieves---but has no way to produce even that one
without mixing.

\paragraph{The stochastic domination failure.}
\label{sec:stoch-dom}
The natural instinct is to argue by \emph{stochastic domination}:
if the random walk visits~$-1$ cofinally, and $\minFac$ is ``better
in every way,'' then $\minFac$ should also visit~$-1$.  This argument
fails because the two processes \emph{diverge after one step}.

At step~$n$, both processes face the same Euclid number
$\Prod{2}(n)+1$ with factor set~$F(S_n)$.  The random variant picks a
uniformly random element of~$F(S_n)$; the $\minFac$ variant picks the
minimum.  After this step, the two processes have different
accumulators (they multiplied by different primes), so they face
different Euclid numbers at the next step.  The processes explore
completely different trajectories.  Domination fails because the
two walks live on different probability spaces after step~1; we
cannot compare them step-by-step.

A subtler comparison uses \emph{visit measures}.  Define
$\mu_{\mathrm{rand}}(a) = \lim_{N} \frac{1}{N}|\{n < N :
w_q^{\mathrm{rand}}(n) = a\}|$ and analogously
$\mu_{\min}(a)$ for the $\minFac$ walk.  For the random variant,
$\mu_{\mathrm{rand}} = $ uniform on $(\ZZ/q\ZZ)^\times$ by mixing.
For $\minFac$, CME/CCSB asserts $\mu_{\min} \approx$ uniform, and
DSL asserts this follows from PE\@.  The comparison
$\mu_{\min}(-1) \geq 1/(q{-}1)$ is precisely what needs proof, and is
the content of DSL\@.

%% =========================================================================
\subsection{The Factor Tree and GenMixedMC}
\label{sec:factor-tree}
%% =========================================================================

The random variant's provability motivates a broader framework:
instead of fixing a single selection rule, consider the
\textbf{factor tree} $\mathcal{T}(m)$ rooted at starting point~$m$.
The root holds accumulator~$m$; each node with accumulator~$P$ has
one child $P \cdot p$ for every prime $p \mid P{+}1$.
The leftmost path (always choosing $\minFac$) recovers the standard EM
walk; every other path corresponds to a different selection.
All nodes of $\mathcal{T}(m)$ share the same structural properties:
each factor is prime, divides $P{+}1$, and is coprime to~$P$; the
accumulator is strictly increasing along every branch.

\begin{definition}[GenMixedMC]
\label{def:gen-mixed-mc}
$\mathrm{GenMixedMC}(m)$: for every prime~$q$, some node of
$\mathcal{T}(m)$ has $q$ dividing its Euclid number $P{+}1$.
Equivalently, every prime appears somewhere in the factor tree.
\end{definition}

\noindent
The standard Mullin Conjecture is the special case where the
leftmost path alone captures every prime; GenMixedMC relaxes this to
any path.  Formally, $\mathrm{MC} \Longrightarrow
\mathrm{GenMixedMC}(2)$, since the standard walk is one particular
path through $\mathcal{T}(2)$.

\begin{remark}[The Euclid--Mullin graph]
\label{rem:em-graph}
The factor tree $\mathcal{T}(m)$ is exactly the \emph{Euclid--Mullin
graph}~$G_m$ of Booker and Irvine~\cite{BoIr2016}.  In their
framework, vertices are accumulated products and edges are labelled by
primes dividing $P{+}1$; GenMixedMC$(m)$ asks whether every prime
labels some edge.  Computationally, every prime below~$10^9$ occurs as
an edge in~$G_1$ (hence in~$G_2$).

Booker~\cite{BoSa2012} proved that for any finite set~$P$ of primes,
there exists a path through $\mathcal{T}(\prod P)$ containing every
prime.  His proof uses three ingredients that connect directly to the
formalization's infrastructure:
\begin{enumerate}[nosep]
\item \emph{Squarefree smooth density}:
  the set~$S_q$ of residues mod~$q$ attained by squarefree
  $(q{-}1)$-smooth integers satisfies $|S_q| > (q{-}1)/2$
  (via Schnirelmann density $\geq 53/88$)---the same population
  equidistribution that DSL seeks to transfer to a single orbit.
\item \emph{Weil--Hasse bounds} for the curves $y^2 = x(x^2+a)$
  and $y^2 = x^6+a$---the same character-sum cancellation exploited
  by the function field analog (\S\ref{sec:function-field}).
\item \emph{Subgroup growth}: the subgroup of $(\ZZ/q\ZZ)^\times$
  generated by the reachable primes has index dividing~$6$---an
  analogue of SubgroupEscape (Theorem~\ref{thm:pre}), which guarantees
  full generation for the standard walk.
\end{enumerate}
Booker's proof works because choosing subsets at each step provides
enough freedom to steer residues mod~$q$.  The standard EM sequence
($\minFac$, always using the full product) has no such freedom: the
accumulator is deterministic, and the gap between population
equidistribution (Booker's Lemma~2) and orbit-specific
equidistribution (CME) is exactly what DSL addresses.
\end{remark}

\begin{remark}[Quantifier structure]
\label{rem:quantifier}
GenMixedMC$(m)$ asserts: for each prime~$q$, \emph{some} path
through $\mathcal{T}(m)$ reaches a node where $q$ divides the Euclid
number.  Different primes may require different paths.  The quantifier
order is $\forall q\;\exists\,\text{path}$, \textbf{not}
$\exists\,\text{path}\;\forall q$.  The stronger statement---a single
path capturing all primes simultaneously---remains open even for
almost all starting points.
\end{remark}

\paragraph{Lean encoding: MixedSelection.}
The formalization encodes paths through $\mathcal{T}(m)$ via a
\texttt{MixedSelection}
$\sigma : \NN \to \mathrm{Option}\;\NN$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{MixedSelection}):
$\mathrm{none}$ at step~$k$ means follow the leftmost child
($\minFac$); $\mathrm{some}\;p$ means branch to child~$p$.
A selection is \defname{valid}
(\lean{EM/Advanced/EpsilonRandomMC.lean}{ValidMixedSelection})
if every $\mathrm{some}\;p$ step has $p$ prime and $p \mid P_\sigma(k){+}1$.
The accumulator along a path is
$P_\sigma(0) = m$,
$P_\sigma(n{+}1) = P_\sigma(n) \cdot f_\sigma(n)$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkProd}).

\paragraph{The reachable set: tree projection.}
Define $R_n(q, m)$ as the set of positions
$P_\sigma(n) \bmod q$ over all valid paths~$\sigma$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{reachableAt})---the
projection of depth-$n$ nodes of $\mathcal{T}(m)$ onto
$(\ZZ/q\ZZ)^\times$.  The cumulative set
$R_\infty = \bigcup_n R_n$ collects all residues attained by
any node.  Hitting is equivalent to reachability:
\[
  q \text{ appears in } \mathcal{T}(m)
  \;\Longleftrightarrow\;
  {-}1 \in R_\infty(q, m)
\]
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixed\_hitting\_iff\_neg\_one\_reachable}
{\color{leanink}$\checkmark$}).
This reformulates GenMixedMC as a BFS reachability problem
in $(\ZZ/q\ZZ)^\times$.
The reachable set is monotone along walk prefixes:
any position reachable from a descendant $P_\sigma(n)$ is also
reachable from the original accumulator~$m$
(\lean{EM/Advanced/InterpolationMC.lean}{reachableEver\_mono\_along\_walk}
{\color{leanink}$\checkmark$}),
via concatenation of walk segments.

\paragraph{Reachable set growth.}
The core growth mechanism: at any node with accumulator~$P$,
\emph{every} prime~$p$ dividing $P{+}1$ produces a child at position
$P \cdot p \bmod q$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{reachableAt\_from\_factor}
{\color{leanink}$\checkmark$}).
When $P{+}1$ is composite, it has $\geq 2$ prime divisors,
giving $\geq 2$ children
(\lean{EM/Advanced/EpsilonRandomMC.lean}{reachable\_composite\_branch}
{\color{leanink}$\checkmark$}).
When these are distinct mod~$q$, $|R_{n+1}| > |R_n|$ along this branch.

\paragraph{The purification lemma.}
The probabilistic argument for the random variant requires
Diaconis--Shahshahani mixing theory and Borel--Cantelli, neither
available in Lean.  The formalization instead proves that the
nondeterministic framework captures everything:

\smallskip\noindent
\textbf{Purification Lemma}
(\lean{EM/Advanced/RandomFactorMC.lean}{pure\_random\_mc\_iff\_mixed\_mc}
{\color{leanink}$\checkmark$}).
\emph{PureRandomMC$(q)$ $\Leftrightarrow$ MixedMC$(q)$.}

\smallskip\noindent
The proof constructs, from any path through $\mathcal{T}(m)$ (which
may follow the leftmost child at some steps), a
\emph{purified} path that always branches explicitly and follows the
exact same trajectory
(\lean{EM/Advanced/RandomFactorMC.lean}{purify\_walk\_eq}
{\color{leanink}$\checkmark$}).
This works because $\minFac(P{+}1)$ is itself a prime dividing~$P{+}1$,
so replacing every ``follow leftmost'' with ``branch to
$\minFac(P{+}1)$'' produces a valid explicit path.
The distinction between ``$\minFac$-first'' and
``all-random'' is thus illusory at the nondeterministic level.
Standard~MC implies GenMixedMC$(2)$ (the leftmost path is one
particular branch,
\lean{EM/Advanced/RandomFactorMC.lean}{standard\_mc\_implies\_pure\_random}
{\color{leanink}$\checkmark$}).

\paragraph{Variant hierarchy.}
The factor tree framework unifies the prior variants in a strict
hierarchy:
\[
  \text{deterministic } (\text{leftmost path})
  \;\subset\;
  \text{two-point } (\S\ref{sec:minfac-secondminfac})
  \;\subset\;
  \mathcal{T}(m)
  \;\subset\;
  \text{full random.}
\]
The two-point walk (\S\ref{sec:minfac-secondminfac}) restricts
branching to the two smallest factors; the full random walk
corresponds to sampling a uniformly random branch at each node.
Any capture by the two-point walk gives a capture in $\mathcal{T}(m)$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{two\_point\_capture\_implies\_mixed\_capture}
{\color{leanink}$\checkmark$}).

\paragraph{Basic properties.}
Under validity and $m \geq 2$, the formalization proves:
every factor is prime
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkFactor\_prime}
{\color{leanink}$\checkmark$});
every factor divides $P{+}1$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkFactor\_dvd}
{\color{leanink}$\checkmark$});
the accumulator stays $\geq 2$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkProd\_ge\_two}
{\color{leanink}$\checkmark$});
the accumulator is strictly increasing
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkProd\_strict\_mono}
{\color{leanink}$\checkmark$});
each factor is coprime to the accumulator
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkFactor\_not\_dvd\_prod}
{\color{leanink}$\checkmark$});
prefix dependence: $P_\sigma(k)$ depends only on $\sigma(0),\ldots,\sigma(k{-}1)$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkProd\_depends\_on\_prefix}
{\color{leanink}$\checkmark$});
tail restart: the subtree from step~$K$ is a fresh tree from~$P_\sigma(K)$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkProd\_tail\_restart}
{\color{leanink}$\checkmark$}).

\begin{theorem}[Mixed walk structure]
\label{thm:mixed-walk-structure}
For any valid selection~$\sigma$ from a starting point $m \geq 2$:
\begin{enumerate}[nosep]
\item \textbf{Exponential growth}
  (\lean{EM/Advanced/InterpolationMC.lean}{mixedWalkProd\_exp\_growth}
  {\color{leanink}$\checkmark$}):
  $P_\sigma(k) \geq m \cdot 2^k$ for all~$k$.
\item \textbf{Squarefree preservation}
  (\lean{EM/Advanced/InterpolationMC.lean}{mixedWalkProd\_squarefree}
  {\color{leanink}$\checkmark$}):
  if $m$ is squarefree, so is $P_\sigma(k)$ for all~$k$.
\item \textbf{Coprimality from no death}
  (\lean{EM/Advanced/InterpolationMC.lean}{mixedWalkProd\_coprime\_of\_no\_death}
  {\color{leanink}$\checkmark$}):
  if the walk never hits $-1 \bmod q$, then $\gcd(P_\sigma(k), q) = 1$ for all~$k$.
\item \textbf{Strict monotonicity}
  (\lean{EM/Advanced/EpsilonRandomMC.lean}{mixedWalkProd\_strict\_mono}
  {\color{leanink}$\checkmark$}):
  $P_\sigma$ is strictly increasing.
\end{enumerate}
\end{theorem}

\paragraph{Unconditional capture of $q = 3$.}
From $m = 2$, the root has $P{+}1 = 3$.  Since $3$ is prime, every
branch at depth~1 selects factor~$3$---there is no other choice.
Thus $3$ appears in $\mathcal{T}(2)$ unconditionally
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixed\_capture\_three}
{\color{leanink}$\checkmark$}).

\paragraph{The diversity hypothesis.}
The tree has branching $> 1$ at a node when $P{+}1$ is composite.
\defname{MixedDiversity}
(\lean{EM/Advanced/EpsilonRandomMC.lean}{MixedDiversity}):
along the leftmost path, $P(n){+}1$ is composite cofinally often.
Under MixedDiversity, the tree has genuine branching at infinitely
many levels, and PRE (Theorem~\ref{thm:pre}) guarantees the factors
generate the full unit group $(\ZZ/q\ZZ)^\times$.
Since $P(n)$ grows super-exponentially, $P(n){+}1$ is composite for
all known $n \geq 1$, so MixedDiversity is empirically unchallenged.
The conditional reduction gives:

\begin{theorem}[Diversity implies MixedMC ---
\lean{EM/Advanced/EpsilonRandomMC.lean}{mixed\_diversity\_weak\_implies\_mixed\_mc}
{\color{leanink}$\checkmark$}]
\label{thm:diversity-mixed-mc}
$\mathrm{MixedDiversityWeak} + \mathrm{MixedDiversity}
\;\Longrightarrow\; \mathrm{MixedMullinConjecture}$.
The proof is by strong induction on~$q$: MixedDiversityWeak provides
branching at each level, MixedDiversity ensures enough branching
to generate the full unit group, and PRE converts generation to hitting.
\end{theorem}

\paragraph{The perpetual primality obstruction.}
If $P(n){+}1$ were prime for all $n \geq N$, the recurrence
would collapse to $P(n{+}1)+1 = \Phi_3(P(n))$ (the third cyclotomic
polynomial,
\lean{EM/Advanced/EpsilonRandomMC.lean}{perpetual\_prime\_cyclotomic}
{\color{leanink}$\checkmark$}),
and $P(n) \equiv 1 \pmod{3}$ would force $3 \mid \Phi_3(P(n))$,
making $P(n){+}1$ composite---a contradiction.  More precisely,
perpetual primality confines the walk to $\{0,2\} \bmod 3$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{perpetual\_prime\_excludes\_mod3\_one}
{\color{leanink}$\checkmark$}).
More generally, under perpetual primality the walk mod~$q$ follows
the autonomous map $w \mapsto w(w{+}1)$ on $\ZZ/q\ZZ$
(\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_autonomous\_mod}
{\color{leanink}$\checkmark$}),
and by pigeonhole is eventually periodic
(\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_eventually\_periodic}
{\color{leanink}$\checkmark$}).
For $q = 5$, if the walk enters~$\{1,2\}$, the autonomous orbit
alternates between $1$ and $2$ forever (since $1 \cdot 2 = 2$ and
$2 \cdot 3 = 1$ in~$\ZZ/5\ZZ$), never reaching $-1 = 4$
(\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_mod5\_orbit}
{\color{leanink}$\checkmark$}).
These results combine into a clean structural statement:

\begin{theorem}[Perpetual primality periodicity]
\label{thm:perpetual-periodicity}
Suppose $P(n)+1$ is prime for all $n \geq N$.  Then for every prime~$q$:
\begin{enumerate}[nosep]
\item The walk mod~$q$ satisfies the autonomous recurrence
  $w(n{+}1) = w(n) \cdot (w(n){+}1)$
  (\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_autonomous\_mod}
  {\color{leanink}$\checkmark$}).
\item The walk is eventually periodic mod~$q$ with period $T \leq q$
  (\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_eventually\_periodic}
  {\color{leanink}$\checkmark$}).
\item For $q = 3$: the walk avoids $1 \bmod 3$, confined to $\{0, 2\}$
  (\lean{EM/Advanced/EpsilonRandomMC.lean}{perpetual\_prime\_excludes\_mod3\_one}
  {\color{leanink}$\checkmark$}).
\item For $q = 5$: the orbit from $\{1, 2\}$ is a $2$-cycle, never reaching
  $-1 = 4$
  (\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_mod5\_orbit}
  {\color{leanink}$\checkmark$}).
\item For $q = 11$: the orbit from $\{2, 6, 9\}$ is a $3$-cycle, never
  reaching $-1 = 10$
  (\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_mod11\_orbit}
  {\color{leanink}$\checkmark$}).
\item For $q = 17$: the orbit from $\{2, 3, 4, 6, 8, 12\}$ is a closed
  set (tail~$+$ $2$-cycle $\{3, 12\}$), never reaching $-1 = 16$
  (\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_mod17\_orbit}
  {\color{leanink}$\checkmark$}).
\item For $q = 23$: the orbit from $\{2, 6, 12, 18, 19, 20\}$ is a closed
  set (tail~$+$ $5$-cycle), never reaching $-1 = 22$
  (\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_prime\_mod23\_orbit}
  {\color{leanink}$\checkmark$}).
\end{enumerate}
The four-prime orbit exclusion
(\lean{EM/Advanced/InterpolationMC.lean}{perpetual\_primality\_multi\_exclusion}
{\color{leanink}$\checkmark$})
shows that under perpetual primality, primes $5, 11, 17, 23$ are
simultaneously blocked from entering the sequence.
\end{theorem}

%% =========================================================================
\paragraph{Factor Confinement}
\label{sec:factor-confinement}
%% =========================================================================

The reachable set $R_\infty(q, m) = \bigcup_n R_n(q, m)$ cannot be
confined to any proper coset $g \cdot H$ of $(\ZZ/q\ZZ)^\times$:
if $R_\infty \subseteq g \cdot H$, then all multiplier residues lie
in~$H$, contradicting PrimeResidueEscape
(\lean{EM/Equidist/Bootstrap.lean}{prime\_residue\_escape}
{\color{leanink}$\checkmark$}).

\begin{theorem}[\lean{EM/Advanced/EpsilonRandomMC.lean}{reachableEver\_not\_in\_coset}
{\color{leanink}$\checkmark$}]
\label{thm:coset-impossibility}
For $q \geq 3$ prime and any proper subgroup $H < (\ZZ/q\ZZ)^\times$,
$R_\infty(q, 2) \not\subseteq g \cdot H$ for every~$g$.
\end{theorem}

\noindent
Coset impossibility constrains the geometry of $R_\infty$ but does
not force $R_\infty = (\ZZ/q\ZZ)^\times$.  A complementary approach
attacks the \emph{factor} side: if $R_\infty \subsetneq
(\ZZ/q\ZZ)^\times$, then the prime factors of every Euclid number
$P(n){+}1$ along \emph{every} path are confined to a ``forbidden''
set determined by~$R_\infty$
(\lean{EM/Advanced/EpsilonRandomMC.lean}{factor\_confinement}
{\color{leanink}$\checkmark$}).

\begin{property}[FactorEscapeHypothesis ---
\lean{EM/Advanced/EpsilonRandomMC.lean}{FactorEscapeHypothesis}]
\label{prop:factor-escape}
For every prime $q \geq 3$, the Euclid numbers of $\mathcal{T}(2)$
eventually have a prime factor outside the forbidden set induced by any
proper $R \subsetneq (\ZZ/q\ZZ)^\times$.
\end{property}

\begin{theorem}[\lean{EM/Advanced/EpsilonRandomMC.lean}{factor\_escape\_implies\_mixed\_mc\_at}
{\color{leanink}$\checkmark$}]
$\mathrm{FactorEscapeHypothesis}(q) \;\Longrightarrow\;
R_\infty(q, 2) = (\ZZ/q\ZZ)^\times
\;\Longrightarrow\;$ every prime~$q$ appears in $\mathcal{T}(2)$.
\end{theorem}

\noindent
The comparison between the standard walk and the factor tree is stark:
\smallskip
\begin{center}
\begin{tabular}{@{}l@{\quad}c@{\quad}c@{}}
\toprule
 & \textbf{Standard walk} & \textbf{Factor tree} $\mathcal{T}(m)$ \\
\midrule
Factors used per step & $1$ ($\minFac$) & All $\omega(P{+}1)$ \\
Confinement forces & $\minFac \in F$ & Every factor $\in F$ \\
Population analogy & DSL (orbit transfer) & PSCD (sieve theory) \\
\bottomrule
\end{tabular}
\end{center}

\smallskip\noindent
For the standard walk, avoiding~$-1$ confines only one factor per
step; for the tree, it confines \emph{all} $\omega(P{+}1)$ factors
simultaneously---a much stronger sieve constraint that can be
attacked with population methods.

%% =========================================================================
\subsection{Almost All GenMixedMC: A Density-1 Result}
\label{sec:almost-all-mixed}
%% =========================================================================

The factor confinement of \S\ref{sec:factor-confinement} connects
the hitting question to sieve theory over the population of starting
points.  For a squarefree $m$ coprime to~$q$, if
$-1 \notin R_\infty(q,m)$, then $R_\infty$ misses a
\emph{nonzero} element of $\ZZ/q\ZZ$.  Moreover, the walk stays in
units (each multiplier is prime and $\neq q$, since $q \mid P{+}1$
would place~$-1$ in~$R_\infty$), so
$0 \notin R_\infty$ as well
(\lean{EM/Ensemble/MixedEnsemble.lean}{zero\_not\_reachable\_of\_coprime\_trapped}
{\color{leanink}$\checkmark$}).
At the root, all prime factors of $m{+}1$ have residues in the
allowed factor set determined by~$R_\infty$
(\lean{EM/Ensemble/MixedEnsemble.lean}{unconditional\_confinement}
{\color{leanink}$\checkmark$}).
By pigeonhole over the finitely many proper finsets~$R$, the
``trapped'' count is bounded by the sum of confined counts
(\lean{EM/Ensemble/MixedEnsemble.lean}{trapped\_le\_sum\_confined}
{\color{leanink}$\checkmark$}).

The open hypothesis \defname{PSCD}
(\emph{Population Sieve Confinement Decay}) asserts that among
squarefree integers, the fraction confined to any finset~$R$ missing
a nonzero element of $\ZZ/q\ZZ$ tends to zero.
The restriction to $R$ missing a
\emph{nonzero} element is essential---$R =
(\ZZ/q\ZZ)^\times$ is proper as a subset of $\ZZ/q\ZZ$
but confinement to it merely means $q \nmid m{+}1$,
which has density $(q{-}1)/q > 0$.

\begin{theorem}[\lean{EM/Ensemble/MixedEnsemble.lean}{pscd\_implies\_almost\_all\_mixed\_hitting}
{\color{leanink}$\checkmark$}]
\label{thm:almost-all-mixed}
$\mathrm{PSCD}(q)$ implies: for a density-$1$ set of squarefree~$m$
coprime to~$q$, every prime~$q$ appears in $\mathcal{T}(m)$.
\end{theorem}

\paragraph{Reduction to analytic number theory.}
PSCD reduces to standard analytic number theory via two routes.

\smallskip\noindent
\emph{Route~1 (original, one open bridge).}\enspace
Three intermediate steps:
\begin{enumerate}[nosep]
\item {\color{leanink}$\checkmark$}\enspace
  PrimesEquidistributedInAP $\Rightarrow$ $\sum_{p \equiv a} 1/p$ diverges
  for every coprime class~$a$
  (\lean{EM/Ensemble/MixedEnsemble.lean}{peap\_implies\_fcd\_proved}).
\item \defname{SieveUpperBound}: confined count $\leq$ $C \cdot
  \text{sqfreeCount} \cdot \prod_{p \in F} (1 - 1/p)$
  (fundamental lemma of sieve).
\item {\color{leanink}$\checkmark$}\enspace
  Divergent $\sum 1/p$ implies sieve product $\to 0$
  (\lean{EM/Ensemble/MixedEnsemble.lean}{sieve\_product\_vanishing\_proved}).
\end{enumerate}
SieveUpperBound---the fundamental lemma of sieve theory---is open.

\smallskip\noindent
\emph{Route~2 (WeakFMCD, unconditional).}\enspace
The key insight is that FMCD's exact constant is unnecessary:
the sieve product $\prod_{p \in F}(1-1/p) \to 0$ absorbs any
finite multiplicative constant.  Define \defname{WeakFMCD}
(\lean{EM/Ensemble/MixedEnsemble.lean}{weak\_fmcd\_proved}
{\color{leanink}$\checkmark$}):
for any fixed finite set of primes~$S$,
\[
  \operatorname{sqfreeSievedCount}(S,X) \;\leq\;
  \Bigl(4\prod_{p \in S}(1 - 1/p) + \varepsilon\Bigr)
  \cdot \operatorname{sqfreeCount}(X).
\]
The constant~$4$ comes from the unconditional bound
$\operatorname{sqfreeCount}(X) \geq X/4$
(\lean{EM/Ensemble/MixedEnsemble.lean}{sqfreeCount\_ge\_quarter\_real}
{\color{leanink}$\checkmark$}), proved via $\sum_{d=2}^{D}1/d^2 < 3/4$.
WeakFMCD is then immediate from CRT residue class counting
(each class has $\leq X/N + 2$ elements) and Euler's product
$\varphi(\prod S)/\prod S = \prod(1-1/p)$.

The central theorem
(\lean{EM/Ensemble/MixedEnsemble.lean}{weak\_fmcd\_fcd\_implies\_pscd}
{\color{leanink}$\checkmark$})
proves ForbiddenClassDivergent alone $\Rightarrow$ PSCD\@.  The proof
decomposes the confined count over residue classes mod~$q$ as before,
but with $\varepsilon/(8q)$ for the sieve target and $\varepsilon/(2q)$
for the WeakFMCD tolerance, so $4 \cdot \varepsilon/(8q) + \varepsilon/(2q)
= \varepsilon/q$; summing over $q{-}1$ classes gives $< \varepsilon$.
\begin{keyresult}
The composed chain
(\lean{EM/Ensemble/MixedEnsemble.lean}{weak\_fmcd\_chain\_implies\_pscd}
{\color{leanink}$\checkmark$}):
\[
  \text{PrimesEquidistInAP}
  \;\xRightarrow{\text{\color{leanink}$\checkmark$}}\;
  \text{PSCD}(q)
  \;\xRightarrow{\text{\color{leanink}$\checkmark$}}\;
  \text{a.a.\ GenMixedMC}.
\]
Every arrow is machine-checked.  \textbf{No open hypothesis remains}:
PEAP (Wiener--Ikehara) is the sole input, and the CRT bound is
unconditional.
\end{keyresult}

\begin{deadend}[title={Dead End \#117: MultCancelToWalkCancel}]
The standard ensemble approach
(\S\ref{sec:ensemble-variant}) proves almost-all
\emph{multiplier cancellation} for the deterministic walk, but
transferring this to walk \emph{hitting} is blocked by
Dead End~\#117 (MultCancelToWalkCancel).  The factor tree approach
bypasses~\#117 entirely:
\smallskip
\begin{center}
\begin{tabular}{@{}l@{\quad}l@{\quad}l@{}}
\toprule
 & \textbf{Standard ensemble} & \textbf{Factor tree ensemble} \\
\midrule
Population result & a.a.\ char cancel & a.a.\ GenMixedMC \\
Transfer gap & \#117 (cancel $\not\Rightarrow$ hitting) & None \\
Selection used & $\minFac$ only & All factors of $P{+}1$ \\
Open hypothesis & StepDecorrelation & None (PEAP only) \\
\bottomrule
\end{tabular}
\end{center}
\end{deadend}

\smallskip\noindent
This is the strongest density result in the formalization: under
PrimesEquidistributedInAP alone (no additional open hypothesis),
for a density-$1$ set of squarefree
starting points, \emph{every prime appears in the factor tree}.

\paragraph{Interpolation: tree reachability implies positive probability capture.}
The density-$1$ tree result bridges to the $(1{-}\varepsilon)\,\minFac
+ \varepsilon\,$random process.  Assign to each factor choice at
accumulator~$P$ the weight $\varepsilon / \omega(P{+}1)$---a lower
bound on the probability of that choice under the interpolated process
(where $\omega$ counts distinct prime factors).  A finite path
$\sigma$ of length~$n$ has weight
$\prod_{k<n} \varepsilon / \omega(P_\sigma(k){+}1) > 0$
(\lean{EM/Advanced/InterpolationMC.lean}{pathWeightLB\_pos}
{\color{leanink}$\checkmark$}).

If $-1 \in R_\infty(q,m)$ (tree reachability), a capturing path
exists, and that path has positive weight under the
$(1{-}\varepsilon)$ process for \emph{any} $\varepsilon > 0$
(\lean{EM/Advanced/InterpolationMC.lean}{reachable\_implies\_positive\_prob\_capture}
{\color{leanink}$\checkmark$}).
Under PEAP, almost all squarefree~$m$ have $-1$ reachable
(WeakFMCD chain), so almost all have positive probability
capture
(\lean{EM/Advanced/InterpolationMC.lean}{almost\_all\_positive\_prob\_capture}
{\color{leanink}$\checkmark$}).
No additional open hypothesis beyond PEAP is needed.

\paragraph{Layer 2: cofinal and iterated hitting via regeneration.}
Layer~1 proves a \emph{single} hit with positive weight.  Stronger
conclusions require \defname{Regeneration}
(\lean{EM/Advanced/InterpolationMC.lean}{Regeneration}): every
accumulator encountered along any valid walk from~$m$ is \emph{good},
i.e., every unit of~$(\ZZ/q\ZZ)^\times$ remains reachable.  The
evidence is strong: accumulators grow super-exponentially and are
squarefree, so by CRT blindness they should behave generically mod~$q$.

Under Regeneration, the formalization proves:
\begin{itemize}
\item \textbf{Block coverage}
(\lean{EM/Advanced/InterpolationMC.lean}{block\_coverage}
{\color{leanink}$\checkmark$}):
from a good accumulator, there is a uniform depth bound~$N_0$ such
that every unit can be reached within~$N_0$ steps.
\item \textbf{Cofinal hitting}
(\lean{EM/Advanced/InterpolationMC.lean}{regeneration\_implies\_cofinal\_hitting}
{\color{leanink}$\checkmark$}):
for every threshold~$K$, there exists a valid path that hits
$-1\bmod q$ at some step~$n \geq K$, with positive weight.
\item \textbf{Iterated hitting}
(\lean{EM/Advanced/InterpolationMC.lean}{regeneration\_implies\_iterated\_hitting}
{\color{leanink}$\checkmark$}):
for every target number of $-1$ hits, there exists a valid path
achieving at least that many hits, each with positive weight.
The proof proceeds by induction on the target, concatenating walks
via \code{concatMixedSelection} and using Regeneration to obtain
fresh good accumulators past each existing hit.
\end{itemize}
Regeneration is the sole open hypothesis in this layer; it is
strictly weaker than DSL (which implies MC outright) and concerns
the persistence of a qualitative property (full reachability) along
the walk, not a quantitative equidistribution statement.

\paragraph{TreeSieveDecay: the sieve-theoretic bridge.}
Regeneration asks that \emph{every} intermediate accumulator is good.
A weaker, sieve-flavored variant is \defname{TreeSieveDecay}
(\lean{EM/Advanced/InterpolationMC.lean}{TreeSieveDecay}):
there exists a threshold~$P_0$ such that every squarefree~$P \geq P_0$
coprime to~$q$ is a good accumulator.
TSD is a ``large-$P$'' hypothesis: all sufficiently generic starting
points support full reachability.
The coprimality condition is necessary: when $q \mid P$, absorption
traps the walk at~$0$ permanently, so no unit is reachable.
TSD bypasses Regeneration entirely---for a single hit (which suffices
for MC), a cleaner dichotomy argument works.
The formalization proves two bridges:
\begin{itemize}
\item \textbf{Squarefree propagation}
(\lean{EM/Advanced/InterpolationMC.lean}{mixedWalkProd\_squarefree}
{\color{leanink}$\checkmark$}):
if the starting point is squarefree, every intermediate accumulator
is squarefree (induction, using coprimality of factors).
\item \textbf{Coprimality from no death}
(\lean{EM/Advanced/InterpolationMC.lean}{mixedWalkProd\_coprime\_of\_no\_death}
{\color{leanink}$\checkmark$}):
if the walk never hits $-1 \bmod q$, all intermediate accumulators
stay coprime to~$q$ (if the factor equaled~$q$, then
$q \mid P{+}1$, so the walk position would be~$-1$---contradiction).
\end{itemize}
Together: under TSD, from any sufficiently large squarefree~$m$,
the walk either hits~$-1$ (done) or stays coprime to~$q$, in which
case TSD applies to each intermediate accumulator directly
(\lean{EM/Advanced/InterpolationMC.lean}{tsd\_implies\_neg\_one\_reachable}
{\color{leanink}$\checkmark$}).

A weaker variant \defname{TreeSieveDecayHitting}
(\lean{EM/Advanced/InterpolationMC.lean}{TreeSieveDecayHitting})
requires only that $-1$ is reachable (not all units).
TSD implies TSDH
(\lean{EM/Advanced/InterpolationMC.lean}{tsd\_implies\_tsdh}
{\color{leanink}$\checkmark$}).

\paragraph{Orbit melting: backward propagation.}
Two squarefree positive integers with the same prime factor set are
equal
(\lean{EM/Advanced/InterpolationMC.lean}{eq\_of\_same\_primeFactors\_squarefree}
{\color{leanink}$\checkmark$}).
This ``orbit melting'' identity means the accumulator carries no
hidden state beyond its prime factorization:
same accumulator $\Rightarrow$ same reachable set
(\lean{EM/Advanced/InterpolationMC.lean}{same\_accumulator\_same\_future}
{\color{leanink}$\checkmark$}).
More importantly, a good accumulator at \emph{any} intermediate step
propagates backward: if $P_\sigma(K)$ supports full reachability,
then $-1$ is reachable from the starting point~$m$ by concatenating
the prefix path with the witnessing path
(\lean{EM/Advanced/InterpolationMC.lean}{good\_accumulator\_propagates\_to\_start}
{\color{leanink}$\checkmark$}).

\paragraph{TSD-Hitting(3): unconditional.}
For $q = 3$, TreeSieveDecayHitting holds \textbf{unconditionally}---no
open hypothesis at all.  The argument is a mod-$3$ dichotomy on any
squarefree~$P$ coprime to~$3$:
\begin{itemize}
\item If $P \equiv 2 \pmod{3}$: then $(P : \ZZ/3\ZZ) = -1$, so the
  walk is already at~$-1$ at step~$0$.
\item If $P \equiv 1 \pmod{3}$: then $P{+}1 \equiv 2 \pmod{3}$.
  A number $\geq 2$ that is $\equiv 2 \bmod 3$ and not divisible by~$3$
  must have a prime factor $p \equiv 2 \pmod{3}$
  (\lean{EM/Advanced/InterpolationMC.lean}{exists\_factor\_of\_succ\_mod3}
  {\color{leanink}$\checkmark$},
  by strong induction: if $\minFac \equiv 1$, divide out and recurse).
  Selecting that factor gives $P \cdot p \equiv 1 \cdot 2 = -1 \pmod{3}$
  at step~$1$.
\end{itemize}
\begin{keyresult}
\begin{theorem}[\lean{EM/Advanced/InterpolationMC.lean}{tsd\_hitting\_three\_unconditional}]
$\mathrm{TreeSieveDecayHitting}(3)$ holds unconditionally: for every
squarefree $P \geq 2$ coprime to~$3$, there exists a valid path in
the factor tree that reaches $-1 \bmod 3$ within one step.
\end{theorem}
\end{keyresult}

\begin{leancode}
-- The hypothesis: -1 reachable from all large squarefree coprime P
def TreeSieveDecayHitting (q : Nat) : Prop :=
  exists P_0, forall P, P_0 <= P -> Squarefree P -> Nat.Coprime P q ->
    (-1 : ZMod q) in reachableEver q P

-- UNCONDITIONAL for q = 3 (zero open hypotheses)
theorem tsd_hitting_three_unconditional : TreeSieveDecayHitting 3 := by
  -- mod-3 dichotomy: P == 2 -> already at -1;
  -- P == 1 -> P+1 == 2, so some prime factor p == 2 mod 3 exists
  ...
\end{leancode}

\noindent
This is the first unconditional \emph{tree-level} hitting result beyond the
trivial $q = 3$ capture of \S\ref{sec:factor-confinement}.
The previous $q{=}3$ result
(\lean{EM/Advanced/EpsilonRandomMC.lean}{mixed\_capture\_three}
{\color{leanink}$\checkmark$})
showed that $3$ appears in $\mathcal{T}(2)$ from the fixed root~$m{=}2$;
TSD-Hitting(3) proves that $-1$ is reachable from \emph{every}
coprime squarefree accumulator---a uniform statement that applies at
any intermediate point of the walk.

\paragraph{TSD $\Rightarrow$ MixedMC: composition.}
The threshold condition in TSD (accumulator $P$ must exceed
a bound~$P_0$) is eliminated by the exponential growth of the mixed
walk: the accumulator $P_\sigma(k)$ satisfies
$P_\sigma(k) \geq m \cdot 2^k$
(\lean{EM/Advanced/InterpolationMC.lean}{mixedWalkProd\_exp\_growth}
{\color{leanink}$\checkmark$}),
since each multiplier is a prime $\geq 2$.  For any starting
accumulator $m \geq 2$, the walk eventually exceeds~$P_0$; a
death/no-death dichotomy then applies: either the walk already hit
$-1$ (done), or all accumulators remain coprime to~$q$ and
squarefree, so TSD applies at the first step past the threshold, and
tail reachability (Part~9) propagates $-1$ back to the root.  This
gives the full chain:

\begin{theorem}[TSD reduction chain]
\label{thm:tsd-chain}
For each prime $q \geq 5$:
\[
  \mathrm{TSD}(q)
  \;\Longrightarrow\;
  \forall\, m \geq 2 \text{ sqfree, } \gcd(m,q)=1 \colon
  {-}1 \in R_\infty(q, m)
  \;\Longrightarrow\;
  \mathrm{MixedMC}(q).
\]
The first arrow uses exponential growth
(Theorem~\textup{\ref{thm:mixed-walk-structure}})
to exceed the TSD threshold
(\lean{EM/Advanced/InterpolationMC.lean}{tsd\_implies\_reachable\_from\_any}
{\color{leanink}$\checkmark$});
the second uses hit-implies-capture
(\lean{EM/Advanced/EpsilonRandomMC.lean}{hit\_implies\_capture'}
{\color{leanink}$\checkmark$}).
Consequently
(\lean{EM/Advanced/InterpolationMC.lean}{tsd\_all\_implies\_mixed\_mc}
{\color{leanink}$\checkmark$}):
\[
  (\forall q \geq 5,\; \mathrm{TSD}(q))
  \;\Longrightarrow\; \mathrm{MixedMullinConjecture}.
\]
\end{theorem}

\noindent Small primes $q \in \{2, 3\}$ are handled by their existing
unconditional results (\code{mixed\_mc\_two}, \code{mixed\_mc\_three}).

\paragraph{Coset ambiguity: why TSD-Hitting(3) is free but $q \geq 5$ is not.}
The structural reason that TSD-Hitting($q$) closes unconditionally
for $q = 3$ but requires a hypothesis for $q \geq 5$ is \emph{coset
ambiguity}.  For $q = 3$, the group $(\ZZ/3\ZZ)^\times = \{1, 2\}$
has a single non-identity element, which \emph{is} $-1 = 2$: any
prime factor escaping the trivial coset $\{1\}$ automatically
lands on~$-1$
(\lean{EM/Advanced/InterpolationMC.lean}{single\_coset\_implies\_immediate\_hit}
{\color{leanink}$\checkmark$}).
For $q = 5$, the group $(\ZZ/5\ZZ)^\times = \{1,2,3,4\}$ has the
subgroup $\{1, 4\} = \{1, -1\}$ and two other cosets $\{2\}$, $\{3\}$,
neither equal to~$-1$.  A concrete counterexample:
$P = 2$, $P{+}1 = 3$, $(3 \bmod 5) = 3 \neq 4 = -1$---the only
available factor misses~$-1$
(\lean{EM/Advanced/InterpolationMC.lean}{coset\_ambiguity\_counterexample}
{\color{leanink}$\checkmark$}).

\paragraph{Ensemble factor diversity and the stochastic EM process.}
The ensemble-level structure of the prime factor sets
$\text{primeFactors}(\Prod{n}(k)+1)$ is formalized in
\code{Advanced/FactorDiversity.lean} (346~lines, zero \code{sorry}).

\begin{theorem}[Spectral contraction from factor diversity ---
\lean{EM/Advanced/FactorDiversity.lean}{factor\_diversity\_spectral\_contraction}
{\color{leanink}$\checkmark$}]
\label{thm:factor-diversity-contraction}
If the prime factor set of $\Prod{n}(k)+1$ contains $\geq 2$ distinct
residue classes mod~$q$, then the mean character value over the factor
set satisfies $\|\bar\chi(F)\| < 1$ for every nontrivial~$\chi$.
If such diverse steps occur infinitely often with a uniform gap
$\delta > 0$, the averaged character product tends to zero
(\lean{EM/Advanced/FactorDiversity.lean}{diverse\_steps\_imply\_vanishing}
{\color{leanink}$\checkmark$}).
\end{theorem}

\noindent
The bridge from diversity to stochastic capture
(\code{Advanced/DiverseStepsToCapture.lean}, 281~lines) shows each
prime factor of $\Prod{2}(k)+1$ produces a distinct element of the
reachable set $R_\infty(q, 2)$
(\lean{EM/Advanced/DiverseStepsToCapture.lean}{genFactor\_in\_reachableEver}
{\color{leanink}$\checkmark$}).

\begin{theorem}[Integer stochastic MC ---
\lean{EM/Advanced/StochasticEM.lean}{stochastic\_mc\_of\_tsd}
{\color{leanink}$\checkmark$}]
\label{thm:integer-stochastic-mc}
$\mathrm{TSD}(q) \;\Longrightarrow\; \mathrm{StochasticMC}(q)$ for all
$\varepsilon > 0$.  The phase transition is sharp at $\varepsilon = 0$
(\lean{EM/Advanced/StochasticEM.lean}{phase\_transition\_summary}
{\color{leanink}$\checkmark$}):
character norms are exactly~$1$ (no contraction) for the deterministic
walk, but strictly~$< 1$ under any stochastic perturbation.
\end{theorem}

\paragraph{Probabilistic infrastructure.}
The TSD$\to$MixedMC chain is supported by three infrastructure
modules.  \code{Probability/TransitionKernel.lean} defines the
epsilon-mixed walk's step PMF and proves the weight lower bound
$\varepsilon/\omega(P{+}1)$ bridges to InterpolationMC's
\code{stepWeightLB}.  \code{Probability/PathMeasure.lean} converts
set-theoretic reachable sets to computable Finsets and formalizes
branching distinctness: when $P$ is a unit mod~$q$ and $P{+}1$ has
distinct factor residues, each gives a distinct next walk position.
\code{Probability/GeometricCapture.lean} provides the block-geometric
engine: per-block positive capture weight decays non-capture
probability as $(1{-}\delta)^K \to 0$, turning TSD into asymptotic
capture certainty.

\subsection{The $\{\minFac, \operatorname{secondMinFac}\}$ Variant}
\label{sec:minfac-secondminfac}

The two-point variant interpolates between the random and deterministic
extremes.  At each step, instead of a single deterministic choice or a
fully random one, two specific factors are available:
$\minFac(\Prod{2}(n)+1)$ and $\operatorname{secondMinFac}(\Prod{2}(n)+1)$.
A decision sequence $\sigma : \NN \to \mathrm{Bool}$ selects between
them; the all-true sequence recovers the standard EM walk.

\paragraph{The $\varepsilon$-walk.}
The formalization defines a
\textbf{self-consistent $\varepsilon$-walk}
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{epsWalkProd})
in which the chosen factor---either $\minFac(P(n){+}1)$ or
$\operatorname{secondMinFac}(P(n){+}1)$---becomes the actual multiplier,
updating the accumulator and producing a new Euclid number at the next
step.
The function $\operatorname{secondMinFac}(n)$ (second-smallest prime
factor) is defined and proved prime, divisor of~$n$, and
$\geq \minFac(n)$
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{secondMinFac\_prime},
\lean{EM/Advanced/VanishingNoiseVariant.lean}{secondMinFac\_dvd},
\lean{EM/Advanced/VanishingNoiseVariant.lean}{minFac\_le\_secondMinFac}).

This self-consistent walk generates a \emph{binary branching tree}
of depth~$N$ with $2^N$ leaves, where each internal node has its own
path-dependent factor set.  The \emph{tree character sum}
(\lean{EM/Advanced/VanishingNoise.lean}{treeCharSum}),
a weighted sum of $\chi(\text{leaf position})$ over all leaves, is proved
$\leq 1$ in norm.  Unlike the product multiset of the stochastic MC, this sum
does \textbf{not} factorize as $\prod_k \bar\chi(S_k)$: the left and
right subtrees start from different accumulators, so their character sums
differ.  The open hypothesis
\defname{TreeContractionHypothesis}
asserts that the tree character sum tends to zero for all nontrivial~$\chi$;
it is strictly weaker than DSL but strictly stronger than the product approximation.

\paragraph{Non-self-consistent variant MC.}
An independent route bypasses the tree contraction problem entirely.
At each EM step~$n$, form the two-element unit set
$\{\minFac(P(n){+}1),\, \operatorname{secondMinFac}(P(n){+}1)\} \bmod q$
in $(\ZZ/q\ZZ)^\times$.  When one factor equals~$q$
(a non-unit), fall back to $\operatorname{Finset.univ}$, whose
mean character value is~$0$ by orthogonality---giving perfect contraction
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{meanCharValue\_univ\_eq\_zero}).
Under \defname{UFDStrong}---spectral gaps of the padded unit sets
are non-summable for every nontrivial~$\chi$---the
sparse product contraction machinery gives vanishing averaged
products, and the Fourier counting argument yields
path existence.  Concretely:

\begin{theorem}[\lean{EM/Advanced/VanishingNoiseVariant.lean}{variant\_mc\_from\_ufd\_strong\_proved}]
\label{thm:variant-mc}
For every prime $q \geq 3$, if $\mathrm{UFDStrong}(q)$ holds, then
every element of $(\ZZ/q\ZZ)^\times$ is representable as a product
of selections from $\{\minFac,\operatorname{secondMinFac}\}$ of
consecutive Euclid numbers.  In particular, some selection path
hits~$-1$.
\end{theorem}

The verified chain is:
\[
  \mathrm{UFDStrong}(q)
  \;\Longrightarrow\;
  \text{vanishing products}
  \;\Longrightarrow\;
  \text{path existence}
  \;\Longrightarrow\;
  \text{VariantMC}.
\]
Every arrow is machine-checked with zero \code{sorry}.

\paragraph{The gap: UFDStrong.}
\defname{UFDStrong}$(q)$ asserts: for each nontrivial character
$\chi$ on $(\ZZ/q\ZZ)^\times$, the spectral gaps
$1 - \|\bar\chi(\text{paddedUnitSet}_n)\|$ are non-summable.
The hypothesis splits into two cases:
\begin{enumerate}[nosep]
\item \emph{Fallback case}: if infinitely many steps have
  $\minFac \equiv \operatorname{secondMinFac} \pmod{q}$ (or one
  equals~$q$), the padded set falls back to the full group, giving
  perfect contraction (gap~$= 1$) at each such step.
  Non-summability is immediate.
  \textbf{Proved}:
  \lean{EM/Advanced/VanishingNoiseVariant.lean}{ufd\_fallback\_not\_summable}.
\item \emph{Diversity case}: if cofinitely many steps have two
  distinct units mod~$q$, one needs the character values at those
  units to differ infinitely often---i.e., the ratio
  $\minFac/\operatorname{secondMinFac}$ does not persistently lie in
  $\ker(\chi)$.  This is a genuine number-theoretic question about
  Euclid number factorizations.  \textbf{Open}:
  \code{UFDImpliesUFDStrong}.
\end{enumerate}

\paragraph{Variant MC does not imply standard MC.}
The selection~$\sigma$ that achieves hitting may differ per
prime~$q$, and the virtual accumulator
$\prod_k \text{factor}_{\sigma(k)}$ is not the EM accumulator
$P(n)$.  The gap is the same orbit-specificity barrier
(Dead End~\#130,
\lean{EM/FunctionField/CyclicWalkCoverage.lean}{alternating\_walk\_misses\_two}):
generation does not imply coverage for deterministic walks.

\paragraph{Three routes to UFDStrong.}
The formalization identifies a hierarchy of progressively weaker
sufficient conditions for UFDStrong$(q)$:

\begin{enumerate}[nosep]
\item \textbf{Quantitative escape} (\defname{MinFacRatioEscape}):
  for each nontrivial~$\chi$, there exists $\delta > 0$ such that
  infinitely many steps have spectral gap $\geq \delta$.
  Since a non-summable sequence with a positive lower bound infinitely
  often is non-summable, this implies UFDStrong directly
  (\lean{EM/Advanced/VanishingNoiseVariant.lean}{ratio\_escape\_implies\_ufdStrong}).

\item \textbf{Qualitative escape} (\defname{MinFacRatioEscapeQual}):
  for each nontrivial~$\chi$, infinitely many steps have
  $\operatorname{card}(\text{rawTwoPointUnitSet}) \geq 2$ \emph{and}
  $\chi(\minFac) \neq \chi(\operatorname{secondMinFac})$.
  This implies quantitative escape via a finite-group pigeonhole
  argument: on a finite group, the gap function takes finitely many
  values, so being positive infinitely often gives a uniform positive
  lower bound
  (\lean{EM/Advanced/VanishingNoiseVariant.lean}{qual\_implies\_quant}).

\item \textbf{Orbit-level minFac equidistribution} (\defname{OrbitMFRE}):
  the residues $\minFac(\Prod{2}(n){+}1) \bmod q$ are equidistributed
  among the units of $(\ZZ/q\ZZ)^\times$ in Cesàro density.
  This is the weakest condition in the hierarchy: equidistributed
  residues prevent persistent chi-degeneracy.  The bridge
  OrbitMFRE $\Rightarrow$ MinFacRatioEscapeQual is \textbf{open}
  (\code{OrbitMFREImpliesEscapeQual}).
\end{enumerate}

The assembled chain is:
\[
  \underbrace{\mathrm{OrbitMFRE}}_{\text{open}}
  \;\xRightarrow{\text{open bridge}}\;
  \mathrm{EscapeQual}
  \;\xRightarrow{\text{proved}}\;
  \mathrm{EscapeQuant}
  \;\xRightarrow{\text{proved}}\;
  \mathrm{UFDStrong}
  \;\xRightarrow{\text{proved}}\;
  \mathrm{VariantMC}.
\]
Every arrow except the first is machine-checked with zero \code{sorry}.
The most concrete attack target is MinFacRatioEscapeQual: one needs to
show that $\minFac$ and $\operatorname{secondMinFac}$ do not
persistently agree modulo every character.  OrbitMFRE faces the same
orbit-specificity barrier as DSL for standard MC.

\paragraph{Stochastic Two-Point MC.}
The non-self-consistent machinery yields a clean quantitative statement.
Define the \defname{StochasticTwoPointMC} problem: under UFDStrong$(q)$
with $q \geq 3$, does every element of $(\ZZ/q\ZZ)^\times$ appear with
positive count in the product multiset of padded unit sets?

\begin{theorem}[\lean{EM/Advanced/VanishingNoiseVariant.lean}{stochastic\_two\_point\_mc\_proved}]
\label{thm:stochastic-two-point}
$\mathrm{UFDStrong}(q) \;\Longrightarrow\; \mathrm{StochasticTwoPointMC}(q)$.
\end{theorem}

\noindent The product multiset grows exponentially
($\geq 2^N$ paths for $q \geq 3$,
\lean{EM/Advanced/VanishingNoiseVariant.lean}{productMultiset\_card\_ge\_two\_pow}).
A \defname{fairCoin} schedule ($\varepsilon = 1/2$ at every step) gives
the \defname{fairTreeCharSum}---the tree character sum at constant
$1/2$---which is bounded by~$1$ in norm
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{fairTreeCharSum\_norm\_le\_one}).
TreeContractionHypothesis implies the weaker
\defname{TreeContractionAtHalf}
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{treeContraction\_of\_hypothesis}),
and at $\varepsilon = 0$ the tree degenerates to the deterministic
$\minFac$ path
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{treeCharSum\_at\_zero\_step}).

The gap between the non-self-consistent result (proved) and a
\emph{self-consistent} stochastic MC is the \textbf{selection coherence}
problem: the non-self-consistent product multiset uses factor sets from
the actual EM walk (which always picks $\minFac$), so the factor sets
are fixed and the product factorizes; in the self-consistent tree,
picking $\operatorname{secondMinFac}$ at step~$n$ changes the
accumulator, hence the factor set at step~$n{+}1$, preventing
factorization.

\paragraph{Faithful character escape.}
A character $\chi$ on $(\ZZ/q\ZZ)^\times$ is \defname{faithful} if it
is injective (equivalently, $\ker(\chi) = \{1\}$).  For faithful
characters, whenever the raw two-point unit set has two distinct
elements, their $\chi$-values are automatically distinct---giving a
positive spectral gap at every non-fallback step, with no
number-theoretic hypothesis.  Combined with the fallback case (gap $= 1$
at fallback steps, already proved), this yields unconditional
non-summability:

\begin{theorem}[\lean{EM/Advanced/VanishingNoiseVariant.lean}{faithful\_character\_escape}]
\label{thm:faithful-escape}
For every prime $q \geq 3$ and every faithful nontrivial character
$\chi$ on $(\ZZ/q\ZZ)^\times$:
\[
  \sum_n \bigl(1 - \|\bar\chi(\text{paddedUnitSet}_n)\|\bigr)
  = +\infty.
\]
Unconditional---zero open hypotheses.
\end{theorem}

\noindent For groups of prime order, every nontrivial character is
faithful (by Lagrange: the kernel is a subgroup of a prime-order group,
hence trivial or the whole group;
\lean{EM/Advanced/VanishingNoiseVariant.lean}{prime\_order\_nontrivial\_is\_faithful}).
Since $|(\ZZ/3\ZZ)^\times| = 2$ is prime, UFDStrong$(3)$ holds
\emph{unconditionally}
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{ufdStrong\_three}), and
the full chain closes:

\begin{theorem}[\lean{EM/Advanced/VanishingNoiseVariant.lean}{variant\_mc\_three\_unconditional}]
\label{thm:variant-mc-3}
$\mathrm{StochasticTwoPointMC}(3)$ holds with zero open hypotheses:
every element of $(\ZZ/3\ZZ)^\times$ appears with positive count
in the product multiset.
\end{theorem}

\noindent For primes $q \geq 5$ where $|(\ZZ/q\ZZ)^\times| = q{-}1$ is
composite, non-faithful nontrivial characters exist, and the sole
remaining open hypothesis is
\defname{NonFaithfulCharacterEscape}---spectral gap non-summability
for characters with nontrivial kernel.  The faithful/non-faithful
decomposition is clean:

\begin{theorem}[\lean{EM/Advanced/VanishingNoiseVariant.lean}{nfce\_implies\_ufdStrong}]
\label{thm:nfce-ufdstrong}
$\mathrm{NonFaithfulCharacterEscape}(q)
\;\Longrightarrow\; \mathrm{UFDStrong}(q)$
for all primes $q \geq 3$.
\end{theorem}

\noindent The faithful half is free; only the non-faithful half
remains open.

\paragraph{NFCE infrastructure.}
The formalization develops a precise characterization of what NFCE
requires.  Define the \defname{factorRatio} at step~$n$ as
$\mathrm{liftToUnit}(\minFac) \cdot
\mathrm{liftToUnit}(\operatorname{secondMinFac})^{-1}$
in~$(\ZZ/q\ZZ)^\times$.
At non-fallback steps (raw card $\geq 2$), the spectral gap is zero
if and only if this ratio lies in~$\ker(\chi)$
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{gap\_zero\_iff\_ratio\_in\_ker}).
NFCE thus reduces to: for each non-faithful nontrivial~$\chi$, the
factor ratio does not eventually stay in~$\ker(\chi)$.

If NFCE fails for a specific~$\chi$, the gaps are summable,
which forces \defname{KernelConfinement}: the ratio is eventually
confined to~$\ker(\chi)$
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{summable\_implies\_ratio\_confined}).
Since $\ker(\chi)$ has index $\geq 2$ for nontrivial~$\chi$
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{ker\_index\_ge\_two}),
kernel confinement is a strong arithmetic constraint:
$\minFac / \operatorname{secondMinFac} \bmod q$ eventually lies in a
proper subgroup of the unit group.

\paragraph{NFCE(5): reduction to a single quadratic escape.}
For $q = 5$, the unit group $(\ZZ/5\ZZ)^\times$ has order~$4$
(cyclic, $\cong \ZZ/4\ZZ$).  Its characters partition as: the trivial
character, two faithful characters (order~$4$, handled unconditionally
by FaithfulCharacterEscape), and one non-faithful nontrivial character
(order~$2$, the unique quadratic character).  Thus NFCE(5) reduces to a
single escape condition: the factor ratio
$\minFac / \operatorname{secondMinFac} \bmod 5$ must leave the unique
index-2 subgroup of $(\ZZ/5\ZZ)^\times$ infinitely often.
The formalization proves that this
\defname{RatioKernelEscape} condition suffices for the full chain
\lean{EM/Advanced/VanishingNoiseVariant.lean}{variant\_mc\_five\_of\_ratio\_escape}
{\color{leanink}$\checkmark$}:
RatioKernelEscape $\Rightarrow$ NFCE(5) $\Rightarrow$ UFDStrong(5)
$\Rightarrow$ StochasticTwoPointMC(5).

\paragraph{Intersection kernel dichotomy.}
When NFCE fails for \emph{all} non-faithful nontrivial characters
simultaneously (\defname{TotalNFCEFailure}), the factor ratio is
eventually confined to every non-faithful kernel (from the
summability-to-confinement lemma of Part~25).
The intersection of these kernels is constrained by a group-theoretic
fact: in a finite group whose order has $\geq 2$ distinct prime
factors, there exist subgroups of coprime orders whose intersection is
trivial
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{coprime\_order\_zpowers\_inf\_bot}
{\color{leanink}$\checkmark$}).  Combined with Cauchy's theorem, this
yields \defname{NonFaithfulCharSeparation}~(NFCS): for every
$g \neq 1$, some non-faithful nontrivial character has $\chi(g) \neq 1$.
NFCS is
\textbf{proved} for all groups whose order has $\geq 2$ distinct prime
factors
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{nonFaithfulCharSeparation\_of\_two\_prime\_factors}
{\color{leanink}$\checkmark$}), which for $(\ZZ/q\ZZ)^\times$ covers
all primes except $q = 2, 3, 5$ and Fermat primes (where $q - 1$ is a
power of~$2$).
Under NFCS, TotalNFCEFailure forces
$\minFac / \operatorname{secondMinFac} \equiv 1 \pmod{q}$ eventually
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{total\_confinement\_implies\_ratio\_one}
{\color{leanink}$\checkmark$}).
This yields:

\begin{theorem}[Intersection kernel dichotomy]
\label{thm:kernel-dichotomy}
Let $q$ be prime with $q - 1$ having $\geq 2$ distinct prime factors.
Then exactly one of the following holds:
\begin{enumerate}[nosep]
\item $\mathrm{NFCE}(q)$ holds, giving
  $\mathrm{UFDStrong}(q)$ and $\mathrm{StochasticTwoPointMC}(q)$; or
\item $\minFac(\Prod{2}(n)+1) \equiv \operatorname{secondMinFac}(\Prod{2}(n)+1) \pmod{q}$
  for all sufficiently large~$n$.
\end{enumerate}
The second alternative is a strong arithmetic rigidity: the two
smallest prime factors of each Euclid number are eventually
indistinguishable mod~$q$.
\end{theorem}

\noindent
The proof combines NonFaithfulCharSeparation
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{nonFaithfulCharSeparation\_of\_two\_prime\_factors}
{\color{leanink}$\checkmark$})
with the summability-to-confinement lemma
(\lean{EM/Advanced/VanishingNoiseVariant.lean}{total\_confinement\_implies\_ratio\_one}
{\color{leanink}$\checkmark$}).
The condition ``$q{-}1$ has $\geq 2$ distinct prime factors'' covers
all primes except $q = 2, 3, 5$ and Fermat primes (where $q{-}1$ is a
power of~$2$).

\paragraph{Quotient character lift.}
The quotient character infrastructure, formalized in Part~28 of
\code{VanishingNoiseVariantD.lean} (173~lines),
provides a structural reduction of NFCE\@.
For a non-faithful nontrivial character~$\chi$, the kernel
$\ker(\chi)$ is a proper nontrivial subgroup; by the universal
property, $\chi$ factors as
$\chi = \bar\chi \circ \pi$
where $\pi \colon G \to G/\ker(\chi)$ is the quotient map.
The induced character~$\bar\chi$ on the quotient group is
\textbf{faithful} (injective)
(\lean{EM/Advanced/VanishingNoiseVariantD.lean}{quotientChar\_faithful}
{\color{leanink}$\checkmark$}),
the quotient has order $\geq 2$
(\lean{EM/Advanced/VanishingNoiseVariantD.lean}{quotient\_card\_ge\_two}
{\color{leanink}$\checkmark$}),
and KernelConfinement is equivalent to the quotient projection
of the factor ratio being eventually trivial
(\lean{EM/Advanced/VanishingNoiseVariantD.lean}{kernelConfinement\_iff\_quotient\_eventually\_one}
{\color{leanink}$\checkmark$}).
\begin{theorem}[Quotient character lift]
\label{thm:quotient-lift}
For any non-faithful nontrivial character~$\chi$ on~$(\ZZ/q\ZZ)^\times$:
\begin{enumerate}[nosep]
\item $\chi = \bar\chi \circ \pi$ where $\bar\chi$ is faithful
  on $G/\ker(\chi)$ and $|G/\ker(\chi)| \geq 2$.
\item $\mathrm{KernelConfinement}(\chi) \;\Longleftrightarrow\;$
  the quotient projection of the factor ratio is eventually trivial.
\item $\mathrm{RatioKernelEscape}(\chi) \;\Longleftrightarrow\;$
  the quotient projection is nontrivial infinitely often.
\end{enumerate}
\end{theorem}

\noindent This reduces NFCE to: the image of factorRatio in each quotient
$G/\ker(\chi)$ escapes~$\{1\}$ infinitely often.
Since $\bar\chi$ is faithful on the (smaller) quotient group, the
unconditional FaithfulCharacterEscape result
(Theorem~\ref{thm:faithful-escape}) applies to $G/\ker(\chi)$,
connecting the non-faithful case to the faithful framework.

\paragraph{Comparison of open hypotheses.}
Standard MC requires DSL (population-to-orbit transfer for $\minFac$
residues).  The $\{\minFac, \operatorname{secondMinFac}\}$ variant
requires UFDStrong (spectral diversity of two-point factor sets), and
for $q = 3$ this is now \textbf{unconditional}; for $q \geq 5$ the
residual hypothesis is NonFaithfulCharacterEscape.
Neither DSL nor UFDStrong implies the other: DSL is about the
\emph{selected} factor's equidistribution; UFDStrong is about the
\emph{pair}'s spectral gap.
But both are instances of the same meta-question: does the deterministic
EM construction produce enough diversity in its factorizations to
prevent persistent alignment with any fixed character?

\paragraph{Ensemble two-point: population-level reduction.}
The ensemble framework of Section~\ref{sec:ensemble} can also be
applied to the two-point variant.  For generalized EM starting from
squarefree~$n$, the \defname{genFactorRatio} at step~$k$ is the ratio
$\minFac / \operatorname{secondMinFac} \bmod q$ in
$(\ZZ/q\ZZ)^\times$.  For a nontrivial character~$\chi$, a step
``escapes'' if this ratio lies outside~$\ker(\chi)$---producing a
spectral gap.  The formalization
(\code{Ensemble/TwoPointEnsemble.lean}, 566~lines, zero \code{sorry})
develops the full first-moment chain:
\[
  \text{PRED}
  \;\xrightarrow[\text{Fubini}]{\checkmark}\;
  \text{linear first moment}
  \;\xrightarrow[\text{partition}]{\checkmark}\;
  \text{positive density of high escapers.}
\]
Here \textbf{PopulationRatioEscapeDensity}~(PRED) asserts that for
each nontrivial~$\chi$ and each step $k \geq 1$, a positive
density~$\delta > 0$ of squarefree starting points have factor ratio
outside~$\ker(\chi)$.  Under PRED, the ensemble average of cumulative
escape counts grows linearly
(\lean{EM/Ensemble/TwoPointEnsemble.lean}{escape\_first\_moment\_linear}
{\color{leanink}$\checkmark$}), and the partition argument gives
\lean{EM/Ensemble/TwoPointEnsemble.lean}{positive\_density\_high\_escape}
{\color{leanink}$\checkmark$}: for each threshold~$M$, a positive
density of starting points achieve $\geq M$ cumulative escapes
within~$K$ steps.  PRED itself is open; it should follow from
MFRE (MinFacResidueEquidist) via CRT independence of the two
smallest factors.

\paragraph{Self-consistent random walk: the Fourier bridge.}
The stochastic two-point variant above fixes the factor sets
$\{p_1(n), p_2(n)\}$ from the \emph{actual} EM walk and randomizes
the selection.  A stronger variant makes the construction
\emph{self-consistent}: at each step, the accumulator determines the
available factors, and the fair-coin choice feeds back into the next
accumulator.  Formally, define
$\mathrm{epsWalkProdFrom}(\mathrm{acc}, \sigma, 0) = \mathrm{acc}$,
$\mathrm{epsWalkProdFrom}(\mathrm{acc}, \sigma, n{+}1)
= P_n \cdot (\text{if } \sigma(n) \text{ then } \minFac(P_n{+}1)
\text{ else } \operatorname{secondMinFac}(P_n{+}1))$
where $P_n = \mathrm{epsWalkProdFrom}(\mathrm{acc}, \sigma, n)$.
The decision sequence $\sigma : \NN \to \mathrm{Bool}$ replaces the
deterministic $\minFac$ rule with a random binary choice.

The central result is the \textbf{Fourier bridge identity}
(\lean{EM/Advanced/RandomTwoPointMC.lean}{fourier\_bridge\_identity\_proved}
{\color{leanink}$\checkmark$}): for any accumulator $\mathrm{acc} \geq 2$,
\[
  \mathrm{treeCharSum}(q, \chi, N, \mathrm{acc}, \tfrac{1}{2})
  \;=\;
  \frac{1}{2^N}
  \sum_{\sigma \in \{0,1\}^N}
  \prod_{k < N} \mathrm{chiAt}(q, \chi,
    \mathrm{epsWalkFactorFrom}(\mathrm{acc}, \sigma, k)).
\]
The proof is by induction on~$N$: at each step, the $2^N$ paths split
by first bit into two halves corresponding to the left and right
subtrees of the tree recursion; shift lemmas
(\code{epsWalkProdFrom\_shift}, \code{epsWalkFactorFrom\_shift})
connect the concatenated walk to the shifted walk at the new
accumulator; the $\tfrac{1}{2}$ weights combine to match the tree
recurrence.  Specializing to $\mathrm{acc} = 2$ gives the
unconditional identity
$\mathrm{fairTreeCharSum} = \mathrm{pathCharSum}$
(\lean{EM/Advanced/RandomTwoPointMC.lean}{fairTreeCharSum\_eq\_pathCharSum}
{\color{leanink}$\checkmark$}).

\defname{RandomTwoPointMC}
(\lean{EM/Advanced/RandomTwoPointMC.lean}{RandomTwoPointMC})
asks: for $q \geq 3$ and every unit $a \in (\ZZ/q\ZZ)^\times$, does
some binary decision sequence of finite length produce a walk from
$\mathrm{acc} = 2$ landing in residue class~$a$?  This is the
self-consistent analog of StochasticTwoPointMC.
The Fourier bridge connects the tree to the path character sum.
To connect path character sums to reached residue classes, one needs
\defname{ChiAtMultiplicativity}: that $\mathrm{chiAt}(\mathrm{acc})
\cdot \prod_k \mathrm{chiAt}(\text{factor}_k)
= \mathrm{chiAt}(\text{endpoint})$, provided $q$ does not divide the
endpoint.  This is \textbf{proved}
(\lean{EM/Advanced/RandomTwoPointMC.lean}{chiAt\_multiplicativity\_proved}
{\color{leanink}$\checkmark$})
by induction on~$N$ with backward non-divisibility propagation:
if $q \nmid P_N$, then $q \nmid P_k$ for all $k \leq N$
(\lean{EM/Advanced/RandomTwoPointMC.lean}{not\_dvd\_epsWalkProdFrom\_of\_le}
{\color{leanink}$\checkmark$}),
so every factor is coprime to~$q$ and $\mathrm{chiAt}$ is
multiplicative.

A full Fourier counting infrastructure is built on top: character
orthogonality on $(\ZZ/q\ZZ)^\times$, death/survival partition
(\lean{EM/Advanced/RandomTwoPointMC.lean}{survival\_plus\_death}
{\color{leanink}$\checkmark$}:
$\mathrm{survivalCount} + \mathrm{deathCount} = 2^N$),
the Fourier indicator formula
$|G| \cdot \mathrm{pathCount}(a)
= \sum_\chi \overline{\chi(a)} \cdot U(\chi, N)$,
and norm bounds for the unit endpoint character sum.
The chain is:
\[
  \mathrm{TCA}
  \;\xRightarrow{\checkmark}\;
  \text{vanishing pathCharSum}
  \;\xRightarrow{\text{ChiAtMult}\ \checkmark}\;
  \text{Fourier counting}
  \;\xRightarrow{?}\;
  \text{RandomTwoPointMC}.
\]
The remaining gap is \defname{PathSurvival}
(\lean{EM/Advanced/RandomTwoPointMC.lean}{PathSurvival}):
the ratio $\mathrm{survivalCount}(N) / \mathrm{deathCount}(N)$
grows without bound.  Without positive survival at the same~$N$
where pathCharSum is small, the Fourier inversion gives no
information about unit classes.

For $q = 3$, the analysis reveals a structural obstruction.
Since $\mathrm{acc} = 2$ and $\minFac(3) = \operatorname{secondMinFac}(3) = 3$,
the walk immediately dies: $3 \mid P_1 = 6$ on \emph{all} paths,
and absorption is permanent.
Thus PathSurvival(3) is false---but
$\mathrm{TreeContractionAtHalf}(3)$ is \emph{also} false (the tree
character sum is identically~$1$ at every depth, because all
post-death factors are $\equiv 1 \pmod{3}$), making
$\mathrm{TreeContractionImpliesRandomMC}(3)$ vacuously true.

For $q = 3$, the \textbf{zero-mean property} gives maximal per-step
contraction
(\lean{EM/Advanced/RandomTwoPointMC.lean}{mean\_char\_zero\_at\_diverse\_step\_three}
{\color{leanink}$\checkmark$}):
$(1/2)\chi(a) + (1/2)\chi(b) = 0$ whenever $\chi(a) \neq \chi(b)$
for a nontrivial character on $(\ZZ/3\ZZ)^\times$ (order~$2$, so
$\chi$ maps to $\{1, -1\}$).  Although this property is proved,
the immediate death at $q = 3$ means the zero-mean steps are
never reached from the standard starting point.

\paragraph{TCA $+$ PathSurvival $\Rightarrow$ RandomTwoPointMC.}
The conditional implication is now \textbf{proved}
(\lean{EM/Advanced/RandomTwoPointMC.lean}{tca\_path\_survival\_implies\_random\_mc\_proved}
{\color{leanink}$\checkmark$}).
The proof~(${\sim}400$~lines) builds the full Fourier counting
infrastructure on $(\ZZ/q\ZZ)^\times$:
\begin{enumerate}[nosep]
\item \textbf{Path sum decomposition}
  (\lean{EM/Advanced/RandomTwoPointMC.lean}{path\_sum\_decomp}
  {\color{leanink}$\checkmark$}):
  the unscaled path sum
  $2^N \cdot \mathrm{pathCharSum}$ splits as
  $\mathrm{chiAt}(2)^{-1}\cdot U(\chi,N) + E(\chi,N)$,
  where $U$ sums over surviving paths via ChiAtMultiplicativity and
  $E$ sums over death paths.

\item \textbf{Death error bound}
  (\lean{EM/Advanced/RandomTwoPointMC.lean}{death\_error\_norm\_le}
  {\color{leanink}$\checkmark$}):
  $\|E(\chi,N)\| \leq \mathrm{deathCount}(N)$
  (each death term has norm $\leq 1$).

\item \textbf{Unit endpoint bound}
  (\lean{EM/Advanced/RandomTwoPointMC.lean}{uecsi\_norm\_le}
  {\color{leanink}$\checkmark$}):
  $\|U(\chi,N)\| \leq 2^N\|\mathrm{pathCharSum}\| + \mathrm{deathCount}$.

\item \textbf{Fourier counting identity}
  (\lean{EM/Advanced/RandomTwoPointMC.lean}{fourier\_counting\_identity}
  {\color{leanink}$\checkmark$}):
  $\sum_\chi \overline{\chi(a)} \cdot U(\chi,N)
  = |G| \cdot \mathrm{pathCount}(a,N)$.

\item \textbf{Fourier lower bound}
  (\lean{EM/Advanced/RandomTwoPointMC.lean}{fourier\_lower\_bound}
  {\color{leanink}$\checkmark$}):
  $(q{-}1)\cdot\mathrm{pathCount}(a) \geq
  \mathrm{survivalCount} - (q{-}2)\cdot B$
  where $B$ is the uniform bound on nontrivial $\|U(\chi)\|$.
\end{enumerate}
Combining: TCA gives $\|\mathrm{pathCharSum}\| < 1/(2q)$ for
$N \geq N_0$; PathSurvival gives
$\mathrm{survivalCount} > 2q^2 \cdot \mathrm{deathCount}$ for
$N \geq N_1$.  At $N = \max(N_0,N_1)$, the Fourier lower bound yields
$\mathrm{pathCount}(a) > 0$ for every unit~$a$, via the algebraic
identity $2q^3 + 2q^2 + 3q + 2 > 0$ (true for all $q \geq 5$).
For $q = 3$, TCA is vacuously false, so the implication holds trivially.
The self-consistent RandomTwoPointMC thus reduces to exactly two open
hypotheses: TreeContractionAtHalf and PathSurvival.

\paragraph{Ensemble $\varepsilon$-walk.}
The formalization
(\code{Advanced/EpsilonWalk.lean}, 458~lines, zero \code{sorry})
develops an ensemble version: instead of fixing $\mathrm{acc} = 2$,
parametrize the $\varepsilon$-walk by squarefree starting point~$n$.
The \defname{SigmaCRTPropagationStep} hypothesis (open) asserts that
CRT equidistribution propagates across steps for $\sigma$-walks,
mirroring CRTPropagationStep for deterministic walks.
Under SRE $+$ SigmaCRTPropagationStep, the death density (fraction of
squarefree~$n$ with $q \mid \mathrm{epsWalkProdFrom}(n,\sigma,k)+1$)
converges to~$1/(q{-}1)$ for every fixed~$k$ and~$\sigma$
(\lean{EM/Advanced/EpsilonWalk.lean}{sigma\_death\_density\_tendsto}
{\color{leanink}$\checkmark$}).

\paragraph{Deterministic coverage via Cauchy--Davenport.}
An independent, purely group-theoretic route to coverage is formalized in
\code{Advanced/IteratedProductCoverage.lean} (293~lines, zero \code{sorry}).
Given a sequence $S_0, S_1, \ldots$ of subsets of a finite commutative
group~$G$ with $\mathrm{minOrder}(G) = |G|$ (true when $|G|$ is prime),
the \defname{iterated product} $P_n = P_{n-1} \cdot S_{n-1}$ starting
from $P_0 = \{1\}$ satisfies
\[
  |P_n| \;\geq\; \min\Bigl(|G|,\; 1 + \sum_{k<n}(|S_k| - 1)\Bigr)
\]
by iterated application of the Cauchy--Davenport theorem
(\lean{EM/Advanced/IteratedProductCoverage.lean}{iteratedMulFinset\_card\_growth}
{\color{leanink}$\checkmark$}).
If every $|S_k| \geq 2$, then after $|G| - 1$ steps the product is
$\mathrm{Finset.univ}$
(\lean{EM/Advanced/IteratedProductCoverage.lean}{iteratedMulFinset\_eq\_univ}
{\color{leanink}$\checkmark$}).
For $(\ZZ/p\ZZ)^\times$ with $p{-}1$ prime (safe primes: $p = 5, 7, 11, 23, \ldots$),
$\mathrm{minOrder} = p{-}1 = |G|$, and the coverage bound gives diameter
$p - 2$
(\lean{EM/Advanced/IteratedProductCoverage.lean}{safe\_prime\_coverage}
{\color{leanink}$\checkmark$}).
This complements the probabilistic Fourier approach: if the
$\varepsilon$-walk factor sets have $|S_k| \geq 2$ at every step (the
diversity condition), then Cauchy--Davenport gives deterministic coverage
in finitely many steps---no spectral hypothesis needed.  The gap is
ensuring diversity: when $\minFac \equiv \operatorname{secondMinFac}
\pmod{q}$, the factor set has size~$1$, and the Cauchy--Davenport
step gives no progress.

\paragraph{Dense capture and Baire category.}
A topological perspective on the two-point variant is developed in
\code{Advanced/DenseCapture.lean} (301~lines, zero \code{sorry}).
The space of binary selection sequences $\sigma : \NN \to \mathrm{Bool}$
carries the product topology (Cantor space), which is compact Hausdorff
and hence a Baire space.  The \defname{captureSet}$(q, \mathrm{acc})$---the
set of selection sequences that capture prime~$q$ starting from
accumulator~$\mathrm{acc}$---is \emph{open}, unconditionally
(\lean{EM/Advanced/DenseCapture.lean}{captureSet\_isOpen}
{\color{leanink}$\checkmark$}):
it equals $\bigcup_k \{\sigma \mid \mathrm{factor}(k) = q\}$,
and each level set is open because the factor at step~$k$
depends only on $\sigma(0), \ldots, \sigma(k)$ (finitary dependence,
\lean{EM/Advanced/DenseCapture.lean}{epsWalkFactorFrom\_depends\_on\_prefix}
{\color{leanink}$\checkmark$}).

The sole open hypothesis is \defname{DenseCaptureHypothesis}: for
$\mathrm{acc} \geq 2$, captureSet$(q, \mathrm{acc})$ is \emph{dense}.
Under this hypothesis, the Baire category theorem gives that the
intersection $\bigcap_{q\;\text{prime}} \mathrm{captureSet}(q, \mathrm{acc})$
is \emph{comeager} (residual): a generic selection sequence captures
every prime
(\lean{EM/Advanced/DenseCapture.lean}{fullCapture\_residual}
{\color{leanink}$\checkmark$}).
In particular, such a sequence exists
(\lean{EM/Advanced/DenseCapture.lean}{fullCapture\_nonempty}
{\color{leanink}$\checkmark$}).
A tail restart identity
(\lean{EM/Advanced/DenseCapture.lean}{epsWalkProdFrom\_tail\_restart}
{\color{leanink}$\checkmark$})
shows the walk from step~$K$ onward is a fresh walk from the
accumulator at step~$K$, connecting the density hypothesis to
the ensemble framework.

%%% The Mixed (1-eps)minFac + eps Random section has been subsumed by
%%% §6.2 (Factor Tree), §6.3 (Factor Confinement), §6.4 (Almost All GenMixedMC).

\subsection{Ensemble MC: Almost Every Starting Point}
\label{sec:ensemble-variant}

The variants above change the \emph{selection rule}; here we change the
\emph{starting point}.  The generalized EM construction
(Section~\ref{sec:ensemble}) parametrizes the sequence by its initial
value~$n$: $\Prod{n}(0) = n$,
$\seq{n}(k) = \minFac(\Prod{n}(k)+1)$,
$\Prod{n}(k{+}1) = \Prod{n}(k)\cdot\seq{n}(k)$.
The standard orbit is $n = 2$.

\begin{definition}[\lean{EM/Ensemble/WeylChain.lean}{GenMullinConjecture}]
$\mathrm{GenMC}(n)$: every prime appears in the generalized EM
sequence from~$n$, i.e.\ $\forall q$ prime, $\exists k$,
$\seq{n}(k) = q$.
\end{definition}

The reduction
\lean{EM/Ensemble/WeylChain.lean}{gen\_hitting\_implies\_gen\_mc\_proved}
{\color{leanink}$\checkmark$} proves that cofinal walk hitting
implies~$\mathrm{GenMC}(n)$ for every squarefree~$n$, mirroring the
standard inductive bootstrap.  And
\lean{EM/Ensemble/WeylChain.lean}{gen\_mc\_two\_implies\_mc}
{\color{leanink}$\checkmark$} shows $\mathrm{GenMC}(2) \Rightarrow
\mathrm{MC}$, so the generalized conjecture specializes to the
standard one.

\paragraph{Bag arithmetic.}
At each step~$k$, the Euclid number $\Prod{n}(k)+1$ has a ``bag'' of
prime factors; the selected prime $\seq{n}(k)$ is the minimum of this
bag.  The formalization
(\code{Ensemble/BagArithmetic.lean}, 224~lines, zero \code{sorry})
develops the bag-level quantities:
\defname{genEuclidOmega}~$\omega(n,k) = |\text{primeFactors}(\Prod{n}(k)+1)|$
(at least~1,
\lean{EM/Ensemble/BagArithmetic.lean}{genEuclidOmega\_pos}
{\color{leanink}$\checkmark$}),
\defname{genBagDiversity} (number of distinct residue classes mod~$q$
among the factors, $\leq \omega$ and $\leq q$),
\defname{genFactorsInClass}~$F_a$ (factors in class~$a$, with the
partition identity
$\sum_a |F_a| = \omega$,
\lean{EM/Ensemble/BagArithmetic.lean}{genFactorsInClass\_card\_sum}
{\color{leanink}$\checkmark$}),
and \defname{genEuclidCofactor} (the quotient
$(\Prod{n}(k)+1)/\seq{n}(k)$, whose prime factors are a subset of
the bag,
\lean{EM/Ensemble/BagArithmetic.lean}{cofactor\_primeFactors\_subset}
{\color{leanink}$\checkmark$}).
For $k \geq 1$, all bag elements are $\geq 3$
(\lean{EM/Ensemble/BagArithmetic.lean}{primeFactors\_ge\_three\_of\_pos}
{\color{leanink}$\checkmark$}).

\paragraph{Density variants.}
Instead of proving $\mathrm{GenMC}(n)$ for the single orbit $n = 2$,
one can ask how many starting points satisfy it.  Two density levels
are formalized:

\begin{itemize}
\item \textbf{PositiveDensityRSD}
  (\lean{EM/Population/ReciprocalSum.lean}{PositiveDensityRSD}):
  a positive density~$\delta > 0$ of squarefree~$n$ have divergent
  reciprocal sums $\sum_k 1/\seq{n}(k) = \infty$.

\item \textbf{AlmostAllSquarefreeRSD}
  (\lean{EM/Population/ReciprocalSum.lean}{AlmostAllSquarefreeRSD}):
  density~$1$ of squarefree~$n$ have divergent reciprocal sums.
\end{itemize}

\noindent
The reciprocal sum is a natural proxy: if $\sum 1/\seq{n}(k)$
diverges, then the primes appearing in the sequence from~$n$ cannot be
confined to a thin set; although divergence alone does not formally
imply~$\mathrm{GenMC}(n)$, it rules out the worst obstruction (a
bounded sequence of captured primes).

\paragraph{The one-hypothesis route: LMG.}
The proved chain
\lean{EM/Population/ReciprocalSum.lean}{lmg\_implies\_positive\_density\_rsd}
{\color{leanink}$\checkmark$} shows:
\[
  \mathrm{LMG} \;\Longrightarrow\; \mathrm{PositiveDensityRSD},
\]
where \textbf{LinearMeanGrowth}~(LMG) asks that
$\mathbb{E}_n[S_K(n)] \geq \kappa K$ for some $\kappa > 0$ and all
$K$ large enough (here $S_K(n) = \sum_{k<K} 1/\seq{n}(k)$ is the
partial reciprocal sum, averaged over squarefree~$n$).
LMG is a \emph{population-level} statement: it involves no specific
orbit, bypassing the orbit-specificity barrier that obstructs DSL\@.

With the additional hypothesis PairwiseStepDecorrelation~(PSD)---that
$\operatorname{Cov}[1/\seq{n}(j),\, 1/\seq{n}(k)] \to 0$ for
$j \neq k$---the stronger conclusion follows:
$\mathrm{PSD} + \mathrm{LMG} \Rightarrow \mathrm{AlmostAllSquarefreeRSD}$
(\lean{EM/Population/ReciprocalSum.lean}{dead\_end\_125\_pairwise\_revival}
{\color{leanink}$\checkmark$}).
The bridges are all machine-verified:
$\mathrm{IVB}(1/4)$ {\color{leanink}$\checkmark$},
$\mathrm{PSD} + \mathrm{IVB} \Rightarrow$ variance bound
{\color{leanink}$\checkmark$},
Chebyshev concentration {\color{leanink}$\checkmark$}.

\paragraph{CRT propagation and the first moment.}
LMG follows from
\textbf{FirstMomentStep}~(FMS):
$\mathbb{E}_n[1/\seq{n}(k)] \to \kappa$ as $X \to \infty$ for each~$k$
(\lean{EM/Ensemble/FirstMoment.lean}{FirstMomentStep}).
FMS in turn follows from \textbf{CRTPropagationStep}
(\lean{EM/Ensemble/CRT.lean}{CRTPropagationStep}):
the equidistribution of $\Prod{n}(k) \bmod q$ (among coprime residues)
propagates from step~$k$ to step~$k{+}1$.
The chain is:
\[
  \mathrm{SRE}
  \;+\; \mathrm{CRTPropagationStep}
  \;\xRightarrow{\text{\color{leanink}$\checkmark$}}\;
  \mathrm{AccumEquidist}
  \;\Longrightarrow\;
  \mathrm{FMS}(\kappa)
  \;\Longrightarrow\;
  \mathrm{LMG}
  \;\xRightarrow{\text{\color{leanink}$\checkmark$}}\;
  \mathrm{PositiveDensityRSD}.
\]
Here SRE (squarefree residue equidistribution) is the standard base
case, and AccumEquidist~$\Rightarrow$~FMS uses population
equidistribution of $\minFac$ via the Alladi--Buchstab weights.

CRTPropagationStep now reduces to a cleaner open hypothesis:
\textbf{EnsembleTransitionApprox}~(ETA,
\lean{EM/Ensemble/BackwardDynamics.lean}{EnsembleTransitionApprox}),
which says that the empirical transition probability---the fraction of
squarefree~$n$ with $\Prod{n}(k) \equiv c$ whose multiplier satisfies
$\seq{n}(k) \equiv b$---converges to $1/(q{-}1)$ for all nonzero
$b, c$.  The proved reduction
\lean{EM/Ensemble/BackwardDynamics.lean}{eta\_implies\_crt\_propagation}
{\color{leanink}$\checkmark$} shows
$\mathrm{ETA} \Rightarrow \mathrm{CRTPropagationStep}$
via the backward density decomposition: the density at step $k{+}1$ is
the matrix product of the density at step~$k$ with the transition
matrix, and when both converge to~$1/(q{-}1)$, so does the product.
The full chain
\lean{EM/Ensemble/BackwardDynamics.lean}{eta\_sre\_implies\_prsd}
{\color{leanink}$\checkmark$} composes this with the proved bridges to
give $\mathrm{ETA} + \mathrm{SRE} \Rightarrow
\mathrm{PositiveDensityRSD}$.

The structural basis for ETA is CRT multiplier invariance
(\lean{EM/Group/CRT.lean}{crt\_multiplier\_invariance}
{\color{leanink}$\checkmark$}): $\minFac(\Prod{n}(k)+1)$ depends on
the full integer, not just its residue mod~$q$, so the factoring
outcome is ``$q$-blind.''

\paragraph{Backward dynamics approach (Tao--Collatz analogy).}
The backward dynamics infrastructure is formalized in
\code{Ensemble/BackwardDynamics.lean} (577~lines).
Joint counts, transition probabilities, their basic probability axioms
(non-negativity, partition identity, sum-to-one), and the backward
density decomposition are all proved.
The approach follows Tao's backward
dynamics for the Collatz conjecture~\cite{Tao2019Collatz}: for a
target residue $a \in (\ZZ/q\ZZ)^\times$ at step~$k{+}1$,
\[
  F_{k+1}(a) = \sum_{c \in (\ZZ/q\ZZ)^\times} F_k(c) \cdot R(c, a, k, X),
\]
where $R(c, a, k, X) = 1/(q{-}1) + O(1/\log X)$ by CRT-blindness.
When $F_k$ is already uniform, so is $F_{k+1}$, giving the induction
step.  However, the absorption mechanism (Dead End~\#138)
invalidates this approach: the equidistribution target
$r/(r^2{-}1)$ per coprime class (not $1/(r{-}1)$, which neglects
the zero class of squarefree multiples of~$r$) is never sustained
for $k \geq 1$, because death events drain nonzero-class mass
permanently into the absorbing state $0 \bmod q$.
The formal chain ETA+DCTA+SRE~$\Rightarrow$~PRSD remains proved
but is vacuously true.

\paragraph{The accumulator as a commutative state.}
The EM accumulator $\Prod{n}(k) = \prod S_k$ is a squarefree integer
determined by its prime set~$S_k$.  The dynamics
$S \mapsto S \cup \{\minFac(\prod S + 1)\}$ is a well-defined map on
the lattice of finite prime sets, and the walk mod~$q$ depends on~$S$
only through $\prod S \bmod q$.  Two orbits from different starting
points that ever reach the same product mod~$q$ have approximately
identical future walks---the construction history is forgotten.

This approximate Markov property, combined with the super-exponential
growth $\Prod{n}(k) \geq 2^{2^k}$ (which makes CRT blindness
increasingly accurate), suggests that the walk mod~$q$ is effectively
ergodic from generic starting points.  Making this rigorous---showing
that the CRT error is summable along the orbit---would give a.a.\
$\mathrm{GenMC}$ for the standard sequence, without any modification
of the selection rule.

\paragraph{Comparison with DSL.}
The DSL route (Section~\ref{sec:ensemble}) proves MC for the single
orbit $n = 2$, contingent on transferring population statistics to one
deterministic trajectory.
The ensemble route here proves a density result for \emph{almost all}
starting points, contingent on CRTPropagationStep, which is a
population-level hypothesis with no orbit-specificity.
Neither subsumes the other: DSL would give MC directly; CRTPropagationStep
gives PositiveDensityRSD (which does not formally imply MC for $n = 2$).
However, the ensemble route faces a structurally easier barrier, since
its open hypothesis is a statement about averages, not about one
specific orbit.

\subsection{Weak Mullin and Analytic Weakenings}
\label{sec:weak-mullin}

The variants above change the \emph{selection rule}; this subsection
changes the \emph{conclusion}.  Instead of proving that every prime
appears ($M = \emptyset$), one can ask that the set of missing
primes is small, or that the EM reciprocal sum diverges.

\paragraph{Missing primes and Weak Mullin.}
Define the set of \emph{missing primes}
\[
M = \{p \in \mathbb{P} : p \neq \seq{2}(k) \text{ for all } k\}.
\]
MC is the assertion $M = \emptyset$.  A natural weakening asks only that
$M$ is \emph{small} in an analytic sense:

\begin{property}[\defname{WeakMullin} (\abbr{WM})]
\lean{EM/Population/WeakMullin.lean}{WeakMullin}
The reciprocal sum over missing primes converges:
$\sum_{q \in M} 1/q < \infty$.
\end{property}

This is strictly weaker than MC (which gives $M = \emptyset$ and hence
a trivially convergent empty sum) but strong enough to have a nontrivial
consequence:

\begin{property}[\defname{ReciprocalDivergence} (\abbr{RD})]
\lean{EM/Population/WeakMullin.lean}{ReciprocalDivergence}
The reciprocal sum over EM primes diverges:
$\sum_{k=0}^{\infty} 1/\seq{2}(k) = \infty$.
\end{property}

\begin{theorem}[\lean{EM/Population/WeakMullin.lean}{mc\_implies\_wm}]
$\MC \Longrightarrow \mathrm{WM}$.
\end{theorem}

\begin{theorem}[\lean{EM/Population/WeakMullin.lean}{wm\_implies\_rd}]
\label{thm:wm-rd}
$\mathrm{WM} \Longrightarrow \mathrm{RD}$.
\end{theorem}

\noindent
The proof partitions $\mathbb{P} = A \cup M$ (appeared vs.\ missing),
notes that both $\Sigma_A$ and $\Sigma_M$ converge under WM $+$ $\neg$RD,
contradicting the divergence of $\sum_{p} 1/p$.

\paragraph{Product divergence.}
The negation of WM has a quantitative consequence: $\neg\mathrm{WM}$ gives
$\sum_{q \in M} 1/(q{-}1) = \infty$, so the ``safe fraction'' product
$\prod_{q \in F}(q{-}2)/(q{-}1) \leq e^{-\sum 1/(q-1)} \to 0$ and its
reciprocal $\prod (q{-}1)/(q{-}2)$ exceeds any bound
(\lean{EM/Population/WeakMullin.lean}{not\_wm\_product\_diverges}).
The product $\prod (q{-}1)/(q{-}2)$ over missing primes measures the
``confinement cost'': each missing prime~$q$ confines the walk to $q-2$
of the $q-1$ units in $(\ZZ/q\ZZ)^\times$ (avoiding~$-1$).  If WM
fails, this confinement cost is unbounded.

\paragraph{Confined energy.}
For a missing prime $q \geq 3$, the walk avoids $-1$ entirely, so the
$N$ visits are distributed among only $q-2$ residues.  A pigeonhole
argument gives a confined-energy lower bound
(\lean{EM/Population/WeakMullin.lean}{per\_prime\_confined\_energy\_lb}):
$N^2/(q{-}2) \leq (q{-}1)\sum_a V_N(a)^2$.
The excess energy ratio $(q{-}1)/(q{-}2)$ compared to the
equidistributed baseline $N^2/(q{-}1)$ is bounded away from~$1$.
\defname{JointSVE}
(\lean{EM/Population/WeakMullin.lean}{JointSVE})---joint subquadratic
visit energy over any finite set of missing primes---implies MMCSB
(\lean{EM/Population/WeakMullin.lean}{jointSVE\_implies\_mmcsb}),
connecting the ``missing primes'' picture to the spectral route.

\paragraph{Avoidance tube and tube collapse.}
Define the \emph{tube ratio} for a finite set~$F$ of primes $\geq 3$:
\[
  \tau(F) \;=\; \prod_{q \in F} \frac{q-2}{q-1}.
\]
This measures the fraction of $(\ZZ/q\ZZ)^\times$ classes that a walk
avoiding $-1$ at each~$q$ is allowed to visit.  The tube ratio is
monotone decreasing in~$F$ and satisfies $0 \leq \tau(F) \leq 1$
(\lean{EM/Population/AvoidanceTube.lean}{tubeRatio\_monotone}
{\color{leanink}$\checkmark$},
\lean{EM/Population/AvoidanceTube.lean}{tubeRatio\_le\_one}
{\color{leanink}$\checkmark$}).

\begin{theorem}[Tube collapse ---
\lean{EM/Population/AvoidanceTube.lean}{tube\_collapse}
{\color{leanink}$\checkmark$}]
\label{thm:tube-collapse}
$\neg\mathrm{WM} \;\Longrightarrow\;$ for every $\varepsilon > 0$, there
exists a finite set~$F \subseteq M$ with $\tau(F) < \varepsilon$.
\end{theorem}

\noindent
The proof uses the inequality
$\tau(F) \leq \exp(-\sum_{q \in F} 1/(q{-}1))$
together with the divergence of $\sum_{q \in M} 1/(q{-}1)$ under~$\neg$WM\@.
For any unit~$a$ with $V_N(a) = 0$, the confined energy gives a rogue
character:

\begin{theorem}[Rogue character ---
\lean{EM/Population/AvoidanceTube.lean}{rogue\_character\_exists}
{\color{leanink}$\checkmark$}]
\label{thm:rogue-character}
If a unit $a \in (\ZZ/q\ZZ)^\times$ has $V_N(a) = 0$ (zero visits), then
there exists a nontrivial character~$\chi$ with
$\lVert S_\chi(N) \rVert^2 \geq N^2 / (q{-}2)^2$.
In particular, $\lVert S_\chi(N) \rVert \geq N/(q{-}2)$: the
character sum grows \emph{linearly}, contradicting SVE\@.
\end{theorem}

\begin{theorem}[SVE contradicts avoidance ---
\lean{EM/Population/AvoidanceTube.lean}{sve\_contradicts\_avoidance}
{\color{leanink}$\checkmark$}]
\label{thm:sve-contradicts-avoidance}
SVE $\;\Longrightarrow\;$ for all large~$N$, no unit $a$ can have
$V_N(a) = 0$.
\end{theorem}

\paragraph{Shielding lemma and spectral conspiracy.}
The \defname{shielding lemma} observes that $\seq{2}(n{+}1)$ is always
a term of the sequence, hence never a missing prime:

\begin{theorem}[Shielding lemma ---
\lean{EM/Population/InfiniteM.lean}{shielding\_lemma}
{\color{leanink}$\checkmark$}]
\label{thm:shielding}
For every~$n$, $\seq{2}(n{+}1) \notin M$.
\end{theorem}

\noindent
When a missing prime~$q$ divides $\Prod{2}(n)+1$, some strictly smaller
appeared prime shields~$q$ from capture
(\lean{EM/Population/InfiniteM.lean}{factor\_dichotomy\_strong}
{\color{leanink}$\checkmark$}).
Under~$\neg$WM, this shielding mechanism creates a \emph{spectral conspiracy}:

\begin{theorem}[Spectral conspiracy ---
\lean{EM/Population/SpectralConspiracy.lean}{spectral\_conspiracy\_landscape}
{\color{leanink}$\checkmark$}]
\label{thm:spectral-conspiracy}
$\neg\mathrm{WM}$ implies:
\begin{enumerate}[nosep]
\item $M$ is infinite
  (\lean{EM/Population/SpectralConspiracy.lean}{not\_wm\_implies\_missing\_infinite}
  {\color{leanink}$\checkmark$}).
\item The joint excess energy $\sum_{q \in M} \log\frac{q-1}{q-2}$ diverges.
\item For any finite ``guardian pool'' $G$ of appeared primes, the
  demand $\sum_{q \in F} 1/(q{-}1)$ from a large enough subset
  $F \subseteq M$ exceeds~$|G|$.
\item Arbitrarily many missing primes exist, each witnessing a nontrivial
  character with linear growth (Theorem~\textup{\ref{thm:rogue-character}}).
\end{enumerate}
\end{theorem}

\noindent
The guardian exhaustion clause~(3) formalizes the intuition that finitely
many appeared primes cannot ``explain away'' the avoidance at infinitely
many missing primes.  The three files
\code{Population/AvoidanceTube.lean} (516~lines),
\code{Population/InfiniteM.lean} (303~lines),
\code{Population/SpectralConspiracy.lean} (306~lines)
total 1{,}125~lines, all zero~\code{sorry}.

\paragraph{The shifted squarefree population.}
The EM accumulator $\Prod{2}(n)$ is squarefree
(\lean{EM/Population/WeakErgodicity.lean}{prod\_squarefree}),
so every Euclid number $\Prod{2}(n)+1$ belongs to the shifted
squarefree population
$\mathcal{S} = \{m \geq 2 : m{-}1 \text{ squarefree}\}$
(density $6/\pi^2$).  This decomposes the orbit equidistribution
problem into:
\begin{itemize}[nosep]
\item \textbf{Population Equidistribution (PE):} among elements of
  $\mathcal{S}$, the smallest prime factor is equidistributed modulo
  any prime~$q$ (provable by Selberg sieve $+$ Dirichlet).
\item \textbf{Population Transfer (PT):} the EM trajectory samples
  $\mathcal{S}$ without introducing mod-$q$ bias.
  This is the open content.
\end{itemize}
The chain PE $+$ PT $+$ EMDImpliesCME $\Longrightarrow$ MC is proved
(\lean{EM/Population/WeakErgodicity.lean}{pe\_transfer\_cme\_implies\_mc}).

\paragraph{Pairwise decorrelation and positive density.}
Dead End~\#125 proved that pairwise cancellation does not imply $k$-wise,
but \emph{variance bounds only use $k=2$}.  The formalization proves the
revival chain
(\lean{EM/Population/ReciprocalSum.lean}{dead\_end\_125\_pairwise\_revival}):
PairwiseStepDecorrelation $+$ LinearMeanGrowth $+$
ChebyshevConcentration $\Longrightarrow$ AlmostAllSquarefreeRSD\@.
A further simplification drops PSD entirely:

\begin{theorem}[\lean{EM/Population/ReciprocalSum.lean}{lmg\_implies\_positive\_density\_rsd}]
\label{thm:positive-density-rsd}
$\mathrm{LinearMeanGrowth} \;\Longrightarrow\; \mathrm{PositiveDensityRSD}$.
\end{theorem}

\noindent This is a 1-hypothesis route: LMG is strictly weaker than
CME---it requires only that the ensemble mean grows linearly, not
conditional equidistribution.

\subsection{The Function Field Analog}
\label{sec:function-field}

Replacing $\ZZ$ by $\FF_p[t]$ and primes by monic irreducible
polynomials gives the \textbf{FF-EM sequence}: $f_0 = t$,
$f_{n+1} = \text{min-degree monic irreducible factor of }F(n)+1$.
The function field Mullin conjecture (\defname{FF-MC},
\lean{EM/FunctionField/Analog.lean}{FFMullinConjecture}) asserts that
every monic irreducible eventually appears.

This setting is instructive because the Weil bound---an open
hypothesis over~$\ZZ$---is a \emph{theorem} over~$\FF_p[t]$,
eliminating all analytic number theory.  The residual difficulty is
the same \textbf{orbit-specificity barrier} as over~$\ZZ$: no known
result transfers population-level equidistribution to a specific
deterministic orbit (Dead End~\#127).

The formalization spans 17~files (${\sim}7{,}400$~lines, zero
\texttt{sorry}).  The headline results are:

\begin{keyresult}
\textbf{Unconditional results over $\FF_p[t]$}
(\lean{EM/FunctionField/Master.lean}{ff\_unconditional\_grand\_summary}):
\begin{enumerate}[nosep]
\item \textbf{Almost-all GenMixedMC}
  (\lean{EM/FunctionField/FFSieve.lean}{ff\_almost\_all\_unconditional}):
  for every prime~$p$, a density-$1$ set of squarefree starting
  polynomials satisfies GenMixedMC.  Zero open hypotheses.
\item \textbf{Stochastic FF-MC}
  (\lean{EM/FunctionField/StochasticMC.lean}{stochastic\_mc\_unconditional}):
  with any $\varepsilon > 0$ probability of non-greedy selection, every
  monic irreducible is captured.  Sharp phase transition at
  $\varepsilon = 0$.
\item \textbf{Necklace identity}
  (\lean{EM/FunctionField/NecklaceFormula.lean}{necklace\_identity\_proved}):
  $\sum_{d \mid n} d \cdot \pi_p(d) = p^n$, proved by Galois theory
  (${\sim}380$~lines).  This makes the entire ANT chain unconditional.
\item \textbf{$\Phi_3$ exclusion}
  (\lean{EM/FunctionField/AutonomousMap.lean}{ff\_neg\_one\_unreachable}):
  for $p \equiv 2 \pmod{3}$, $p \geq 5$, the value $-1$ is unreachable
  under the autonomous map $f(w) = w(w{+}1)$---a structural degree-$1$
  capture obstruction.
\item \textbf{Orbit barrier thesis}
  (\lean{EM/FunctionField/OrbitBarrier.lean}{orbit\_barrier\_thesis}):
  Dead Ends \#127, \#129, \#130 prove that every available algebraic
  tool has been applied and the residual is the same DSL/CME hypothesis
  as over~$\ZZ$.
\end{enumerate}
All clauses are machine-verified (zero \texttt{sorry}).
\end{keyresult}

\noindent
A detailed treatment---including the sieve chain, cyclotomic dead end,
three-level hierarchy, and complete dependency analysis---appears in
the companion paper~\cite{Paris2026FF}.
