%% =========================================================================
\section{Why It's Hard}
\label{sec:hard}
%% =========================================================================

\noindent\textbf{Original contribution.}\enspace
The selectability analysis, oracle barrier, and CCSB-as-frontier
argument below are new.  They explain \emph{why} the remaining
hypothesis resists both computation and existing analytic tools.

\subsection{The Selectability Perspective}
\label{sec:selectability}

The Euclid construction guarantees fresh primes at every step: every
prime factor of~$\Prod{2}(n)+1$ is new (coprime to the running product).
The difficulty of MC is not the \emph{existence} of new primes but
whether the $\minFac$ rule eventually \emph{selects} each one.  The
formalization makes this contrast precise.

\begin{theorem}[\lean{EM/Equidist/OrbitAnalysis.lean}{divisor\_not\_yet\_in\_seq}]
\label{thm:divisor-fresh}
If $p \mid \Prod{2}(n)+1$, then $\seq{2}(m) \neq p$ for all $m \leq n$.
\end{theorem}

\begin{proof}
Any $\seq{2}(m)$ with $m \leq n$ divides $\Prod{2}(n)$.  A number $\geq 2$
cannot divide both~$a$ and~$a+1$.
\end{proof}

\begin{theorem}[\lean{EM/Equidist/OrbitAnalysis.lean}{passed\_over\_persists}]
If $p \mid \Prod{2}(n)+1$ but $\seq{2}(n\!+\!1) \neq p$ (the $\minFac$ rule
chose a smaller prime), then $\seq{2}(m) \neq p$ for all $m \leq n+1$.
The prime survives to potentially divide future Euclid numbers.
\end{theorem}

\begin{theorem}[\lean{EM/Equidist/OrbitAnalysis.lean}{selectability\_extinguished}]
\label{thm:extinct}
Once $\seq{2}(m) = p$, we have $p \mid \Prod{2}(n)$ for all $n \geq m$, so
$p \nmid \Prod{2}(n)+1$ ever again.  Selectability is a one-shot resource.
\end{theorem}

\begin{definition}[\defname{InfinitelySelectable}]\label{def:inf-sel}
A prime~$p$ is \emph{infinitely selectable} if $p \mid \Prod{2}(n)+1$ for
cofinally many~$n$: $\forall\, N,\, \exists\, n \geq N,\,
p \mid \Prod{2}(n)+1$.
\end{definition}

By Theorem~\ref{thm:extinct}, MC$(p)$ and
Infinitely\-Selectable$(p)$ are mutually exclusive
(\lean{EM/Equidist/OrbitAnalysis.lean}%
{mc\_implies\_not\_infinitely\_selectable}):
a prime that enters the sequence can never be selectable again.

\begin{theorem}[\lean{EM/Equidist/OrbitAnalysis.lean}{dh\_implies\_infinitely\_selectable}]
\label{thm:dh-inf-sel}
Under DH, every prime that never appears in the sequence (with SE
satisfied) is infinitely selectable.
\end{theorem}

\paragraph{The random-factor variant is easy.}
Consider a variant of the Euclid--Mullin construction where, instead of
the smallest prime factor, one picks a \emph{random} prime factor
of~$\Prod{2}(n)+1$ at each step.  In this variant, MC follows from
classical mixing theory.  For any target
prime~$p$, the residue $r_n = \Prod{2}(n) \bmod p$ performs a
multiplicative walk on $(\ZZ/p\ZZ)^\times$.  In the random-factor
variant, each multiplier is a random element of the group; once the
multipliers generate the full group (which PRE guarantees), the walk
is a genuine random walk with full support.  By classical
equidistribution on finite groups (Diaconis--Shahshahani), the walk
mixes in $O(\log p)$ steps:
$r_n = -1$ (i.e., $p \mid \Prod{2}(n)+1$) occurs with
frequency $\to 1/(p\!-\!1)$---infinitely often with
probability~1.  When $p$ divides the Euclid number, the random rule
selects~$p$ with probability $1/|F(S_n)|$ (which may miss), but
cofinal hitting provides infinitely many chances, so~$p$ is
eventually selected.

\paragraph{The lpf variant is hard---yet structurally better.}
The actual Euclid--Mullin sequence uses
$\seq{2}(n{+}1) = \minFac(\Prod{2}(n){+}1)$, a \emph{deterministic}
function of the walk position.  Compared to the random-factor variant,
$\minFac$ has three measurable advantages:

\smallskip
\begin{center}
\begin{tabular}{l@{\quad}l@{\quad}l}
\toprule
 & \textbf{Random factor} & $\boldsymbol{\minFac}$ \\
\midrule
\textbf{Selection when $q \mid P(n){+}1$}
  & Probability $1/|F(S_n)|$ & Probability~1 \\
\textbf{Accumulator growth}
  & Faster (random $\geq \minFac$) & Slowest possible \\
\textbf{Hits on $-1$ needed}
  & Infinitely many & \textbf{One} (lethal hit) \\
\bottomrule
\end{tabular}
\end{center}

\smallskip\noindent
\emph{Selection efficiency.}\enspace Past the sieve gap, 0-accessibility
guarantees: if $q \mid \Prod{2}(n)+1$ then
$q = \minFac(\Prod{2}(n)+1)$---selection is certain, not
probabilistic.  \emph{Growth rate.}\enspace Since
$\minFac(N) \leq p$ for any factor~$p$ of~$N$, the accumulator
$\Prod{2}(n)$ grows minimally under the $\minFac$ rule, giving
exponentially more steps before it becomes astronomically large---more
chances for~$q$ to divide.  \emph{Lethality.}\enspace The inductive
bootstrap shows one single hit suffices: $\MC({<}\,q)$ ensures~$q$
is minimal whenever it divides, so hitting $-1$ once means~$q$
enters the sequence.  The random variant requires infinitely many
hits to overcome low per-hit selection probability.

\emph{MinFac is strictly better in every measurable dimension.
So why can't we prove it works?}

\paragraph{The independence trade-off.}
The answer is surgical: the random variant possesses a single property
that $\minFac$ lacks, and that property is the \emph{only thing the
proof uses}.

The property is \textbf{independence of multiplier from walk
position}.  In the random-factor variant, the multiplier at step~$n$
is selected independently of which residue class
$\Prod{2}(n) \bmod q$ occupies.  Walk position and multiplier are
independent random variables, and independence is the engine that
drives mixing.  Diaconis--Shahshahani's theorem applies whenever the
multipliers are i.i.d.\ from a distribution whose support generates
the group.

In the $\minFac$ variant, the multiplier is
$\minFac(\Prod{2}(n){+}1) \bmod q$---a deterministic function of the
full integer~$\Prod{2}(n)+1$.  The CRT multiplier invariance
(\lean{EM/Group/CRT.lean}{crt\_multiplier\_invariance}) shows
that the outcome is independent of $\Prod{2}(n) \bmod q$ at the
\emph{population} level: for generic integers, knowing the mod-$q$
residue does not predict $\minFac$.  But for the specific orbit~$n=2$,
the multiplier is a deterministic value, and mixing theorems for
deterministic sequences cannot be invoked.

The random variant thus sacrifices selection efficiency (it may
miss~$q$ even when $q$ divides) in exchange for statistical
independence.  The $\minFac$ variant has perfect selection but no
independence:

\smallskip
\begin{center}
\begin{tabular}{l@{\quad}l@{\quad}l}
\toprule
 & \textbf{Random factor} & $\boldsymbol{\minFac}$ \\
\midrule
Selection & Bad ($1/k$ chance) & Perfect \\
Mixing & Random walk theorem & Unproved \\
\bottomrule
\end{tabular}
\end{center}

\smallskip\noindent
The random variant's proof: mixing $\to$ cofinal hitting $\to$
infinitely many chances compensate for low per-hit probability.
The $\minFac$ variant needs only one hit---absurdly less than the
random variant achieves---but has no way to produce even that one
without mixing.

\paragraph{The stochastic domination failure.}
\label{sec:stoch-dom}
The natural instinct is to argue by \emph{stochastic domination}:
if the random walk visits~$-1$ cofinally, and $\minFac$ is ``better
in every way,'' then $\minFac$ should also visit~$-1$.  This argument
fails because the two processes \emph{diverge after one step}.

At step~$n$, both processes face the same Euclid number
$\Prod{2}(n)+1$ with factor set~$F(S_n)$.  The random variant picks a
uniformly random element of~$F(S_n)$; the $\minFac$ variant picks the
minimum.  After this step, the two processes have different
accumulators (they multiplied by different primes), so they face
different Euclid numbers at the next step.  The processes explore
completely different trajectories.  Domination fails because the
two walks live on different probability spaces after step~1; we
cannot compare them step-by-step.

A subtler comparison uses \emph{visit measures}.  Define
$\mu_{\mathrm{rand}}(a) = \lim_{N} \frac{1}{N}|\{n < N :
w_q^{\mathrm{rand}}(n) = a\}|$ and analogously
$\mu_{\min}(a)$ for the $\minFac$ walk.  For the random variant,
$\mu_{\mathrm{rand}} = $ uniform on $(\ZZ/q\ZZ)^\times$ by mixing.
For $\minFac$, CME/CCSB asserts $\mu_{\min} \approx$ uniform, and
DSL asserts this follows from PE\@.  The comparison
$\mu_{\min}(-1) \geq 1/(q{-}1)$ is precisely what needs proof, and is
the content of DSL\@.

\paragraph{CRT blindness as the organizing principle.}
The self-correcting feedback mechanism (\S\ref{sec:visit-eqd})
provides the intuition for \emph{why} $\mu_{\min} \approx$ uniform.
If the walk concentrates on a class~$a$, then at each visit:
(1)~the Euclid number satisfies $\Prod{2}(n)+1 \equiv a+1 \pmod{q}$;
(2)~by MinFacSelectionIndependence (Level~2 CRT,
\lean{EM/Ensemble/CRT.lean}{MinFacSelectionIndependence}),
$\minFac$ of numbers $\equiv a{+}1 \pmod{q}$ is equidistributed
mod~$q$ in the population; (3)~the next multiplier scatters the
walk away from~$a$ toward all classes.  Concentration generates
negative feedback pushing toward uniformity---proved as
infrastructure
(\code{lyapunov\_one\_step}, \code{uniform\_multiplier\_zero\_drift},
\code{group\_walk\_doubly\_stochastic}).
What remains open is whether the feedback is strong enough to force
equidistribution---i.e., Self-Correcting Drift (SCD), shown
equivalent to SVE (Dead End~\#120).

At the broadest level, the question becomes: \emph{for any
deterministic CRT-blind, 0-accessible selection rule~$\sigma$
applied to the factor set $F(S_n)$ at each step, must the resulting
walk hit~$-1$ cofinally?}  A positive answer would cover $\minFac$
as a special case (and explain the random variant a fortiori).  The
answer is \emph{no} in general---the $\ZZ/4\ZZ$ counterexample
(\lean{EM/FunctionField/CyclicWalkCoverage.lean}%
{alternating\_walk\_misses\_two}, Dead End~\#130)
exhibits a deterministic walk with generating support that
permanently misses an element.  But that counterexample uses an
adversarial selection rule, not a CRT-blind one.  The real question
is: \emph{does CRT-blindness + 0-accessibility + sieve gap +
super-exponential growth force hitting?}  This is DSL, restated in
the language of selection rules.

The inductive bootstrap (Section~\ref{sec:bootstrap}) partially
resolves the comparison:
$\MC({<}\,p)$ ensures all primes below~$p$ are in the sequence,
hence divide~$\Prod{2}(n)$, hence cannot divide~$\Prod{2}(n)+1$.
Past a computable stage, $p$ is the \emph{smallest} available factor
whenever it divides the Euclid number---reducing the $\minFac$
variant to the ``any-factor'' variant for the tail of the sequence.
But the gap between ``any-factor'' and ``random'' remains:
in the any-factor regime, \emph{selection} is solved,
yet \emph{hitting} is not, because hitting requires the population-to-orbit
transfer that DSL encodes.

\paragraph{The Marginal/Joint Barrier}
\label{sec:oracle}

The verified reductions (TailSE, CofinalEscape, QuotientDH) exhaust
what can be proved about the \emph{marginal} distribution of multiplier
residues.

\begin{theorem}[\lean{EM/Equidist/OrbitAnalysis.lean}{emfe\_iff\_tail\_se\_at}]
$\mathrm{EuclidMinFacEscape}(q) \Leftrightarrow \mathrm{TailSE}(q)$.
\end{theorem}

Even perfect per-position equidistribution of multipliers is consistent
with HH failure.  DH is a \emph{joint} statement---the
(position, multiplier) pair must hit the \emph{death curve}
$\multZ{2}{q}{n} = -\walkZ{2}{q}{n}^{-1}$---and no marginal statement can
force this.

\paragraph{The orbit chain gap.}
The cofinal orbit analysis picks one cofinal multiplier~$s_x$ per
walk position, producing a cycle $x_0 \to x_1 \to \cdots \to x_0$
in $(\ZZ/q\ZZ)^\times$.  Even when the cofinal multipliers generate
the full group, the cycle size~$k$ can be less than~$|(\ZZ/q\ZZ)^\times|$.
Example: in $\ZZ/6\ZZ$, the cycle $0 \to 1 \to 0$ has
$\langle 1, 5 \rangle = \ZZ/6\ZZ$ but misses~3.

Closing this gap requires showing that at each cofinally visited
position, \emph{multiple} multiplier classes appear---expanding the
cycle until it covers $-1$.  This is the ``specific-orbit problem'':
transferring generic equidistribution of $\minFac$ residues to the
particular EM orbit.

\subsection{The BRE Impossibility for $d \geq 3$}

\begin{remark}\label{rem:bre-impossible}
Positive escape density (PED) alone does \emph{not} imply
CCSB for characters of order $d \geq 3$.

\emph{Counterexample}: a walk on $\ZZ/3\ZZ$ that alternates between
only two of the three $d$-th roots of unity (phase-aligned escapes)
achieves positive escape density yet has walk sum
$\approx N/2 \cdot (1 + \omega) \neq o(N)$.

For $d = 2$ this degeneracy vanishes: the only non-trivial rotation
is~$-1$, so escape frequency \emph{is} the rotation distribution.
The order-2 BRE from NoLongRuns$(L)$ is proved in the formalization
(\lean{EM/Equidist/SelfCorrecting.lean}{order2\_noLongRuns\_mc}).
But for $d \geq 3$, PED constrains how often the walk rotates without
constraining the \emph{distribution} among $d-1$ non-identity
rotations.  The $\mathrm{PED} \Rightarrow \mathrm{BRE} \Rightarrow
\mathrm{CCSB}$ factorization is invalid for $d \geq 3$.

This barrier is specific to the PED route.  The CME $\to$ CCSB
reduction (\lean{EM/LargeSieve/Spectral.lean}{cme\_implies\_ccsb})
bypasses PED and BRE entirely, working for all character orders $d$
via the telescoping identity.  The $d \geq 3$ problem is therefore
not a barrier for the \emph{reduction}---only for the particular
factorization through PED.
\end{remark}

\paragraph{The Van der Corput Barrier}

The van der Corput inequality (Theorem~\ref{thm:vdc}, now fully proved
in the formalization) converts character sum bounds into autocorrelation
bounds.  Theorem~\ref{thm:shift-one} gives $R_1 = o(N)$ under the
Decorrelation Hypothesis.  VdC with $H = 1$
yields $|S_N|^2 \leq \frac{N+1}{2}(N + 2|R_1|) = N^2/2 + o(N^2)$,
hence $|S_N| \leq N/\sqrt{2}$.  This is non-trivial but \emph{not}
$o(N)$.  To get $o(N)$, one needs higher-order correlations $R_h = o(N)$
for $h \geq 2$, which requires HigherOrderDecorrelation
(Theorem~\ref{thm:hod-mc}).
The telescoping identity $\sum_n \chi(w(n))(\chi(m(n))-1) = O(1)$
is a precise structural constraint.

\paragraph{The Walk Bridge Falsity}
\label{sec:walk-bridge-false}

The BV route decomposes into two stages: sieve transfer
(BV~$\Rightarrow$~EMMultCSB, bounding \emph{multiplier} character sums)
and the walk bridge (EMMultCSB~$\Rightarrow$~MMCSB, converting multiplier
bounds to \emph{walk} bounds).  The walk bridge
\textbf{MultCSBImpliesMMCSB}
(\lean{EM/LargeSieve/Basic.lean\#L1226}{MultCSBImpliesMMCSB})
is stated as an open \code{Prop} and is \textbf{false in general}.

The obstruction is structural: the walk character sum is a
\emph{partial product} $\chi(w(n)) = \prod_{k<n} \chi(m(k))$
of the multiplier characters.  Even when the individual factors
$\chi(m(k))$ are equidistributed on the unit circle (so their
\emph{sum} cancels), their \emph{partial products} perform a
multiplicative walk whose norm grows as~$\sqrt{N}$,
not~$o(N)$.  Cancellation of sums does not imply cancellation
of cumulative products.

This negative result explains why the CME bypass
(Definition~\ref{def:cme}) is essential.  CME uses fiber
decomposition and telescoping to go directly from conditional
multiplier equidistribution to CCSB, circumventing the walk bridge
entirely.

\subsection{The Factorization Independence Heuristic}
\label{sec:hash-heuristic}

The preceding barriers explain why MC is hard to \emph{prove}.  This
subsection explains why it should be \emph{true}---and why the
formalization's sole remaining hypothesis (CME) is the precise
mathematical content of a natural intuition about factorization.

\paragraph{The information bottleneck.}
The information bottleneck of \S\ref{sec:sufficient-conditions}---$O(\log q)$
bits visible to the walk versus ${\sim}2^n$ bits determining the
multiplier---means the mutual information vanishes exponentially.
This is directly analogous to the pseudorandomness of iterated hash
functions: if~$H$ is a cryptographic hash, the sequence
$x, H(x), H(H(x)), \ldots$ is deterministic but statistically
indistinguishable from random, because the hash destroys recoverable
correlations.  For the EM sequence, integer factorization plays the
role of the hash function.  Both processes are fully deterministic
(given the seed, every term is uniquely determined), one-way
(computing forward is trivial; extracting structure from the output
is computationally hard), and \emph{de facto} uncorrelated
(consecutive terms pass every reasonable test for independence).

The analogy is heuristic, not rigorous.  SHA-256 is \emph{designed}
for pseudorandomness; $\minFac$ is not designed for anything.  We do
not claim that computational hardness of factoring implies MC---what
MC requires is a number-theoretic statement (CME), not a
complexity-theoretic one.  But the analogy explains the structure of
the problem: proving pseudorandomness of a deterministic process
requires showing that \emph{no exploitable structure exists}, which is
harder than finding structure.  This is why the conjecture resists
proof despite overwhelming heuristic evidence.

\paragraph{Negative analogies: why the selection rule matters.}
The factorization independence heuristic explains why the EM sequence
($\minFac$ variant) should contain all primes, while related sequences
do not.

\begin{itemize}[nosep]
\item The \textbf{Sylvester sequence} $s(n{+}1) = s(0) \cdots s(n) + 1$
  has density-zero prime divisors~\cite{Odoni1985}.  Its terms grow
  doubly exponentially, giving each prime only $O(1)$ chances to appear
  as a factor.  The ``hash chain'' runs too fast.

\item \textbf{Fermat numbers} $F_n = 2^{2^n} + 1$ have the even stronger
  property that $\sum 1/p$ converges over their prime divisors.  Again,
  doubly exponential growth limits opportunities.

\item The \textbf{second EM sequence} ($\mathrm{maxFac}$ variant)
  provably omits infinitely many primes~\cite{CoxVdP1968,
  PollackTrevino2014}.  Here the ``hash function'' ($\mathrm{maxFac}$
  instead of $\minFac$) produces large multipliers, causing the product
  to grow rapidly---analogous to using a hash function that amplifies
  rather than compresses.

\item The \textbf{first EM sequence} ($\minFac$ variant) keeps the
  product growing slowly---heuristically as $\exp(n^2/2)$
  (see~\S\ref{sec:lean}).  This is analogous to a hash function that
  compresses, giving exponentially many iterations within any fixed
  modulus.  The conjecture is that this compression ensures coverage.
\end{itemize}

\paragraph{What the formalization adds.}
The formalization identifies three equivalent formulations of the
factoring channel's decorrelation:
\begin{enumerate}[nosep]
\item \textbf{Conditional Multiplier Equidistribution (CME)}:
  the distribution of $\minFac(\Prod{2}(n)+1) \bmod q$, conditioned on
  $\Prod{2}(n) \equiv c \pmod{q}$, is asymptotically independent of~$c$
  (Definition~\ref{def:cme}).
\item \textbf{Decorrelation Hypothesis}: the character sum
  $\sum \chi(\multZ{2}{q}{n}) = o(N)$ for nontrivial~$\chi$---the
  multiplier residues have cancelling character sums, as independent
  random variables would.
\item \textbf{ComplexCharSumBound (CCSB)}: the walk character sums
  $\sum \chi(\walkZ{2}{q}{n}) = o(N)$ for nontrivial~$\chi$---ruling out
  permanent avoidance of any residue class.
\end{enumerate}
These are related by proved implications
($\mathrm{CME} \Rightarrow \mathrm{Dec} \Rightarrow \mathrm{PED}$)
and direct bypass ($\mathrm{CME} \Rightarrow \mathrm{CCSB}$).
Each implies MC through the verified reduction chain.  The irreducible
mathematical content is: does the factoring operation destroy enough
correlation for character sums to cancel?

\paragraph{Dead Ends as a Roadmap.}
Over 136~potential approaches were explored and found to be dead ends,
grouped into eight categories: orbit specificity~(OS),
technique mismatch~(TM), scale mismatch~(SM), circularity~(CI),
structurally false~(SF), collapse~(CO), decorrelation gap~(DG),
and aggregate gap~(AG).
Of these, 13~have formal Lean witnesses and 8~have weak-MC revival
score~$\geq 2$---meaning they contribute to weaker forms of the
conjecture despite being dead for full MC\@.
The full catalogue is in \code{Meta/DeadEnds.lean} and
Appendix~\ref{app:dead-ends}.

\paragraph{The Four-Way Blocker.}
The majority of the 136~dead ends reduce to a single meta-obstacle:
every known technique for proving equidistribution of sequences on
finite groups requires at least \emph{one} of (1)~independence of
steps, (2)~multiplicativity of the generating function,
(3)~algebraic-geometric structure (parameter families, monodromy),
or (4)~ergodic stationarity.  The EM walk satisfies \emph{none} of
these: the steps are deterministic and mutually dependent, the walk
is not a multiplicative function, the multipliers have no known
algebraic-geometric parametrization, and the non-autonomous dynamics
(a different multiplier at each step) rule out stationarity.  This
explains why classical tools---random-walk mixing, Hal\'asz's theorem,
Katz monodromy, Birkhoff averages---all fail.

\paragraph{The telescope exhausts the algebra.}
The telescoping identity $\chi(w(n{+}1)) = \chi(w(n)) \cdot
\chi(m(n))$ is the \emph{complete} algebraic content of the walk.
Only two decomposition strategies for the character sum $S_N =
\sum_{n<N} \chi(w(n))$ exist: grouping by \emph{value} (fiber
decomposition $\to$ CME) and grouping by \emph{lag} (autocorrelation
$\to$ HOD).  Every other rearrangement---Abel summation, M\"obius
inversion, Dirichlet series, block decomposition---either reduces to
one of these two or fails outright (Abel gives $O(N^2)$ remainder in
the wrong direction; M\"obius/Dirichlet require multiplicativity).
There is no third algebraic route to CCSB\@.

\smallskip\noindent
The pattern: every approach that avoids the specific EM orbit's
joint distribution either reduces to CCSB or fails.  Since CME
implies CCSB (proved), and CME decomposes as VCB~$+$~Dec
(\S\ref{sec:vcb}), the sharpest targets are now VCB~$+$~PED or CME:
conditional equidistribution of multipliers given walk position, or
the weaker proportionality condition combined with escape density.
CME is strictly weaker than CCSB and is the \emph{irreducible
analytic content}.

\paragraph{The Mathematical Landscape}

We need to prove one of these equivalent statements for every missing
prime~$q$:
\begin{itemize}[nosep]
\item \textbf{DH}: If the multipliers generate $(\ZZ/q\ZZ)^\times$,
  the walk $\walkZ{2}{q}{n} = \Prod{2}(n) \bmod q$ hits $-1$ cofinally.
\item \textbf{CCSB}: For every non-trivial character
  $\chi \colon (\ZZ/q\ZZ)^\times \to \mathbb{C}^\times$, the sum
  $\sum_{n < N} \chi(\walkZ{2}{q}{n}) = o(N)$.
\item \textbf{$d{=}2$ special case} (as a stepping stone): For every
  quadratic character~$\chi$, the $\pm 1$-valued walk character sum is
  $o(N)$.
\end{itemize}
The formalization has conclusively shown that every tool requiring
independence, classical multiplicativity, or ergodic stationarity
fails for a \emph{single} orbit.  Ensemble averaging over squarefree
starting points (Section~\ref{sec:ensemble}) bypasses this, but
introduces a new gap: MultCancelToWalkCancel (Dead End~\#117).
We need ideas that exploit what the EM walk specifically has.

\paragraph{Structural Features of the EM Walk}

The dead ends above show what does not work.  Complementarily, the EM
walk has four structural features---all proved or formalized---that no
dead-end approach has successfully exploited.  Any proof of MC will
almost certainly use at least one.

\medskip\noindent\textbf{Feature~1: Super-exponential growth.}\enspace
$\Prod{2}(n) \geq 2^n$
(\lean{EM/LargeSieve/Basic.lean}{prod\_lower\_bound\_for\_sieve}).  The
Euclid numbers $\Prod{2}(n)+1$ grow absurdly fast.  This means the
\emph{sieve level}---the threshold below which all prime factors have
been excluded---grows super-exponentially.  By step~$n$, the Euclid
number $\Prod{2}(n)+1$ is coprime to each of $\seq{2}(0), \ldots, \seq{2}(n)$,
a growing set of distinct primes.  The pool of ``available'' small
primes as factors of $\Prod{2}(n)+1$ shrinks, but the size of
$\Prod{2}(n)+1$ grows so fast that it must have enormous prime factors
most of the time.

\medskip\noindent\textbf{Feature~2: Mutual coprimality of Euclid
numbers.}\enspace For $m > n$, $\Prod{2}(m)$ is divisible by
$\seq{2}(n+1)$, which divides $\Prod{2}(n)+1$.  So
$\Prod{2}(m)+1 \equiv 1 \pmod{\seq{2}(n+1)}$: successive Euclid numbers
live in different residue classes modulo earlier sequence terms.  This
coprimality structure means the Euclid numbers cannot all ``avoid'' a
residue class in a coordinated way---their residues are forced apart
by the construction.

\medskip\noindent\textbf{Feature~3: The multiplier is the smallest
prime factor.}\enspace  This is the key constraint that everyone
mentions but nobody has quantified.  If $\walkZ{2}{q}{n} \neq -1$ (so
that $q$ does not divide $\Prod{2}(n)+1$ as the smallest factor), then
the multiplier $\multZ{2}{q}{n} = \minFac(\Prod{2}(n)+1)$ satisfies
$\multZ{2}{q}{n} \leq (\Prod{2}(n)+1)^{1/2}$.  For a number of size
${\sim}2^n$, this smallest factor could be as small as~$3$ or as
large as~${\sim}2^{n/2}$.  The $\minFac$ rule creates a deterministic
coupling between walk position and multiplier: the multiplier at
step~$n$ depends on the full value of~$\Prod{2}(n)+1$, not just its
residue.

\medskip\noindent\textbf{Feature~4: Self-correcting feedback.}\enspace
If the walk concentrates on certain residues mod~$q$---say
$\walkZ{2}{q}{n} \equiv a \pmod{q}$ for many~$n$---then
$\Prod{2}(n)+1 \equiv a+1 \pmod{q}$ for many~$n$.  The smallest prime
factor of numbers $\equiv a+1 \pmod{q}$ depends on~$a+1$, creating a
feedback loop: concentration in one residue class biases the
multiplier distribution, which in turn pushes the walk away from that
class.  This self-correcting mechanism has been formalized via a Lyapunov
function (\code{Reduction/SelfCorrecting.lean}): the one-step identity
$L(N{+}1) = L(N) + 2d_{w(N)}(N) + (q{-}2)/(q{-}1)$ telescopes to
$L(N) = 2R(N) + O(N)$, and $R(N) = o(N^2)$ (SCD) implies VE and
hence MC\@.  The zero-bias property under uniform multipliers is
proved; the orbit-specificity gap---does SCD hold for the EM
orbit?---remains open (\S\ref{sec:open}).  A weaker variant
\emph{LinearDrift} ($|R(N)| \leq CN$) also implies MC
(\lean{EM/Reduction/WeakRecurrence.lean}{linearDrift\_implies\_mc})
via a threshold-free route: LinearDrift~$\Rightarrow$~VisitGrowth%
~$\Rightarrow$~HH~$\Rightarrow$~MC, where VisitGrowth ($V(a,N) \to
\infty$ for all~$a$) bypasses the $\mathrm{FiniteMCBelow}$ threshold
entirely.  A free unconditional lower bound
$R(N) \geq -N(q{-}2)/(2(q{-}1))$ is proved
(\code{cumulativeDrift\_lower\_bound}), so only the one-sided
upper bound $R(N) \leq CN$ is needed.

\medskip
These four features---growth, coprimality, the $\minFac$ selection
rule, and self-correcting feedback---are the raw material that any
successful approach must engage with.  The dead ends above fail
precisely because they treat the walk generically (as a random walk,
or as an arbitrary multiplicative walk) rather than exploiting the
specific arithmetic of the EM construction.

%% =========================================================================
\subsection{The NAAD Conjecture}
\label{sec:naad}
%% =========================================================================

The preceding analysis---the four-way blocker, the telescope exhaustion,
the dead-end catalogue---identifies a structural gap that no existing
theory covers.  We now name this gap and state a precise conjecture.

\paragraph{Three theories that almost apply.}
Three mature mathematical theories converge on the class of problems to
which MC belongs, without covering it.

\emph{Arithmetic dynamics} (Silverman, Baker--DeMarco) studies iteration
of algebraic maps $\varphi \colon \mathbb{P}^1 \to \mathbb{P}^1$ over
number fields.  The theory requires the map to be a \emph{fixed}
polynomial or rational function; the EM map
$T(P) = P \cdot \minFac(P{+}1)$ is non-algebraic, since the multiplier
depends on the factorization of~$P{+}1$.

\emph{Multiplicative number theory} (Alladi, Hildebrand, Granville,
Soundararajan) gives precise asymptotics for the distribution of
$\minFac$, $\max\mathrm{Fac}$, $\Omega$, $\omega$ \emph{over the
population} of integers.  But it does not control these functions along
a specific deterministic orbit.

\emph{Computational pseudorandomness} (Blum--Micali, Goldreich--Levin)
studies when deterministic iteration produces outputs
indistinguishable from random.  The theory requires computational
hardness (one-wayness against polynomial-time adversaries) and operates
in the Boolean setting $\{0,1\}^n$, not the arithmetic setting~$\ZZ$.

The EM sequence sits at the intersection: an iterated map on~$\ZZ$
(arithmetic dynamics) whose multiplier is~$\minFac$ (multiplicative
number theory), whose pseudorandom behavior---if true---arises from
the opacity of integer factorization (computational pseudorandomness).

\paragraph{Arithmetic iteration maps.}

\begin{definition}[\defname{Arithmetic iteration map}]
\label{def:aim}
An \emph{arithmetic iteration map} is a function
$T \colon \ZZ_{>0} \to \ZZ_{>0}$ of the form $T(x) = x \cdot m(x)$,
where $m(x) \in \ZZ_{>1}$ is the \emph{multiplier}, determined by the
arithmetic of~$x$.  The orbit from seed~$x_0$ is the sequence
$x_0, T(x_0), T^2(x_0), \ldots$
\end{definition}

\noindent The map is \emph{non-autonomous}: the multiplier $m(x_n)$
depends on the state~$x_n$, which changes at every step.  Unlike
autonomous dynamics (a fixed map applied repeatedly), the effective
transformation changes at each iteration.

\begin{example}[\defname{Examples}]
\begin{enumerate}[nosep]
\item \textbf{Euclid--Mullin}: $T(P) = P \cdot \minFac(P{+}1)$.
\item \textbf{Largest-factor Euclid}: $T(P) = P \cdot \max\mathrm{Fac}(P{+}1)$.
  Same structure, different selection rule; provably misses
  primes~\cite{CoxVdP1968}.
\item \textbf{Random-factor Euclid}: $T(P) = P \cdot
  \mathrm{rand\-Fac}(P{+}1)$.  The walk is a genuine random walk on
  $(\ZZ/q\ZZ)^\times$; equidistribution follows from
  Diaconis--Shahshahani mixing.
\end{enumerate}
\end{example}

\noindent For a prime~$q$, the \emph{walk} is the projection
$w_n = T^n(x_0) \bmod q$ onto $(\ZZ/q\ZZ)^\times$, satisfying the
multiplicative recurrence $w_{n+1} = w_n \cdot m_n \bmod q$ where
$m_n = m(x_n) \bmod q$.  This is the walk framework of
Section~\ref{sec:walk}, now abstracted to a class of maps.

\paragraph{Four structural properties.}
We identify four properties that control the walk's behavior.  All four
are verified for the EM map (the first three are proved; the fourth
follows from standard ANT).

\begin{enumerate}[nosep]
\item \textbf{CRT-blindness} (= Property~1 of \S\ref{sec:crt-equidist}):
  $m(x) \bmod q$ is asymptotically independent of $x \bmod q$ in the
  population.  For EM, this is the algebraic fact
  \lean{EM/Group/CRT.lean}{crt\_multiplier\_invariance}.

\item \textbf{Super-exponential growth}: $T^n(x_0) \geq 2^{2^n}$ for
  $n \geq n_0$.  For EM, $\Prod{2}(n) \geq 2^{n+1}$
  (\lean{EM/LargeSieve/Basic.lean}{prod\_lower\_bound\_for\_sieve}).
  This creates the information bottleneck: the walk position uses
  $O(\log q)$ bits to describe a state with $\Omega(2^n)$ bits.

\item \textbf{Generating multipliers}: the residues
  $\{m(x_n) \bmod q\}_{n \geq 0}$ generate $(\ZZ/q\ZZ)^\times$.
  For EM, this is SubgroupEscape, proved unconditionally via PRE
  (Theorem~\ref{thm:pre}).

\item \textbf{Population equidistribution}: $m(x) \bmod q$ is
  equidistributed over $(\ZZ/q\ZZ)^\times$ in the population.  For EM,
  this follows from Alladi's formula combined with Dirichlet's theorem
  (scaffolded in \code{AlladiDensity.lean}).
\end{enumerate}

\begin{conjecture}[\defname{NAAD Equidistribution}]
\label{conj:naad}
Let $T(x) = x \cdot m(x)$ be an arithmetic iteration map satisfying
Properties~1--4 for all primes $q \geq q_0$.  Let $x_0 \in \ZZ_{>0}$
with $\gcd(x_0, q) = 1$.  Then the walk $w_n = T^n(x_0) \bmod q$
equidistributes in $(\ZZ/q\ZZ)^\times$:
\[
  \text{for every non-trivial } \chi \colon (\ZZ/q\ZZ)^\times \to
  \mathbb{C}^\times,\quad
  \frac{1}{N}\Bigl|\sum_{n < N} \chi(w_n)\Bigr| \;\to\; 0
  \quad\text{as } N \to \infty.
\]
\end{conjecture}

\begin{theorem}
NAAD Equidistribution for the EM map implies MC.
\end{theorem}

\begin{proof}
The EM map satisfies Properties~1--4.  NAAD applied with $x_0 = 2$
gives CCSB, which implies MC through the machine-verified chain
$\mathrm{CCSB} \Rightarrow \mathrm{MC}$
(\lean{EM/Equidist/SelfCorrecting.lean}{complex\_csb\_mc'}).
\end{proof}

\paragraph{The three-level gap.}
NAAD Equidistribution is the assertion that Level~3 of the CRT
hierarchy (\S\ref{sec:crt-equidist}) follows from Levels~1--2.
The conjecture's content is precisely the orbit-specificity
transfer: Properties~1--4 describe the \emph{function}~$m$;
the conjecture asks whether the \emph{orbit} inherits the function's
statistical behavior.  The entire infrastructure of this
formalization---58{,}150~lines---reduces to this single question.

\paragraph{Relation to Collatz.}
Tao~\cite{Tao2019Collatz} proved that almost all Collatz orbits attain almost
bounded values, using a random model for Syracuse random variables.
Both Collatz and EM involve non-autonomous iteration maps with CRT
structure; both have random models that behave well; both face the
transfer problem from random model to actual orbit.  The key
difference: the Collatz map is piecewise linear mod~2 (clean
algebraic structure in the 2-adic coordinate), while the EM map
involves factoring, which has no algebraic shortcut.  This makes EM
harder to analyze but also more ``pseudorandom''---the factoring
channel is a stronger scrambler than the 2-adic valuation.

\paragraph{Why NAAD requires new mathematics.}
The four-way blocker (\S\ref{sec:hash-heuristic}) applies to any
arithmetic iteration map: Properties~1--4 explicitly exclude
independence, multiplicativity, algebraic-geometric structure, and
ergodic stationarity.  NAAD Equidistribution asks whether these four
properties \emph{collectively} substitute for any single classical
assumption.  The 136~dead ends documented in the formalization
(Appendix~\ref{app:dead-ends}) confirm that no existing technique
closes this gap.

\paragraph{NAAD open problems.}
Beyond the full conjecture, several intermediate problems are
independently interesting:

\begin{enumerate}[nosep]
\item \textbf{Weak NAAD} (almost-all starting points): prove that for
  almost all squarefree~$x_0$, the walk equidistributes.  Partially
  achieved: SD~$\Rightarrow$~cancellation for a.a.\ starting points
  (\lean{EM/Ensemble/EnsemblePT.lean}{sd\_implies\_cancellation}).

\item \textbf{Quantitative NAAD} (correlation decay): prove
  $|\mathrm{Cov}(\chi(w_n), \chi(w_{n+k}))| \leq C \alpha^{-k}$
  for some $\alpha > 1$.  This would give NAAD via van der Corput
  with $|\sum \chi(w_n)| = O(\sqrt{N})$.

\item \textbf{CRTPropagationStep}: prove that accumulator
  equidistribution propagates across ensemble steps.  This is a
  population-level statement with no orbit-specificity barrier,
  amenable to CRT~$+$~Alladi.

\item \textbf{NAAD for other maps}: study variants
  $T(x) = x \cdot \max\mathrm{Fac}(x{+}1)$,
  $T(x) = x \cdot \Omega(x{+}1)$, etc.  The comparison illuminates the
  role of the selection rule: minFac gives slow growth and guaranteed
  capture (\S\ref{sec:selectability}); maxFac gives fast growth and
  provable failure.
\end{enumerate}

%% =========================================================================
\subsection{Open Problems and Paths Forward}
\label{sec:open}
%% =========================================================================

The formalization certifies the reductions above and pinpoints the
frontier.  This subsection identifies the open problems whose resolution
would close Mullin's Conjecture.

\paragraph{CCSB as the Precise Frontier}

The formalization identifies \textbf{ComplexCharSumBound} as the
irreducible analytic content.  The question:

\begin{center}
\emph{Are the walk character sums $\sum_{n<N} \chi(\walkZ{2}{q}{n})$ bounded
$o(N)$ for every non-trivial $\chi$?}
\end{center}

CCSB is a single hypothesis that implies MC with no additional
conditions.  As shown in Section~\ref{sec:character}, CCSB rules out
permanent avoidance of~$-1$: the Fourier bridge forces the hit count
at~$-1$ to be $N/(q{-}1) + o(N)$, contradicting the zero hits that
the first missing prime's walk must achieve past the sieve gap.

The walk telescoping identities (Section~\ref{sec:character}) provide
precise structural constraints.  The identity
$\sum_n \chi(w(n))(\chi(m(n))-1) = O(1)$ means that the walk sum
$S_N$ and the multiplier sum $M_N = \sum_n \chi(m(n))$ satisfy
$S_N \approx S_N + (M_N - S_N) = M_N + O(1)$ only in the crude sense;
the telescoping does \emph{not} separate them.

\paragraph{Connection to Bombieri--Vinogradov}

A Bombieri--Vinogradov type result for EM walk residues would give:
\[
\sum_{\substack{q \leq Q \\ q\text{ prime}}}
\max_{a}
\Bigl| |\{n \leq N : w(n) \equiv a \pmod{q}\}|
  - \frac{N}{q-1} \Bigr|
\;\ll\; \frac{NQ}{(\log N)^A}.
\]
For non-exceptional primes, the hit count at~$-1$ would be
$\sim N/(q{-}1) > 0$, contradicting permanent avoidance; the finitely
many exceptional primes below~$Q_0$ are handled by the inductive
bootstrap (\S\ref{sec:bootstrap}), closing the conjecture.

The difficulty is that BV applies to the set of \emph{all} primes,
not to a specific subsequence.  The EM walk is deterministic and
self-referential: the walk at step~$n$ depends on the factorization
of~$\Prod{2}(n)+1$, which depends on all previous walk values.  Standard
BV does not apply.

\paragraph{Connection to Chebotarev}

The \textbf{EffectiveKummerEscape} hypothesis asserts: for each
prime~$\ell$, there exists~$B$ such that for $q \geq B$ with
$\ell \mid q\!-\!1$, some multiplier among the first~$B$ escapes the
$\ell$-th power kernel.  This is a Chebotarev-type statement for the
Kummer extension $\mathbb{Q}(\zeta_\ell, 3^{1/\ell}, \ldots,
53^{1/\ell})$: the Frobenius at~$q$ determines which multiplier
primes are $\ell$-th power residues.

An effective Chebotarev density theorem~\cite{LagariasOdlyzko1977}
for this fixed number field
would give EKE for all but finitely many~$q$ (effectively bounded).
Combined with finite verification for the remaining~$q$, this would
prove PRE and hence SE unconditionally---but SE is
\emph{already} proved unconditionally via the elementary PRE.
The Chebotarev approach would give a stronger \emph{effective}
bound on how quickly SE kicks in, refining the density argument of
\S\ref{sec:bootstrap}.

\paragraph{The Sieve-Theoretic Approach}

\textbf{Mertens\-Escape}: for any prime~$q$ and proper
subgroup~$H$, infinitely many primes outside~$H$ exist
(Dirichlet content).
\textbf{Sieve\-Amplification}: Mertens escape should force
eventual $\minFac(\Prod{2}(n){+}1)$ escape from~$H$, via
super-exponential growth and mutual coprimality of
successive Euclid numbers.

The formally verified chain:
$\mathrm{MertensEscape} + \mathrm{SieveAmplification}
\xrightarrow{\text{proved}}
\mathrm{TailSE} \xrightarrow{\text{proved}}
\mathrm{CofinalEscape} \xrightarrow{\text{proved}}
\mathrm{QuotientDH}$.

The formalization articulates a richer sieve infrastructure in two parallel
routes:

\paragraph{Cumulative route.}
\begin{gather*}
\mathrm{PDE}
\xrightarrow{\text{Alladi}}
\mathrm{GLPFE}
\xrightarrow{\mathrm{SieveTransfer}}
\mathrm{SieveEquidist}
\xrightarrow{\text{open}}
\mathrm{NoLongRuns} \\
\xrightarrow{\text{proved}}
\mathrm{PED}
\xrightarrow{\text{open}}
\mathrm{CCSB}
\xrightarrow{\text{proved}}
\mathrm{MC}.
\end{gather*}
Here PDE is PrimeDensityEquipartition (PNT in arithmetic progressions,
a known theorem not yet in Mathlib), and GLPFE is GenericLPFEquidist
(Alladi's theorem~\cite{Alladi1977} on $\minFac$ distribution of generic integers,
also known but not formalized).  Both ends of the chain---from PDE to
GLPFE via Alladi, and from CCSB to MC via Fourier inversion---are
formally proved.

\paragraph{Window route.}
\[
\mathrm{StrongSieveEquidist}
\xrightarrow{\text{proved}}
\mathrm{NoLongRunsAt}
\xrightarrow{\text{proved}}
\mathrm{PEDAt}
\xrightarrow{\text{open}}
\mathrm{CCSB}
\xrightarrow{\text{proved}}
\mathrm{MC}.
\]
StrongSieveEquidist asserts that EM multipliers are equidistributed
within sliding windows; NoLongRunsAt and PEDAt are per-prime variants
proved by pigeonhole and block-counting respectively
(\lean{EM/Equidist/SieveTransfer.lean\#L185}{strongSieveEquidist\_noLongRunsAt},
\lean{EM/Equidist/SieveTransfer.lean\#L254}{noLongRunsAt\_ped}).

\paragraph{The genuine frontier.}

\textbf{SieveTransfer} is the critical open hypothesis: does the
equidistribution of $\minFac$ residues for generic integers transfer
to the specific EM orbit?  Everything above SieveTransfer is known
mathematics; everything below it is proved.  SieveTransfer is where
``known but not formalized'' meets ``genuinely open.''

The difficulty: for \emph{generic} $q$-rough integers,
$\minFac$ residues are equidistributed (by CRT + Mertens)---the
death channel gets its fair share of multipliers.  But the EM
Euclid numbers are not generic: each is determined by the entire
walk history.  Transferring the generic equidistribution to this
specific orbit is the open step.

\paragraph{Sieve-to-harmonic convergence.}

The sieve hierarchy (\S36--\S39 of \code{Equidist/SelfCorrecting.lean})
and the harmonic hierarchy (\S30--\S35) converge: both produce
DecorrelationHypothesis as output.  The full chain
\[
\mathrm{SieveEquidist}
\xrightarrow{\text{proved}}
\mathrm{Dec}
\xrightarrow{\text{proved}}
\mathrm{PED}
\xrightarrow[\text{sole gap}]{\text{open}}
\mathrm{CCSB}
\xrightarrow{\text{proved}}
\mathrm{MC}
\]
is formalized, with the first two arrows machine-verified
(\lean{EM/LargeSieve/Basic.lean\#L1617}{sieve\_equidist\_implies\_decorrelation},
\lean{EM/Equidist/SelfCorrecting.lean\#L139}{decorrelation\_implies\_ped}).
The sieve route achieves SieveEquidist~$\Rightarrow$~Dec via a
counting-to-character-sum bridge: SieveEquidistribution produces
\code{EMMultCharSumBound} with $Q_0 = 0$, meaning multiplier character
sums cancel for \emph{all} primes~$q$, which is exactly
DecorrelationHypothesis.  The sole remaining gap on this route is
\textbf{PEDImpliesComplexCSB}
(\lean{EM/Equidist/SelfCorrecting.lean\#L109}{PEDImpliesComplexCSB}):
does positive escape density for all primes imply walk character sum
cancellation?  Any proof of SieveEquidistribution (e.g., from PNT
in APs $+$ Alladi's theorem) would immediately yield Dec and PED
for free, isolating this single bridge as the only open step.

\paragraph{DSL: The Sole Remaining Hypothesis}

The ensemble averaging framework (Section~\ref{sec:ensemble}) collapses
the entire reduction landscape to a single open statement:
the \textbf{Deterministic Stability Lemma}~(DSL,
Property~\ref{def:dsl}).  DSL asserts that population
equidistribution of multiplier residues (a consequence of standard
ANT) implies conditional multiplier equidistribution for the
standard EM orbit ($n = 2$).  The assembly
$\mathrm{DSL} \Rightarrow \mathrm{CME} \Rightarrow \mathrm{CCSB}
\Rightarrow \mathrm{MC}$
is fully machine-verified
(\lean{EM/Reduction/Master.lean}{full\_chain\_dsl}).
DSL is the orbit-transfer principle: it bridges from population-level
statistics (which hold for almost all starting points) to the specific
orbit.

A weaker variant also suffices:

\begin{theorem}[\lean{EM/Ensemble/Structure.lean}{pe\_dsl\_hitting\_implies\_mc}]
$\mathrm{PE} + \mathrm{DSLHitting} \;\Longrightarrow\; \MC$,
\end{theorem}

\noindent where DSLHitting only requires that the walk hits~$-1$ for a
positive density of starting points, rather than full equidistribution.

\paragraph{Pairwise Decorrelation and Weak MC}

Dead End~\#125 proved that pairwise cancellation does not imply $k$-wise
(XOR counterexample), killing the CCSB+CPD~$\to$~UPE route.  However,
\emph{variance bounds only use $k=2$}: the identity
$\mathrm{Var}(\sum_j X_j) = \sum_j \mathrm{Var}(X_j) + 2\sum_{j<k}\mathrm{Cov}(X_j, X_k)$
means pairwise decorrelation suffices for concentration.  The
formalization proves this revival chain:

\begin{theorem}[\lean{EM/Population/ReciprocalSum.lean}{dead\_end\_125\_pairwise\_revival}]
\label{thm:pairwise-revival}
PairwiseStepDecorrelation $+$ LinearMeanGrowth $+$
ChebyshevConcentration $\;\Longrightarrow\;$
AlmostAllSquarefreeRSD\@.
\end{theorem}

\noindent The proved intermediate results include:
$\mathrm{Var}[1/\mathrm{genSeq}(n,k)] \leq 1/4$ (from $\mathrm{genSeq} \geq 2$),
the variance decomposition identity, and
PSD$+$IVB~$\Rightarrow$~$\mathrm{Var}[S_K] \leq (V{+}1)K$
(\code{psd\_ivb\_implies\_variance\_bound\_proved}).
The Chebyshev bridge is now proved
(\code{chebyshev\_concentration\_proved}), and PSD$+$IVB~$\Rightarrow$~variance
bound is proved, leaving exactly two open hypotheses: PSD and LMG\@.

A further simplification drops PSD entirely.  Since
$S_K(n) \leq K/2$ deterministically (each term $\leq 1/2$), a partition
argument shows that LMG alone implies a positive density of starting
points have divergent reciprocal sums:

\begin{theorem}[\lean{EM/Population/ReciprocalSum.lean}{lmg\_implies\_positive\_density\_rsd}]
\label{thm:positive-density-rsd}
$\mathrm{LinearMeanGrowth} \;\Longrightarrow\; \mathrm{PositiveDensityRSD}$.
\end{theorem}

\noindent This is a 1-hypothesis route (vs.\ 2 for density~1),
and the hypothesis LMG is strictly weaker than CME---it requires
only that the ensemble mean grows linearly, not conditional
equidistribution.

\paragraph{Function Field Analog}

Over $\FF_p[t]$, the FF-EM sequence has additional structural
advantages.  The formalization proves
(\code{FunctionField/WeakMC.lean}, 471~lines, zero \code{sorry}):
\begin{itemize}[nosep]
\item \textbf{Degree escape}
(\lean{EM/FunctionField/WeakMC.lean}{ffSeq\_degree\_tendsto\_atTop}):
$\deg(\mathrm{ffSeq}(n)) \to \infty$, unconditional from finiteness
$+$ injectivity.
\item \textbf{Capture count stabilization}: for each degree~$d$, the
number of degree-$d$ irreducibles captured eventually stabilizes.
\item \textbf{Pool partition identity}
(\lean{EM/FunctionField/WeakMC.lean}{captureCount\_plus\_missing}):
captured $+$ unsieved $=$ total, an exact accounting.
\item \textbf{FFCaptureExhaustive $\Leftrightarrow$ FFMullinConjecture}:
the two natural formulations are equivalent.
\end{itemize}
The pool partition identity gives quantitative control:
if the total number of monic irreducibles of degree~$\leq D$ is~$M$,
then among the first~$N > M$ terms, at least one has degree~$> D$.
Dead End~\#129 (FFLM structurally false---cyclotomic Galois groups)
is sidestepped entirely: degree escape uses only finiteness and
injectivity, not monodromy.

\paragraph{What Would Close the Conjecture}

The cleanest paths to MC:
\begin{enumerate}
\item \textbf{Prove DSL} (sole remaining hypothesis): show that the
  character energy of generalized EM sequences concentrates when averaged
  over squarefree starting points.  This is the weakest known hypothesis
  that implies MC, and the only one that does so without additional
  conditions.  The assembly
  DSL~$\Rightarrow$~CME~$\Rightarrow$~CCSB~$\Rightarrow$~MC is fully
  machine-verified (\lean{EM/Reduction/Master.lean}{full\_chain\_dsl}).

\item \textbf{Prove CME}: show that the multiplier
  character sum $\sum_{\substack{n < N \\ w(n) = c}} \chi(m(n))$ is
  $o(N)$ for each walk position~$c$.  CME is strictly weaker than
  CCSB, and $\mathrm{CME} \Rightarrow \mathrm{CCSB}$ is proved
  (\lean{EM/LargeSieve/Spectral.lean}{cme\_implies\_ccsb}).
  CME asks only about the \emph{conditional} distribution of
  multipliers given walk state---it does not require controlling the
  walk character sum itself.  This bypasses the $d \geq 3$ barrier
  entirely.

\item \textbf{Prove CCSB directly}: show that no non-trivial
  character sum can sustain the $\Omega(N)$ bias that permanent
  avoidance of~$-1$ would require.  The self-correcting sieve
  (concentration of EM primes in a residue class is exponentially
  self-limiting) is the strongest heuristic argument.

\item \textbf{Prove a BV-type estimate}: even an averaged bound
  over~$q$ would suffice---for $q \geq Q_0$ it rules out permanent
  avoidance, and the finitely many $q < Q_0$ are handled by the
  inductive bootstrap (\S\ref{sec:bootstrap}).

\item \textbf{Close the orbit chain gap}: show that at each cofinally
  visited walk position, at least two distinct multiplier classes
  appear.  This would force the orbit chain to expand to the full group.

\item \textbf{Prove DH directly}: show that a multiplicative walk
  on a cyclic group with a generating set of multipliers must hit
  every element cofinally.  This is a combinatorial question about
  deterministic walks.

\item \textbf{Prove SieveTransfer}: show that the EM orbit's $\minFac$
  distribution matches that of generic integers, at least on average.
  The cumulative sieve route (\S38) reduces this to known number theory
  (PNT in APs + Alladi's theorem); closing SieveTransfer gives CME
  (and hence CCSB) via the conditional multiplier equidistribution
  framework.

\item \textbf{Prove BVImpliesMMCSB or PrimeArithLSImpliesMMCSB}: the
  large sieve route (\S41--\S65) reduces MC to a transfer hypothesis.
  Given Bombieri--Vinogradov or the analytic large sieve, the remaining
  open step is showing that the EM orbit's multipliers receive their
  fair share of each residue class---in particular, that the death
  channel is not systematically avoided.  This is the same
  orbit-specificity gap as SieveTransfer, approached from a different
  mathematical toolkit.
\end{enumerate}

\paragraph{The vanishing noise route.}
A complementary perspective recasts the walk as a product of random
and deterministic steps.  At each step, $\Prod{2}(n)+1$ has a set~$F$
of prime factors; $\minFac$ selects the smallest, but the
\emph{factor set} $F \bmod q$ acts as a ``noisy generating set'' in
$(\ZZ/q\ZZ)^\times$.  The spectral gap lemma
(\lean{EM/Advanced/VanishingNoise.lean}{spectral\_gap\_with\_identity})
proves: for any generating set $S \ni 1$ in a finite abelian group
and any nontrivial character~$\chi$,
$\|\sum_{s \in S} \chi(s)\| < |S|$
(strict inequality, using strict convexity of~$\mathbb{C}$).
Deterministic walk steps---multiplying by a fixed element---preserve
this contraction
(\lean{EM/Advanced/VanishingNoise.lean}{char\_sum\_mul\_right\_norm}).
If the factor set has $|F| \geq 2$ elements in $(\ZZ/q\ZZ)^\times$
infinitely often (with non-summable spectral gap), the product
$\prod(1 - \gamma_k) \to 0$
(\lean{EM/Advanced/VanishingNoise.lean}{product\_contraction\_tendsto}),
which would give CCSB---but only if the walk actually \emph{averages}
over~$F$.  A proved counterexample
(\lean{EM/Advanced/VanishingNoise.lean}{selection\_vs\_average\_gap})
shows the gap cleanly: for any fixed selection $s \in S$, the
pointwise norm $\|\chi(s)\| = 1$ (zero contraction), even when the
average over~$S$ contracts strictly below~$1$.
Since $\minFac$ selects \emph{one specific element} of the factor set,
the spectral gap alone does not imply CCSB\@.

The formalization defines the factor set residues
(\lean{EM/Advanced/VanishingNoise.lean}{factorSetResidues})
and proves that the walk multiplier lies in it
(\lean{EM/Advanced/VanishingNoise.lean}{multZ\_in\_factorSetResidues}).
Two open hypotheses remain:
\textbf{InfinitelyManyLargeFactorSets} (that the Euclid numbers
$\Prod{2}(n)+1$ have at least two prime factors $\geq q$ for
infinitely many~$n$) and
\textbf{MinFacUnbiased} (that $\minFac$ selection is asymptotically
unbiased over the squarefree ensemble).
The latter is the orbit-specificity barrier
restated in spectral language.

\paragraph{Stochastic MC: path existence.}
What \emph{can} be proved is a weaker, combinatorial statement:
under InfinitelyManyLargeFactorSets, there \emph{exists} a selection
path through the factor sets that hits~$-1$ for every prime~$q$.
The formalization proves the complete chain:
\begin{enumerate}[nosep]
\item Define the \emph{mean character value} $\bar\chi(S) = |S|^{-1}\sum_{s \in S}\chi(s)$
  and the \emph{averaged character product} $\prod_{k < N}\bar\chi(S_k)$.
\item When $S_k$ contains elements with distinct $\chi$-values,
  $\|\bar\chi(S_k)\| < 1$
  (\lean{EM/Advanced/VanishingNoise.lean}{meanCharValue\_norm\_lt\_one\_of\_distinct}).
  If the spectral gaps $1 - \|\bar\chi(S_k)\|$ diverge in sum, the
  product tends to~$0$.
\item The \emph{character sum factorization identity}
  (\lean{EM/Advanced/VanishingNoise.lean}{char\_sum\_productMultiset})
  shows $\sum_\sigma \chi(\prod_k \sigma_k)
  = \prod_k \bigl(\sum_{s \in S_k}\chi(s)\bigr)$,
  where $\sigma$ ranges over all selection paths.
\item A \textbf{Fourier counting argument}
  (\lean{EM/Advanced/VanishingNoise.lean}{pathExistenceFromVanishing\_proved},
  ${\sim}160$~lines)
  proves: if the averaged products vanish for all nontrivial~$\chi$,
  then every element of~$G$ appears in the product multiset.
  The proof constructs the MulChar~$\leftrightarrow$~$(G \to^* \mathbb{C}^\times)$ bridge,
  derives character orthogonality for abstract groups, and applies the
  Fourier identity $|G| \cdot \operatorname{count}(a, M)
  = \sum_\chi \overline{\chi(a)}\cdot(\sum_{g \in M}\chi(g))$.
\end{enumerate}
This gives a machine-verified ``stochastic MC'': if the factor sets
of $\Prod{2}(n)+1$ have sufficiently many distinct residues mod~$q$
infinitely often, then \emph{some} selection from those factor sets
reaches every residue class---but not necessarily the $\minFac$ selection.
An independent two-point formulation
(\lean{EM/Advanced/VanishingNoise.lean}{twoPointCharValue\_norm\_lt\_one})
shows that even a tiny $\varepsilon$-bias coin flip between two factors
with distinct character values gives strict spectral contraction, using
strict convexity of the unit ball in~$\mathbb{C}$.

\paragraph{Self-consistent $\varepsilon$-walk.}
The stochastic MC above is \emph{physically inconsistent}: the factor
sets are taken from the $\minFac$ walk, but the selections do not feed
back into the accumulator.  The formalization also defines a
\textbf{self-consistent $\varepsilon$-walk}
(\lean{EM/Advanced/VanishingNoise.lean}{epsWalkProd})
in which the chosen factor---either $\minFac(P(n){+}1)$ or
$\operatorname{secondMinFac}(P(n){+}1)$---becomes the actual multiplier,
updating the accumulator and producing a new Euclid number at the next
step.  A decision sequence $\sigma : \mathbb{N} \to \mathrm{Bool}$
determines the choice; the all-true sequence recovers the EM walk
(\lean{EM/Advanced/VanishingNoise.lean}{epsWalkProd\_emDecision}).
The function $\operatorname{secondMinFac}(n)$ (second-smallest prime
factor) is defined and proved prime, divisor of~$n$, and
$\geq \minFac(n)$.

This self-consistent walk generates a \emph{binary branching tree}
of depth~$N$ with $2^N$ leaves, where each internal node has its own
path-dependent factor set.  The \emph{tree character sum}
(\lean{EM/Advanced/VanishingNoise.lean}{treeCharSum}),
a weighted sum of $\chi(\text{leaf position})$ over all leaves, is proved
$\leq 1$ in norm.  Unlike the product multiset of Parts~9--15, this sum
does \textbf{not} factorize as $\prod_k \bar\chi(S_k)$: the left and
right subtrees start from different accumulators, so their character sums
differ.  The open hypothesis
\defname{TreeContractionHypothesis}
asserts that the tree character sum tends to zero for all nontrivial~$\chi$;
it is strictly weaker than DSL (which requires the deterministic path to
mix) but strictly stronger than the product approximation.

\paragraph{Does the Walk Hit $-1$?}

The irreducible open question is not whether DH holds (cofinal
hitting is a convenient sufficient condition) but whether the walk
hits~$-1$ \emph{at least once} past the sieve gap.  The Single Hit
Theorem (Theorem~\ref{thm:single-hit}) shows that a single hit at
each prime suffices for MC; DH, CCSB, and CME are strategies for
producing that hit.

\paragraph{Evidence for.}
Three independent lines of evidence suggest the walk does hit~$-1$
(see also Remark~\ref{rem:why-dsl}).
(1)~\emph{The information bottleneck}: $\Prod{2}(n)$ has
${\sim}2^n$ bits, but the walk position carries only $O(\log q)$ bits.
The smallest prime factor of $\Prod{2}(n)+1$ depends on the full
value, not on its residue mod~$q$, so conditioning on walk position
constrains almost nothing about the factoring outcome---the
conditional and unconditional $\minFac$ distributions should agree.
(2)~\emph{Ensemble evidence}: for almost all squarefree starting
points, CME holds (the proved SD~$\Rightarrow$~cancellation chain).
The standard orbit $n = 2$ has no special arithmetic property that
would make it an outlier.  DSL is a typicality assertion, not a
structural miracle.
(3)~\emph{Analogy with Artin's conjecture}: Artin's conjecture
(that every non-square integer is a primitive root for infinitely
many primes) has the same orbit-specificity structure and is believed
true; Hooley~\cite{Hooley1967} proved it conditional on GRH.  The
walk-hitting question is the analogous statement for the EM walk
and would follow from an analogous uniformity hypothesis.

\paragraph{Evidence against.}
Two features of the EM sequence give pause.
(1)~\emph{Cox--van der Poorten}~\cite{CoxVdP1968}: the ``largest factor'' variant of
the Euclid--Mullin sequence provably misses primes.  The EM
sequence's completeness is not a soft consequence of the Euclid
construction but depends sensitively on the $\minFac$ selection
rule.  This fragility means heuristic arguments
(``it should work because Euclid numbers have many factors'') are
not reliable.
(2)~\emph{The $d \geq 3$ barrier}: the formalization proves that
the most natural route from multiplier escape density to walk
character sum cancellation (PED~$\Rightarrow$~BRE~$\Rightarrow$~CCSB) is
\emph{impossible} for character orders $d \geq 3$.  This is not
evidence against the hit, but it shows that producing one
requires mechanisms beyond the simplest equidistribution
framework.  The CME bypass sidesteps this barrier, but the barrier's
existence means any proof must be genuinely subtle.

\paragraph{Assessment.}
We believe the walk does hit~$-1$ at every prime, primarily because
the self-correcting sieve mechanism provides a concrete dynamical
reason (not merely a probabilistic heuristic) for the walk to
eventually hit~$-1$.  The strongest form of this belief: CME should
hold because the EM multipliers, conditioned on walk position, have
no arithmetic reason to avoid the death channel systematically.
But we acknowledge that no existing technique can prove this.
The formalization has exhaustively explored all known proof strategies
across 136~dead ends: the $d \geq 3$ barrier, the MultCancelToWalkCancel
counterexample (Dead End~\#117), the cofactor identity's failure to
escape the Marginal/Joint Barrier (Dead End~\#115), and the exhaustion
of all four DSL proof directions (algebraic identities, cofactor
completeness, exponential sums, sum-product theory) collectively show
that the proof, when found, will require genuinely new mathematical
techniques---likely exploiting the specific arithmetic of the
$\minFac$ rule on super-exponentially growing squarefree numbers in
ways that current analytic number theory does not.

One direction remains genuinely untried: exploiting the
\emph{arithmetic of the seed} $x_0 = 2$.
The multiplicative order $\mathrm{ord}_q(2)$ is large for most
primes~$q$ (Artin's conjecture, proved under GRH by Hooley), and
character sums $\sum \chi(g^n)$ cancel when $\mathrm{ord}(g)$ is
large (Konyagin--Shparlinski).  This is exactly the pattern CCSB
demands.  The fatal gap: the walk position
$\Prod{2}(n) \bmod q = 2 \cdot \seq{2}(1) \cdots \seq{2}(n) \bmod q$
is \emph{not} the geometric progression $2^n \bmod q$---the
varying multipliers $\seq{2}(k)$ destroy the structure that makes
exponential-sum bounds work (Dead End~\#135).  No mechanism has been
found to convert Artin-type information about $\mathrm{ord}_q(2)$
into walk equidistribution through a route that avoids Fourier
analysis and Cauchy--Schwarz (Dead End~\#136).  Whether such a
mechanism exists is, in our view, the most promising open question.

